### 2015-09-29 untitled.tex
edited untitled.tex

\section{Introduction} 
This report presents some discussions to explain how to use the period enforcer algorithm proposed by Rajkumar \cite{Rajkumar_1991}

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 \section{Introduction} 
-This report presents some discussions to explain how to use the period enforcer algorithm proposed by Rajkumar \cite{Rajkumar_1991}
+This report presents some discussions to explain how to use the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle self-suspending real-time tasks.
+The report explains briefly the underlying assumption of the period enforcer algorithm, its limitations, and how it should be correctly used to handle self-suspending sporadic real-time tasks.
+
+Self-suspending tasks can be classified into two models: \emph{dynamic} self-suspension and \emph{segmented} (or \emph{multi-segment}) self-suspension models. 
+The dynamic self-suspension sporadic task model characterizes each
+task $\tau_i$ as a $4$-tuple $(C_i,S_i,T_i,D_i)$: $T_i$ denotes the minimum inter-arrival time (or period) of $\tau_i$, $D_i$ is the relative deadline,
+$C_i$ denotes the upper bound on total execution time of each job of $\tau_i$,
+and $S_i$ denotes the upper bound on total suspension time of each job of $\tau_i$.  In addition to the above $4$-tuple, the segmented sporadic task model further 
+characterizes the computation segments and suspension intervals as an array
+$(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals.
+
+The basic notation in the report in \cite{Raj:suspension1991} is a \emph{deferrable task}.
+In Section 3 in the report in \cite{Raj:suspension1991}, it is assumed that "\emph{Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
+hat is, a task τi executes for Ci units with no suspensions once it begins execution.
+Any task that does suspend after it executes for a while can be considered to be two or more tasks each with its own worst- case execution time.
+The only difference is that if a task $\tau_i$ is split into two tasks $\tau_i'$ followed by $\tau_i''$, then $\tau_i''$ has the same deadlines as $\tau_i′$.
+}".
+
+The period enforcer algorithm can only be applied for the segmented case.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 The basic notation in the report in \cite{Raj:suspension1991} is a \emph{deferrable task}.
-In Section 3 in the report in \cite{Raj:suspension1991}, it is assumed that "\emph{Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
+In Section 3 in the report in \cite{Raj:suspension1991}, it is assumed that "\emph{With deferred execution, a task $\tau_i$ can execute its $C_i$ units of execution in discrete amounts $C_i^1, C_i^2$, $\cdots$ with suspension inbetween $C_i^j$ and $C_i^{j+1}$.\footnote{The notations are alerted here to be consistent.} Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
 hat is, a task τi executes for Ci units with no suspensions once it begins execution.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 
+The deferrable execution of a deferrable task imposes a scheduling penalty because one instance of a deferrable task can first defer defer its execution by some amount of time and for the next instance without any deferrable execution.
+The purpose of the period enforcer algorithm is to reduce such penalty.
+One of the main contributions in the report in \cite{Raj:suspension1991} is its Theorem 5, as follows:
+
+\begin{quote}
+Theorem 5 \cite{Raj:suspension1991}: A deferrable task that is schedulable under its worst-case conditions is also schedulable under the period enforcer algorithm.
+\end{quote}
+
 The period enforcer algorithm can only be applied for the segmented case.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 \begin{quote}
-Theorem 5 \cite{Raj:suspension1991}: A deferrable task that is schedulable under its worst-case conditions is also schedulable under the period enforcer algorithm.
+{\bf Theorem 5} \cite{Raj:suspension1991}: A deferrable task that is schedulable under its worst-case conditions is also schedulable under the period enforcer algorithm.
 \end{quote}

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 In Section 3 in the report in \cite{Raj:suspension1991}, it is assumed that "\emph{With deferred execution, a task $\tau_i$ can execute its $C_i$ units of execution in discrete amounts $C_i^1, C_i^2$, $\cdots$ with suspension inbetween $C_i^j$ and $C_i^{j+1}$.\footnote{The notations are alerted here to be consistent.} Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
-hat is, a task τi executes for Ci units with no suspensions once it begins execution.
+That is, a task $\tau_i$ executes for $C_i$ units with no suspensions once it begins execution.
 Any task that does suspend after it executes for a while can be considered to be two or more tasks each with its own worst- case execution time.
@@
 One of the main contributions in the report in \cite{Raj:suspension1991} is its Theorem 5, as follows:
-
 \begin{quote}
@@
 
-The period enforcer algorithm can only be applied for the segmented case.
+Theorem 5 in \cite{Raj:suspension1991} provides a very positive result to handle deferrable task systems.
+That is, if a deferrable task system can be schedulable by using a fixed-priority scheduling policy, the sporadic enforcer algorithm also provides a feasible schedule.
+However, recall that the deferrable task systems are based on the assumption that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
+If we would like to use this result to handle segmented self-suspending task systems, the first question we have to answer is \emph{"What is the corresponding deferrable task system of a sporadic segmented self-suspending task system?"}.
+Unfortunately, the report in \cite{Raj:suspension1991} did not provide any explanation about the above question.
+In this report, we will explain three concepts:
+\begin{itemize}
+\item First, there exists (at least) a sporadic segmented self-suspending task system that is schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using period enforcer algorithm is not feasible.
+This shows that the result in Theorem 5 in \cite{Raj:suspension1991} has to be carefully used only based on the schedulability of \emph{deferrable task systems} instead of \emph{self-suspending task systems}.
+\item Second, deriving the corresponding deferrable task system of a sporadic segmented self-suspending task system is an open problem to be done in polynomial time.
+The recent findings by Netlissen et al.
+\cite{ecrts15nelissen} can be applied, but it takes exponential time.
+\item Third, the result in Theorem 5 in \cite{Raj:suspension1991} can be very useful for handling self-suspending task systems if there exist \emph{efficient} schedulability tests of the corresponding deferrable task systems or the period enforcer algorithm.
+However, there is no such schedulability test up to now.
+\end{itemize}
+
+This report will explain
+
+\section{Mismatche of Self-Suspending Tasks and Deferrable Tasks}

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 
-This report will explain
+\section{Schedulable Self-Suspending Tasks but Unschedulable Period Enforcer}
 
-\section{Mismatche of Self-Suspending Tasks and Deferrable Tasks}
+Let's consider a system with $2$ tasks.
+Let $C_1 = 2$ and $T_1=D_1=10$ (without self-suspension).
+Let $C_{2,1} = 1, S_{2,1} = 6, C_{2,2}=1, T_2=D_2=11$.
+Suppose that we use the rate-monotonic priority assignment, i.e., $\tau_1$ has higher priority than $\tau_2$.
+This task system is feasible without any enforcement since at most one computation segment of a job of $\tau_2$ can be affected by $\tau_1$.
+The worst-case response time of task $\tau_2$ is hence $10$.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 
-Let's consider a system with $2$ tasks.
+Let's consider a task system with $2$ tasks.
 Let $C_1 = 2$ and $T_1=D_1=10$ (without self-suspension).
@@
 The worst-case response time of task $\tau_2$ is hence $10$.
+
+Figure \ref{fig:example} is the resulting schedule by using the period enforcer algorithm.
+The first job of task $\tau_2$ arrives at time $0$ is executed as if there is no period enforcer.
+The second job of task $\tau_2$ requires some attention to understand how the period enforcer algorithm works.
+Note that the first computation segment of task $\tau_2$ does not suffer from any deferrable execution (it is postponed due to higher-priority block).
+Therefore, it is always placed to the ready queue and scheduled as early as possible.
+
+Let's now look at the computation segment $C_2^2$ of the second job of task $\tau_2$.
+This segment is placed back to the ready queue at time $19$.
+If there is no period enforcer, this computation segment should be executed from $19$ to $20$.
+However, the period enforcer algorithm intentionally defers this computation segment to ensure \emph{the periodic behaviour} of the computation segment $C_2^2$.
+Therefore, this computation segment is \emph{activated} to the ready queue at time $9+T_2=9+11=20$.
+This results in a deadline miss.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 Figure \ref{fig:example} is the resulting schedule by using the period enforcer algorithm.
-The first job of task $\tau_2$ arrives at time $0$ is executed as if there is no period enforcer.
-The second job of task $\tau_2$ requires some attention to understand how the period enforcer algorithm works.
+The first job of task $\tau_2$ (that arrives at time $0$) is executed as if there is no period enforcer.
+The second job (that releases at time $11$) of task $\tau_2$ requires some attention to understand how the period enforcer algorithm works.
 Note that the first computation segment of task $\tau_2$ does not suffer from any deferrable execution (it is postponed due to higher-priority block).
@@
 Let's now look at the computation segment $C_2^2$ of the second job of task $\tau_2$.
-This segment is placed back to the ready queue at time $19$.
-If there is no period enforcer, this computation segment should be executed from $19$ to $20$.
-However, the period enforcer algorithm intentionally defers this computation segment to ensure \emph{the periodic behaviour} of the computation segment $C_2^2$.
+This segment is placed back to the ready queue at time $19$ after its self-suspension.
+Without period enforcer, this computation segment should be executed from $19$ to $20$.
+However, the period enforcer algorithm intentionally defers this computation segment to ensure \emph{the periodic demand} of the computation segment $C_2^2$.
 Therefore, this computation segment is \emph{activated} to the ready queue at time $9+T_2=9+11=20$.
-This results in a deadline miss.
+This results in a deadline miss.\footnote{By using the notations of \cite{Raj:suspension1991}, $ET_{2,1}^2=9, ET_{2,2}^2 = \max\{9+11, 19\}=20$.}

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 Therefore, this computation segment is \emph{activated} to the ready queue at time $9+T_2=9+11=20$.
-This results in a deadline miss.\footnote{By using the notations of \cite{Raj:suspension1991}, $ET_{2,1}^2=9, ET_{2,2}^2 = \max\{9+11, 19\}=20$.}
+This results in a deadline miss.\footnote{By using the notations in \cite{Raj:suspension1991}, we have $ET_{2,1}^2=9, ET_{2,2}^2 = \max\{9+11, 19\}=20$.
+The superscript $^2$ is for denoting the second computation segment of task $\tau_2$.}

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 The superscript $^2$ is for denoting the second computation segment of task $\tau_2$.}
+
+Therefore, the above example shows that there exists (at least) a sporadic segmented self-suspending task system that is schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using period enforcer algorithm is not feasible.
+In this example, it is also clear that we can pessimistically convert the two computation segments of task $\tau_2$ into two deferrable tasks $\tau_2^1$ and $\tau_2^2$.
+The deferrable time of task $\tau_2^1$ is $0$, whereas the deferrable time of task $\tau_2^2$ is at most \emph{$9$}.
+With the above conversion, we can clearly see that the deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable by the rate-monotonic priority assignment since we can release a job of $\tau_1$ together with a job of $\tau_2^2$ (after its worst-case deferrable time).

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 $(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals.
+For the simplicity of presentation, we will implicitly assume that a task $\tau_i$ always starts with a computation segment.
+The arguments can be easily extended to handle tasks that start with self-suspensions.
 
@@
 Unfortunately, the report in \cite{Raj:suspension1991} did not provide any explanation about the above question.
-In this report, we will explain three concepts:
+In this report, we will explain two concepts:
 \begin{itemize}
@@
 \cite{ecrts15nelissen} can be applied, but it takes exponential time.
-\item Third, the result in Theorem 5 in \cite{Raj:suspension1991} can be very useful for handling self-suspending task systems if there exist \emph{efficient} schedulability tests of the corresponding deferrable task systems or the period enforcer algorithm.
-However, there is no such schedulability test up to now.
 \end{itemize}
@@
 \section{Schedulable Self-Suspending Tasks but Unschedulable Period Enforcer}
+\label{sec:unschedulable}
 
@@
 With the above conversion, we can clearly see that the deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable by the rate-monotonic priority assignment since we can release a job of $\tau_1$ together with a job of $\tau_2^2$ (after its worst-case deferrable time).
+
+\section{How to Convert to the Corresponding Deferrable Task Set}
+\label{sec:convert}
+
+The example in Section \ref{sec:unschedulable} can be easily converted to a corresponding deferrable task set.
+However, it is in general not an easy problem to convert a self-suspending task system into a deferrable task set.
+Suppose that the system has $k-1$ ordinary sporadic tasks and only one segmented self-suspending task $\tau_k$.
+Converting a computation segment into a deferrable task requires to derive the \emph{worst-case resume time of a computation segment}, denoted as $R_k^j$ for the $j$-th computation segment of task $\tau_k$.
+Suppose that the worst-case response time of the $j$-th computation segment of task $\tau_k$ is $W_k^j$.
+It is not difficult to see that $R_k^1=0$ and $R_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k-1$.
+So, we need to derive the worst-case response times of the computation segments of task $\tau_k$.
+
+With the above discussions, for the most simple case defined above, the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems.
+It has been recently shown by Nelissen et al.
+\cite{ecrts15nelissen} that calculating the worst-case response time of such a problem is a challenging problem.
+Nelissen et al.
+\cite{ecrts15nelissen} explain misconceptions conducted by using wrong assumptions in the critical instant theorem.
+However, their approaches require exponential time complexity to derive $W_k^{j}$.
+
+\section{Concluding Remarks}
+
+This report explains how to use the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle segmented self-suspending real-time tasks.
+One key assumption in the report in \cite{Raj:suspension1991} is that the deferrable task systems are based on the assumption that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
+This creates some mismatches between the original sporadic task system and the corresponding deferrable task system.
+The report in  \cite{Raj:suspension1991} did not explain how to convert a segmented self-suspending task system to a corresponding deferrable task system.
+We show in this report that this is non-trivial and the main claim of Theorem 5 in \cite{Raj:suspension1991} does not reflect the schedulability of the original self-suspending task system.
+
+Nevertheless, the result in Theorem 5 in \cite{Raj:suspension1991} can be very useful for handling self-suspending task systems if there exist \emph{efficient} schedulability tests of the corresponding deferrable task systems or the period enforcer algorithm.
+However, such tests do not exist yet and remain as open problems.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 
-The deferrable execution of a deferrable task imposes a scheduling penalty because one instance of a deferrable task can first defer defer its execution by some amount of time and for the next instance without any deferrable execution.
+The deferrable execution of a deferrable task imposes a scheduling penalty because one job of a deferrable task can first defer defer its execution by some amount of time and for the next job without any deferrable execution.
 The purpose of the period enforcer algorithm is to reduce such penalty.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 Theorem 5 in \cite{Raj:suspension1991} provides a very positive result to handle deferrable task systems.
-That is, if a deferrable task system can be schedulable by using a fixed-priority scheduling policy, the sporadic enforcer algorithm also provides a feasible schedule.
+That is, if a deferrable task system can be schedulable by using a fixed-priority scheduling policy, the perioc enforcer algorithm also provides a feasible schedule.
 However, recall that the deferrable task systems are based on the assumption that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
@@
 \begin{itemize}
-\item First, there exists (at least) a sporadic segmented self-suspending task system that is schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using period enforcer algorithm is not feasible.
+\item First, there exists (at least) a sporadic segmented self-suspending task system that is schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using the period enforcer algorithm is not feasible.
 This shows that the result in Theorem 5 in \cite{Raj:suspension1991} has to be carefully used only based on the schedulability of \emph{deferrable task systems} instead of \emph{self-suspending task systems}.
@@
 Figure \ref{fig:example} is the resulting schedule by using the period enforcer algorithm.
-The first job of task $\tau_2$ (that arrives at time $0$) is executed as if there is no period enforcer.
+The first job of task $\tau_2$ (that arrives at time $0$) is executed as if there is no period enforcement.
 The second job (that releases at time $11$) of task $\tau_2$ requires some attention to understand how the period enforcer algorithm works.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 This shows that the result in Theorem 5 in \cite{Raj:suspension1991} has to be carefully used only based on the schedulability of \emph{deferrable task systems} instead of \emph{self-suspending task systems}.
-\item Second, deriving the corresponding deferrable task system of a sporadic segmented self-suspending task system is an open problem to be done in polynomial time.
+\item Second, deriving the corresponding deferrable task system of a sporadic segmented self-suspending task system in polynomial time is an open problem.
 The recent findings by Netlissen et al.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 Note that the first computation segment of task $\tau_2$ does not suffer from any deferrable execution (it is postponed due to higher-priority block).
-Therefore, it is always placed to the ready queue and scheduled as early as possible.
+Therefore, it is always placed to the ready queue immediately when it arrives and scheduled as early as possible.
 
@@
 This segment is placed back to the ready queue at time $19$ after its self-suspension.
-Without period enforcer, this computation segment should be executed from $19$ to $20$.
+Without period enforcement, this computation segment should be executed from $19$ to $20$.
 However, the period enforcer algorithm intentionally defers this computation segment to ensure \emph{the periodic demand} of the computation segment $C_2^2$.
@@
 
-Therefore, the above example shows that there exists (at least) a sporadic segmented self-suspending task system that is schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using period enforcer algorithm is not feasible.
+Therefore, the above example shows that there exists (at least) a sporadic segmented self-suspending task system that is schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using the period enforcer algorithm is not feasible.
 In this example, it is also clear that we can pessimistically convert the two computation segments of task $\tau_2$ into two deferrable tasks $\tau_2^1$ and $\tau_2^2$.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 The example in Section \ref{sec:unschedulable} can be easily converted to a corresponding deferrable task set.
-However, it is in general not an easy problem to convert a self-suspending task system into a deferrable task set.
+However, it is in general not an easy problem to convert a self-suspending task system into a deferrable task set \emph{precisely}.
+We can demonstrate the difficulty by looking at a special case.
 Suppose that the system has $k-1$ ordinary sporadic tasks and only one segmented self-suspending task $\tau_k$.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 It has been recently shown by Nelissen et al.
-\cite{ecrts15nelissen} that calculating the worst-case response time of such a problem is a challenging problem.
+\cite{ecrts15nelissen}  calculating the worst-case response time in the above case is already a very challenging problem.
 Nelissen et al.
-\cite{ecrts15nelissen} explain misconceptions conducted by using wrong assumptions in the critical instant theorem.
-However, their approaches require exponential time complexity to derive $W_k^{j}$.
+\cite{ecrts15nelissen} identify several misconceptions used in the literature for this problem.
+After excluding those misconceptions, deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very difficult problem.
+Due to \cite{ecrts15nelissen}, the only existing solution to derive the exact $W_k^{j}$ requires exponential time complexity.
 

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 After excluding those misconceptions, deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very difficult problem.
-Due to \cite{ecrts15nelissen}, the only existing solution to derive the exact $W_k^{j}$ requires exponential time complexity.
+Due to \cite{ecrts15nelissen}, the only existing solution to derive the exact $W_k^{j}$ (and hence $R_k^j$) requires exponential time complexity.
 
@@
 This report explains how to use the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle segmented self-suspending real-time tasks.
-One key assumption in the report in \cite{Raj:suspension1991} is that the deferrable task systems are based on the assumption that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
+One key assumption in the report in \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
 This creates some mismatches between the original sporadic task system and the corresponding deferrable task system.

### 2015-09-29 untitled.tex
edited untitled.tex

@@
 Due to \cite{ecrts15nelissen}, the only existing solution to derive the exact $W_k^{j}$ (and hence $R_k^j$) requires exponential time complexity.
+And, also recall the example in \ref{sec:unschedulable}.
+Even if the conversion is done precisely, the deferrable task system can be more pessimistic than the original self-suspending task system with respect to the schedulability.
 
@@
 One key assumption in the report in \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
-This creates some mismatches between the original sporadic task system and the corresponding deferrable task system.
+This creates some mismatches between the original self-suspending task system and the corresponding deferrable task system.
 The report in  \cite{Raj:suspension1991} did not explain how to convert a segmented self-suspending task system to a corresponding deferrable task system.

### 2015-10-03 untitled.tex
edited untitled.tex

@@
 \section{Introduction} 
-This report presents some discussions to explain how to use the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle self-suspending real-time tasks.
-The report explains briefly the underlying assumption of the period enforcer algorithm, its limitations, and how it should be correctly used to handle self-suspending sporadic real-time tasks.
+This report revists the \emph{period enforcer} algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle self-suspending real-time tasks.
+The report briefly reviews the period enforcer algorithm, explains its underlying assumptions and limitations, and discusses how it may be analyzed to correctly determine the schedulability of self-suspending sporadic real-time tasks subject to period enforcement.
 
-Self-suspending tasks can be classified into two models: \emph{dynamic} self-suspension and \emph{segmented} (or \emph{multi-segment}) self-suspension models. 
+The main contributions of this report are two observations that have not been previously reported in the real-time literature:
+
+\begin{enumerate}
+	\item period enforcement can render self-suspending tasks sets unschedulable that are schedulable otherwise (i.e., period enforcement can induce deadline misses); and
+	\item with current techniques, schedulability analysis of the period enforcer algorithm requires a task set transformation that is subject to exponential time complexity.
+\end{enumerate}
+
+\subsection{Preliminaries}
+
+To date, the real-time literature on self-suspending tasks has focused on two task models: the \emph{dynamic} and the \emph{segmented} (or \emph{multi-segment}) self-suspension model. 
 The dynamic self-suspension sporadic task model characterizes each
-task $\tau_i$ as a $4$-tuple $(C_i,S_i,T_i,D_i)$: $T_i$ denotes the minimum inter-arrival time (or period) of $\tau_i$, $D_i$ is the relative deadline,
-$C_i$ denotes the upper bound on total execution time of each job of $\tau_i$,
-and $S_i$ denotes the upper bound on total suspension time of each job of $\tau_i$.  In addition to the above $4$-tuple, the segmented sporadic task model further 
-characterizes the computation segments and suspension intervals as an array
-$(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals.
-For the simplicity of presentation, we will implicitly assume that a task $\tau_i$ always starts with a computation segment.
+task $\tau_i$ as a $4$-tuple $(C_i,S_i,T_i,D_i)$: 
+$C_i$ denotes the upper bound on the total execution time of any job of $\tau_i$,
+$S_i$ denotes the upper bound on the total self-suspension time of any job of $\tau_i$,
+$T_i$ denotes the minimum inter-arrival time (or period) of $\tau_i$, and $D_i$ is the relative deadline.
+The dynamic self-suspension model does not impose a bound on the maximum number of self-suspensions, nor does it make any assumptions as to where during a job's execution self-suspensions occur.
+
+In contrast, the segmented sporadic task model extends the above $4$-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals.
+These intervals are represented as a tuple
+$(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, which is composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals.
+For the simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment.
 The arguments can be easily extended to handle tasks that start with self-suspensions.
 
-The basic notation in the report in \cite{Raj:suspension1991} is a \emph{deferrable task}.
-In Section 3 in the report in \cite{Raj:suspension1991}, it is assumed that "\emph{With deferred execution, a task $\tau_i$ can execute its $C_i$ units of execution in discrete amounts $C_i^1, C_i^2$, $\cdots$ with suspension inbetween $C_i^j$ and $C_i^{j+1}$.\footnote{The notations are alerted here to be consistent.} Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
+The advantage of the dynamic model is that it is more flexible since it does not impose any assumptions on control flow.
+The advantage of the segmented model is that is allows for more accurate analysis.
+The period enforcer algorithm and its analysis fundamentally applies (only) to the segmented model.
+
+The central notation in Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{deferrable task}, which matches our notion of multi-segment tasks.  Specifically, Rajkumar states that:
+\begin{quote}
+With deferred execution, a task $\tau_i$ can execute its $C_i$ units of execution in discrete amounts $C_i^1, C_i^2$, $\ldots$ with suspension in between $C_i^j$ and $C_i^{j+1}$.
+\cite[Section 3]{Raj:suspension1991}\footnote{The notation has been altered here for the sake of consistency.}
+\end{quote}
+%
+Central to Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{task set transformation} that splits each deferrable task with multiple segments into a corresponding number of single-segment deferrable tasks.
+In the words of Rajkumar~\cite[Section 3]{Raj:suspension1991}:
+
+\begin{quote}
+	 Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
 That is, a task $\tau_i$ executes for $C_i$ units with no suspensions once it begins execution.
-Any task that does suspend after it executes for a while can be considered to be two or more tasks each with its own worst- case execution time.
-The only difference is that if a task $\tau_i$ is split into two tasks $\tau_i'$ followed by $\tau_i''$, then $\tau_i''$ has the same deadlines as $\tau_i′$.
-}".
+Any task that does suspend after it executes for a while can be considered to be two or more tasks each with its own worst-case execution time.
+The only difference is that if a task $\tau_i$ is split into two tasks $\tau_i'$ followed by $\tau_i''$, then $\tau_i''$ has the same deadlines as $\tau_i{{'}}$.
+\end{quote}
+%
+In other words, the transformation can be understood as splitting each self-suspending task into a matching number of non-self-suspending sporadic tasks subject to \emph{release jitter}, which can be easily analyzed with classic fixed-priority response-time analysis~\cite{ABRTW:93}.
 
-The deferrable execution of a deferrable task imposes a scheduling penalty because one job of a deferrable task can first defer defer its execution by some amount of time and for the next job without any deferrable execution.
-The purpose of the period enforcer algorithm is to reduce such penalty.
-One of the main contributions in the report in \cite{Raj:suspension1991} is its Theorem 5, as follows:
+It is well known that uncontrolled deferred execution (i.e, release jitter) can impose a scheduling penalty because of the potential for ``back-to-back'' execution~\cite{ABRTW:93}.
+That is, if a job of a deferrable task that maximally defers its execution is directly followed by a job that executes immediately without deferring its execution, then lower-priority tasks may suffer increased interference.
+
+The purpose of the period enforcer algorithm is to reduce such penalties for lower-priority tasks without detrimentally affecting the schedulability of self-suspending, higher-priority tasks.
+The latter aspect --- no detrimental effects for self-suspending tasks --- is captured concisely by Theorem 5 in the original analysis of the period enforcer algorithm \cite{Raj:suspension1991}.
 \begin{quote}
-{\bf Theorem 5} \cite{Raj:suspension1991}: A deferrable task that is schedulable under its worst-case conditions is also schedulable under the period enforcer algorithm.
+{\bf Theorem 5}: A deferrable task that is schedulable under its worst-case conditions is also schedulable under the period enforcer algorithm.  \cite{Raj:suspension1991}
 \end{quote}
 
-Theorem 5 in \cite{Raj:suspension1991} provides a very positive result to handle deferrable task systems.
-That is, if a deferrable task system can be schedulable by using a fixed-priority scheduling policy, the perioc enforcer algorithm also provides a feasible schedule.
-However, recall that the deferrable task systems are based on the assumption that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
-If we would like to use this result to handle segmented self-suspending task systems, the first question we have to answer is \emph{"What is the corresponding deferrable task system of a sporadic segmented self-suspending task system?"}.
-Unfortunately, the report in \cite{Raj:suspension1991} did not provide any explanation about the above question.
-In this report, we will explain two concepts:
-\begin{itemize}
-\item First, there exists (at least) a sporadic segmented self-suspending task system that is schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using the period enforcer algorithm is not feasible.
-This shows that the result in Theorem 5 in \cite{Raj:suspension1991} has to be carefully used only based on the schedulability of \emph{deferrable task systems} instead of \emph{self-suspending task systems}.
-\item Second, deriving the corresponding deferrable task system of a sporadic segmented self-suspending task system in polynomial time is an open problem.
-The recent findings by Netlissen et al.
-\cite{ecrts15nelissen} can be applied, but it takes exponential time.
-\end{itemize}
+\subsection{Questions answered in this report}
+
+Theorem 5 in \cite{Raj:suspension1991} provides a very positive result for the analysis of deferrable task sets subject to period enforcement.
+That is, if the corresponding transformed task set can be shown to be schedulable under fixed-priority scheduling using \emph{any} applicable analysis, then the period enforcer algorithm also provides a feasible schedule.
+
+However, note that Theorem 5 in \cite{Raj:suspension1991} applies only to the \emph{transformed} task set: recall that, in the original analysis~\cite{Raj:suspension1991}, deferrable tasks are assumed to defer their entire execution time either completely or not at all (but not parts of it).
+Therefore, if we would like to use the period enforcer algorithm to handle segmented self-suspending task sets, we first have to answer the following question: ``\emph{Given a set of sporadic segmented self-suspending tasks, what is the corresponding set of (single-segment) deferrable tasks?}'' That is, how do we convert given self-suspension segments into equivalent bounds on release jitter such that we may apply Theorem 5 to conclude that the system remains schedulable despite period enforcement?
+Unfortunately, the original proposal \cite{Raj:suspension1991} does not provide an answer to this central question.
+
+In this report, we make two observations pertaining to this question:
+\begin{enumerate}
+	\item There exist sporadic segmented self-suspending task sets that are schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using the period enforcer algorithm is not feasible.
+This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, and not in the context of the original segmented self-suspending task set.
+
+	\item Deriving a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem.
+Recent findings by Nelissen et al.
+\cite{ecrts15nelissen} can be applied, but their method takes exponential time.
+\end{enumerate}
 

### 2015-10-03 untitled.tex
edited untitled.tex

@@
 
-In this report, we make two observations pertaining to this question:
+In this report, we make two observations pertinent to this question:
 \begin{enumerate}
@@
 
-\section{Schedulable Self-Suspending Tasks but Unschedulable Period Enforcer}
+\section{Deadline Missses Caused by Period Enforcement}
 \label{sec:unschedulable}
 
-Let's consider a task system with $2$ tasks.
+In this section, we demonstrate with an example that there exist sets of segmented self-suspending sporadic tasks that both \textbf{(i)} are schedulable \emph{without} period enforcement and \textbf{(ii)} are not schedulable under the period enforcer.
+
+Consider a task system with $2$ tasks.
 Let $C_1 = 2$ and $T_1=D_1=10$ (without self-suspension).

### 2015-10-03 untitled.tex
edited untitled.tex

@@
 
-In this section, we demonstrate with an example that there exist sets of segmented self-suspending sporadic tasks that both \textbf{(i)} are schedulable \emph{without} period enforcement and \textbf{(ii)} are not schedulable under the period enforcer.
+In this section, we demonstrate with an example that there exist sporadic  segmented self-suspending task sets that both \textbf{(i)} are schedulable \emph{without} period enforcement and \textbf{(ii)} are not schedulable under the period enforcer.
 
@@
 
-\section{How to Convert to the Corresponding Deferrable Task Set}
+\section{How to Derive the Corresponding Deferrable Task Set}
 \label{sec:convert}

### 2015-10-03 untitled.tex
edited untitled.tex

@@
 Due to \cite{ecrts15nelissen}, the only existing solution to derive the exact $W_k^{j}$ (and hence $R_k^j$) requires exponential time complexity.
-And, also recall the example in \ref{sec:unschedulable}.
-Even if the conversion is done precisely, the deferrable task system can be more pessimistic than the original self-suspending task system with respect to the schedulability.
+Furthermore, as demonstrated with the example in Section \ref{sec:unschedulable}, even if the conversion is done precisely, the transformed deferrable task set can be more pessimistic than the original self-suspending task set with respect to schedulability.
 

### 2015-10-07 untitled.tex
edited untitled.tex

--- 2015-10-03 untitled.tex
@@
 
-Theorem 5 in \cite{Raj:suspension1991} provides a very positive result for the analysis of deferrable task sets subject to period enforcement.
-That is, if the corresponding transformed task set can be shown to be schedulable under fixed-priority scheduling using \emph{any} applicable analysis, then the period enforcer algorithm also provides a feasible schedule.
+Theorem 5 in \cite{Raj:suspension1991} provides a very positive result for the analysis of deferrable task sets subject to period enforcement: if the corresponding transformed task set can be shown to be schedulable under fixed-priority scheduling using \emph{any} applicable analysis, then the period enforcer algorithm also yields a correct schedule.
 
@@
 Therefore, if we would like to use the period enforcer algorithm to handle segmented self-suspending task sets, we first have to answer the following question: ``\emph{Given a set of sporadic segmented self-suspending tasks, what is the corresponding set of (single-segment) deferrable tasks?}'' That is, how do we convert given self-suspension segments into equivalent bounds on release jitter such that we may apply Theorem 5 to conclude that the system remains schedulable despite period enforcement?
-Unfortunately, the original proposal \cite{Raj:suspension1991} does not provide an answer to this central question.
 
-In this report, we make two observations pertinent to this question:
+In this report, which is motivated by the fact that the original proposal \cite{Raj:suspension1991} does not provide an answer to this central question, 
+we make two pertinent observations.
 \begin{enumerate}
@@
 
-\section{Deadline Missses Caused by Period Enforcement}
+\section{Period Enforcement Can Induce Deadline Misses}
 \label{sec:unschedulable}
 
-In this section, we demonstrate with an example that there exist sporadic  segmented self-suspending task sets that both \textbf{(i)} are schedulable \emph{without} period enforcement and \textbf{(ii)} are not schedulable under the period enforcer.
+In this section, we demonstrate with an example that there exist sporadic  segmented self-suspending task sets that both \textbf{(i)} are schedulable \emph{without} period enforcement and \textbf{(ii)} are not schedulable with period enforcement.
 
 Consider a task system with $2$ tasks.
-Let $C_1 = 2$ and $T_1=D_1=10$ (without self-suspension).
-Let $C_{2,1} = 1, S_{2,1} = 6, C_{2,2}=1, T_2=D_2=11$.
+Let $\tau_1$ denote a sporadic task without self-suspensions and parameters $C_1 = 2$ and $T_1=D_1=10$, and let $\tau_2$ denote a self-suspending task consisting of two segments with parameters  $C_{2,1} = 1$,  $S_{2,1} = 6$, $C_{2,2}=1$, and $ T_2=D_2=11$.
 Suppose that we use the rate-monotonic priority assignment, i.e., $\tau_1$ has higher priority than $\tau_2$.
-This task system is feasible without any enforcement since at most one computation segment of a job of $\tau_2$ can be affected by $\tau_1$.
-The worst-case response time of task $\tau_2$ is hence $10$.
+This task set is schedulable without any enforcement since at most one computation segment of a job of $\tau_2$ can be delayed by $\tau_1$:
+\begin{itemize}
+	\item if the first segment of a job of $\tau_2$ is interfered with by $\tau_1$, then the second segment resumes at most after $9$ time units after the release of the job and the response time of task $\tau_2$ is hence $10$; otherwise,
+	\item  if the first segment of a job of $\tau_2$ is not interfered with by $\tau_1$, then the second segment resumes at most $7$ time units after the release of the job and hence the  response time of task $\tau_2$ is at most $10$ even if the second segment is interfered with by $\tau_1$.
+\end{itemize}
 
-Figure \ref{fig:example} is the resulting schedule by using the period enforcer algorithm.
-The first job of task $\tau_2$ (that arrives at time $0$) is executed as if there is no period enforcement.
-The second job (that releases at time $11$) of task $\tau_2$ requires some attention to understand how the period enforcer algorithm works.
-Note that the first computation segment of task $\tau_2$ does not suffer from any deferrable execution (it is postponed due to higher-priority block).
-Therefore, it is always placed to the ready queue immediately when it arrives and scheduled as early as possible.
+\textbf{XXX TODO:} It would be nice to show a schedule for the uncontrolled example.\textbf{XXX}
 
-Let's now look at the computation segment $C_2^2$ of the second job of task $\tau_2$.
-This segment is placed back to the ready queue at time $19$ after its self-suspension.
-Without period enforcement, this computation segment should be executed from $19$ to $20$.
-However, the period enforcer algorithm intentionally defers this computation segment to ensure \emph{the periodic demand} of the computation segment $C_2^2$.
-Therefore, this computation segment is \emph{activated} to the ready queue at time $9+T_2=9+11=20$.
-This results in a deadline miss.\footnote{By using the notations in \cite{Raj:suspension1991}, we have $ET_{2,1}^2=9, ET_{2,2}^2 = \max\{9+11, 19\}=20$.
-The superscript $^2$ is for denoting the second computation segment of task $\tau_2$.}
+The period enforcer algorithm determines for each segment (i.e., for each converted single-segment deferrable task) an \emph{eligibility time}.
+If a segment resumes (i.e., a job of a single-segment deferrable task arrives) before its eligibility time, the execution of the segment (i.e., the single-segment job) is delayed until the eligibility time is reached.
 
-Therefore, the above example shows that there exists (at least) a sporadic segmented self-suspending task system that is schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using the period enforcer algorithm is not feasible.
-In this example, it is also clear that we can pessimistically convert the two computation segments of task $\tau_2$ into two deferrable tasks $\tau_2^1$ and $\tau_2^2$.
-The deferrable time of task $\tau_2^1$ is $0$, whereas the deferrable time of task $\tau_2^2$ is at most \emph{$9$}.
-With the above conversion, we can clearly see that the deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable by the rate-monotonic priority assignment since we can release a job of $\tau_1$ together with a job of $\tau_2^2$ (after its worst-case deferrable time).
+Let $ET_{2,j}^1$ and $ET_{2,j}^2$ denote the eligibility times of the first and second segment, respectively, of the $j^{\mathrm{th}}$ job of $\tau_2$.
+Further, let $a^1_{2,j}$ and $a^2_{2,j}$ denote the arrival times\footnote{A segment \emph{arrives} when it becomes available for execution.
+The first segment arrives immediately when the job is released; the segment arrives when the job resumes from its self-suspension.} of the job's first and second segment, respectively.
+Finally, let $\mathit{busy}(\tau_2, t')$ denote the last time that a level-2 busy interval began on or prior to time $t'$.
+According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm then defines the segment eligibility times as
+\begin{align*}
+	ET_{2,j}^1 & = \max\left(ET_{2,j-1}^1 + T_2,\ \mathit{busy}(\tau_2, a^1_{2,j})\right)  \text{ and }
+\\
+	ET_{2,j}^2 & = \max\left(ET_{2,j-1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,j}) \right),
+\end{align*}
+where $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$.
 
-\section{How to Derive the Corresponding Deferrable Task Set}
+Figure \ref{fig:example} shows the resulting schedule assuming a periodic release pattern.
+The first job of task $\tau_2$ (which arrives at time $a^1_{2,1} =  0$) is executed as if there is no period enforcement since the definition $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$ ensures that both segments are immediately eligible.
+Note that the first segment of $\tau_2$'s first job is delayed due to interference from $\tau_1$.
+As a result, the second segment of $\tau_2$'s first job does not resume until time $a^2_{2,1} = 9$.
+Thus, we have
+\begin{align*}
+	ET_{2,1}^1 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 0)\right) = 0  \text{ and }
+\\
+	ET_{2,1}^2 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 9\right) ) = 9.
+\end{align*}
+
+In contrast to the first job, the second job of task $\tau_2$ (which is released at time $11$) is affected by period enforcement.
+The first segment of the second job arrives at time $a^1_{2,2} = 11$, is interfered with for one time unit, and suspends at time $13$.
+The segment second of the second job hence resumes only at time $a^2_{2,2} = 19$.
+Thus, we have
+\begin{align*}
+	ET_{2,2}^1 & = \max\left(0 + 11,\ \mathit{busy}(\tau_2, 11)\right) = 11  \text{ and }
+\\
+	ET_{2,2}^2 & = \max\left(9 + 11,\ \mathit{busy}(\tau_2, 19\right) ) = 20.
+\end{align*}
+According to the rules of the period enforcer algorithm, the processor therefore remains idle at time $19$ because the segment is not eligible to execute until time $ET_{2,2}^2 = 20$.
+However, at time $20$, the third job of $\tau_1$ is released.
+As a result, the second job of $\tau_2$ incurs additional interference and misses its deadline at time $22$.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=0.9\columnwidth]{figures/example}
+\caption{An illustrative example demonstrating a deadline miss at time $22$ under the period enforcer algorithm.
+\label{fig:example}%
+}
+\end{center}
+\end{figure}
+
+This example shows that there exist sporadic segmented self-suspending task sets that   \textbf{(i)} are schedulable under fixed-priority scheduling without any enforcement, but \textbf{(ii)} are not schedulable under the period enforcer algorithm.
+
+Furthermore, the example also demonstrates that the conversion to single-segment deferrable tasks does incur a loss of generality since it introduces pessimism: if we convert the two computation segments of task $\tau_2$ into two deferrable tasks $\tau_2^1$ and $\tau_2^2$, where  task $\tau_2^1$ never defers its execution and task $\tau_2^2$ defers its execution by at most \emph{$9$} time units,  then it is immediately obvious that the resulting single-segment deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable under a rate-monotonic priority assignment since we can time the release of a job of $\tau_1$ to coincide with the arrival of a job of $\tau_2^2$ after it has maximally deferred its execution, which clearly results in a deadline miss.
+
+\clearpage
+
+\section{Deriving a Corresponding Deferrable Task Set}
 \label{sec:convert}
 
-The example in Section \ref{sec:unschedulable} can be easily converted to a corresponding deferrable task set.
-However, it is in general not an easy problem to convert a self-suspending task system into a deferrable task set \emph{precisely}.
-We can demonstrate the difficulty by looking at a special case.
+The example in Section \ref{sec:unschedulable} can be easily converted to a corresponding single-segment deferrable task set.
+However, in general,  \emph{precisely} converting a self-suspending task system into a corresponding single-segment  deferrable task set is not an easy problem.
+We can demonstrate the inherent difficulty by looking at a special case.
+
 Suppose that the system has $k-1$ ordinary sporadic tasks and only one segmented self-suspending task $\tau_k$.
-Converting a computation segment into a deferrable task requires to derive the \emph{worst-case resume time of a computation segment}, denoted as $R_k^j$ for the $j$-th computation segment of task $\tau_k$.
-Suppose that the worst-case response time of the $j$-th computation segment of task $\tau_k$ is $W_k^j$.
+Converting a computation segment into a deferrable task requires to derive the \emph{worst-case resume time of a computation segment}, denoted as $R_k^j$ for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
+Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$.
 It is not difficult to see that $R_k^1=0$ and $R_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k-1$.
-So, we need to derive the worst-case response times of the computation segments of task $\tau_k$.
+We therefore need to derive the worst-case response times of the computation segments of task $\tau_k$.
 
-With the above discussions, for the most simple case defined above, the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems.
-It has been recently shown by Nelissen et al.
-\cite{ecrts15nelissen}  calculating the worst-case response time in the above case is already a very challenging problem.
-Nelissen et al.
-\cite{ecrts15nelissen} identify several misconceptions used in the literature for this problem.
-After excluding those misconceptions, deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very difficult problem.
-Due to \cite{ecrts15nelissen}, the only existing solution to derive the exact $W_k^{j}$ (and hence $R_k^j$) requires exponential time complexity.
-And, also recall the example in \ref{sec:unschedulable}.
-Even if the conversion is done precisely, the deferrable task system can be more pessimistic than the original self-suspending task system with respect to the schedulability.
+Based on these considerations, it appears that, at least for the simple example,  the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems.
+However, it has been recently shown by Nelissen et al.\ \cite{ecrts15nelissen} that calculating the worst-case response time in the above ``simple'' case is already a very challenging problem.
+In particular, Nelissen et al.\ \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.
+
+In the context of the period enforcer, we consequently observe that the only existing solution to derive the \emph{precise} bound $W_k^{j}$ (and hence $R_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity.
+Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.
 
@@
 
-This report explains how to use the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle segmented self-suspending real-time tasks.
-One key assumption in the report in \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
-This creates some mismatches between the original self-suspending task system and the corresponding deferrable task system.
-The report in  \cite{Raj:suspension1991} did not explain how to convert a segmented self-suspending task system to a corresponding deferrable task system.
-We show in this report that this is non-trivial and the main claim of Theorem 5 in \cite{Raj:suspension1991} does not reflect the schedulability of the original self-suspending task system.
+We have revisited the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle segmented self-suspending real-time tasks.
+One key assumption in the original report \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
+This creates some mismatches between the original self-suspending task set and the corresponding deferrable task set, which we have demonstrated with an example that shows that Theorem 5 in \cite{Raj:suspension1991} does not reflect the schedulability of the original self-suspending task system.
 
-Nevertheless, the result in Theorem 5 in \cite{Raj:suspension1991} can be very useful for handling self-suspending task systems if there exist \emph{efficient} schedulability tests of the corresponding deferrable task systems or the period enforcer algorithm.
-However, such tests do not exist yet and remain as open problems.
+Furthermore, the original report \cite{Raj:suspension1991} left open the question of how to convert a segmented self-suspending task system to a corresponding deferrable task system.
+Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a task set transformation is non-trivial in the general case.
+
+Nevertheless, Theorem 5 in \cite{Raj:suspension1991} can be useful for handling self-suspending task systems should there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm.
+However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.

### 2015-10-07 untitled.tex
edited untitled.tex

@@
 
-Nevertheless, Theorem 5 in \cite{Raj:suspension1991} can be useful for handling self-suspending task systems should there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm.
+Nevertheless, Theorem 5 in \cite{Raj:suspension1991} can be useful for handling self-suspending task systems if there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm.
 However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.

### 2015-10-07 untitled.tex
text for original schedule
revise caption
revise caption…

@@
 
-\textbf{XXX TODO:} It would be nice to show a schedule for the uncontrolled example.\textbf{XXX}
-
+Let's release these two tasks periodically, starting from time $0$. The fixed-priority schedule is presented in Figure \ref{fig:example-original}.
 The period enforcer algorithm determines for each segment (i.e., for each converted single-segment deferrable task) an \emph{eligibility time}.
@@
 
-In the context of the period enforcer, we consequently observe that the only existing solution to derive the \emph{precise} bound $W_k^{j}$ (and hence $R_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity.
+In the context of the period enforcer, we consequently observe that the only existing solution to derive the \emph{precise} bound $W_k^{j}$ (and hence $R_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
 Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.

### 2015-10-08 untitled.tex
edited untitled.tex

@@
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.9\columnwidth]{figures/example}
-\caption{An illustrative example demonstrating a deadline miss at time $22$ under the period enforcer algorithm.
-\label{fig:example}%
-}
-\end{center}
-\end{figure}
+This example shows that there exist sporadic segmented self-suspending task sets that   \textbf{(i)} are schedulable under fixed-priority scheduling without any enforcement, but \textbf{(ii)} are not schedulable under the period enforcer algorithm.
 
-This example shows that there exist sporadic segmented self-suspending task sets that   \textbf{(i)} are schedulable under fixed-priority scheduling without any enforcement, but \textbf{(ii)} are not schedulable under the period enforcer algorithm.
+One may consider to enrich the period enforcer with the following power: when the processor becomes idle, the task becomes ready to run right away regardless of the settings of eligibility time.
+Even with this, the above example remains valid by introducing a lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from $3$ to $9$ and $13$ to $20$) and $T_3=D_3=100$.
 

### 2015-10-08 untitled.tex
edited untitled.tex

@@
 One may consider to enrich the period enforcer with the following power: when the processor becomes idle, the task becomes ready to run right away regardless of the settings of eligibility time.
-Even with this, the above example remains valid by introducing a lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from $3$ to $9$ and $13$ to $20$) and $T_3=D_3=100$.
+Even with this, the above example remains valid by introducing one additional lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from $3$ to $9$ and $13$ to $20$) and $T_3=D_3=100$.
+With task $\tau_3$, the processor is always busy from time $0$ to time $23$.
 

### 2015-10-08 conclusion.tex
file splitting

--- 2015-10-07 untitled.tex
@@
-\section{Introduction} 
-This report revists the \emph{period enforcer} algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle self-suspending real-time tasks.
-The report briefly reviews the period enforcer algorithm, explains its underlying assumptions and limitations, and discusses how it may be analyzed to correctly determine the schedulability of self-suspending sporadic real-time tasks subject to period enforcement.
-
-The main contributions of this report are two observations that have not been previously reported in the real-time literature:
-
-\begin{enumerate}
-	\item period enforcement can render self-suspending tasks sets unschedulable that are schedulable otherwise (i.e., period enforcement can induce deadline misses); and
-	\item with current techniques, schedulability analysis of the period enforcer algorithm requires a task set transformation that is subject to exponential time complexity.
-\end{enumerate}
-
-\subsection{Preliminaries}
-
-To date, the real-time literature on self-suspending tasks has focused on two task models: the \emph{dynamic} and the \emph{segmented} (or \emph{multi-segment}) self-suspension model. 
-The dynamic self-suspension sporadic task model characterizes each
-task $\tau_i$ as a $4$-tuple $(C_i,S_i,T_i,D_i)$: 
-$C_i$ denotes the upper bound on the total execution time of any job of $\tau_i$,
-$S_i$ denotes the upper bound on the total self-suspension time of any job of $\tau_i$,
-$T_i$ denotes the minimum inter-arrival time (or period) of $\tau_i$, and $D_i$ is the relative deadline.
-The dynamic self-suspension model does not impose a bound on the maximum number of self-suspensions, nor does it make any assumptions as to where during a job's execution self-suspensions occur.
-
-In contrast, the segmented sporadic task model extends the above $4$-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals.
-These intervals are represented as a tuple
-$(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, which is composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals.
-For the simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment.
-The arguments can be easily extended to handle tasks that start with self-suspensions.
-
-The advantage of the dynamic model is that it is more flexible since it does not impose any assumptions on control flow.
-The advantage of the segmented model is that is allows for more accurate analysis.
-The period enforcer algorithm and its analysis fundamentally applies (only) to the segmented model.
-
-The central notation in Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{deferrable task}, which matches our notion of multi-segment tasks.  Specifically, Rajkumar states that:
-\begin{quote}
-With deferred execution, a task $\tau_i$ can execute its $C_i$ units of execution in discrete amounts $C_i^1, C_i^2$, $\ldots$ with suspension in between $C_i^j$ and $C_i^{j+1}$.
-\cite[Section 3]{Raj:suspension1991}\footnote{The notation has been altered here for the sake of consistency.}
-\end{quote}
-%
-Central to Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{task set transformation} that splits each deferrable task with multiple segments into a corresponding number of single-segment deferrable tasks.
-In the words of Rajkumar~\cite[Section 3]{Raj:suspension1991}:
-
-\begin{quote}
-	 Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
-That is, a task $\tau_i$ executes for $C_i$ units with no suspensions once it begins execution.
-Any task that does suspend after it executes for a while can be considered to be two or more tasks each with its own worst-case execution time.
-The only difference is that if a task $\tau_i$ is split into two tasks $\tau_i'$ followed by $\tau_i''$, then $\tau_i''$ has the same deadlines as $\tau_i{{'}}$.
-\end{quote}
-%
-In other words, the transformation can be understood as splitting each self-suspending task into a matching number of non-self-suspending sporadic tasks subject to \emph{release jitter}, which can be easily analyzed with classic fixed-priority response-time analysis~\cite{ABRTW:93}.
-
-It is well known that uncontrolled deferred execution (i.e, release jitter) can impose a scheduling penalty because of the potential for ``back-to-back'' execution~\cite{ABRTW:93}.
-That is, if a job of a deferrable task that maximally defers its execution is directly followed by a job that executes immediately without deferring its execution, then lower-priority tasks may suffer increased interference.
-
-The purpose of the period enforcer algorithm is to reduce such penalties for lower-priority tasks without detrimentally affecting the schedulability of self-suspending, higher-priority tasks.
-The latter aspect --- no detrimental effects for self-suspending tasks --- is captured concisely by Theorem 5 in the original analysis of the period enforcer algorithm \cite{Raj:suspension1991}.
-\begin{quote}
-{\bf Theorem 5}: A deferrable task that is schedulable under its worst-case conditions is also schedulable under the period enforcer algorithm.  \cite{Raj:suspension1991}
-\end{quote}
-
-\subsection{Questions answered in this report}
-
-Theorem 5 in \cite{Raj:suspension1991} provides a very positive result for the analysis of deferrable task sets subject to period enforcement: if the corresponding transformed task set can be shown to be schedulable under fixed-priority scheduling using \emph{any} applicable analysis, then the period enforcer algorithm also yields a correct schedule.
-
-However, note that Theorem 5 in \cite{Raj:suspension1991} applies only to the \emph{transformed} task set: recall that, in the original analysis~\cite{Raj:suspension1991}, deferrable tasks are assumed to defer their entire execution time either completely or not at all (but not parts of it).
-Therefore, if we would like to use the period enforcer algorithm to handle segmented self-suspending task sets, we first have to answer the following question: ``\emph{Given a set of sporadic segmented self-suspending tasks, what is the corresponding set of (single-segment) deferrable tasks?}'' That is, how do we convert given self-suspension segments into equivalent bounds on release jitter such that we may apply Theorem 5 to conclude that the system remains schedulable despite period enforcement?
-
-In this report, which is motivated by the fact that the original proposal \cite{Raj:suspension1991} does not provide an answer to this central question, 
-we make two pertinent observations.
-\begin{enumerate}
-	\item There exist sporadic segmented self-suspending task sets that are schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using the period enforcer algorithm is not feasible.
-This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, and not in the context of the original segmented self-suspending task set.
-
-	\item Deriving a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem.
-Recent findings by Nelissen et al.
-\cite{ecrts15nelissen} can be applied, but their method takes exponential time.
-\end{enumerate}
-
-\section{Period Enforcement Can Induce Deadline Misses}
-\label{sec:unschedulable}
-
-In this section, we demonstrate with an example that there exist sporadic  segmented self-suspending task sets that both \textbf{(i)} are schedulable \emph{without} period enforcement and \textbf{(ii)} are not schedulable with period enforcement.
-
-Consider a task system with $2$ tasks.
-Let $\tau_1$ denote a sporadic task without self-suspensions and parameters $C_1 = 2$ and $T_1=D_1=10$, and let $\tau_2$ denote a self-suspending task consisting of two segments with parameters  $C_{2,1} = 1$,  $S_{2,1} = 6$, $C_{2,2}=1$, and $ T_2=D_2=11$.
-Suppose that we use the rate-monotonic priority assignment, i.e., $\tau_1$ has higher priority than $\tau_2$.
-This task set is schedulable without any enforcement since at most one computation segment of a job of $\tau_2$ can be delayed by $\tau_1$:
-\begin{itemize}
-	\item if the first segment of a job of $\tau_2$ is interfered with by $\tau_1$, then the second segment resumes at most after $9$ time units after the release of the job and the response time of task $\tau_2$ is hence $10$; otherwise,
-	\item  if the first segment of a job of $\tau_2$ is not interfered with by $\tau_1$, then the second segment resumes at most $7$ time units after the release of the job and hence the  response time of task $\tau_2$ is at most $10$ even if the second segment is interfered with by $\tau_1$.
-\end{itemize}
-
-Let's release these two tasks periodically, starting from time $0$. The fixed-priority schedule is presented in Figure \ref{fig:example-original}.
-The period enforcer algorithm determines for each segment (i.e., for each converted single-segment deferrable task) an \emph{eligibility time}.
-If a segment resumes (i.e., a job of a single-segment deferrable task arrives) before its eligibility time, the execution of the segment (i.e., the single-segment job) is delayed until the eligibility time is reached.
-
-Let $ET_{2,j}^1$ and $ET_{2,j}^2$ denote the eligibility times of the first and second segment, respectively, of the $j^{\mathrm{th}}$ job of $\tau_2$.
-Further, let $a^1_{2,j}$ and $a^2_{2,j}$ denote the arrival times\footnote{A segment \emph{arrives} when it becomes available for execution.
-The first segment arrives immediately when the job is released; the segment arrives when the job resumes from its self-suspension.} of the job's first and second segment, respectively.
-Finally, let $\mathit{busy}(\tau_2, t')$ denote the last time that a level-2 busy interval began on or prior to time $t'$.
-According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm then defines the segment eligibility times as
-\begin{align*}
-	ET_{2,j}^1 & = \max\left(ET_{2,j-1}^1 + T_2,\ \mathit{busy}(\tau_2, a^1_{2,j})\right)  \text{ and }
-\\
-	ET_{2,j}^2 & = \max\left(ET_{2,j-1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,j}) \right),
-\end{align*}
-where $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$.
-
-Figure \ref{fig:example} shows the resulting schedule assuming a periodic release pattern.
-The first job of task $\tau_2$ (which arrives at time $a^1_{2,1} =  0$) is executed as if there is no period enforcement since the definition $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$ ensures that both segments are immediately eligible.
-Note that the first segment of $\tau_2$'s first job is delayed due to interference from $\tau_1$.
-As a result, the second segment of $\tau_2$'s first job does not resume until time $a^2_{2,1} = 9$.
-Thus, we have
-\begin{align*}
-	ET_{2,1}^1 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 0)\right) = 0  \text{ and }
-\\
-	ET_{2,1}^2 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 9\right) ) = 9.
-\end{align*}
-
-In contrast to the first job, the second job of task $\tau_2$ (which is released at time $11$) is affected by period enforcement.
-The first segment of the second job arrives at time $a^1_{2,2} = 11$, is interfered with for one time unit, and suspends at time $13$.
-The segment second of the second job hence resumes only at time $a^2_{2,2} = 19$.
-Thus, we have
-\begin{align*}
-	ET_{2,2}^1 & = \max\left(0 + 11,\ \mathit{busy}(\tau_2, 11)\right) = 11  \text{ and }
-\\
-	ET_{2,2}^2 & = \max\left(9 + 11,\ \mathit{busy}(\tau_2, 19\right) ) = 20.
-\end{align*}
-According to the rules of the period enforcer algorithm, the processor therefore remains idle at time $19$ because the segment is not eligible to execute until time $ET_{2,2}^2 = 20$.
-However, at time $20$, the third job of $\tau_1$ is released.
-As a result, the second job of $\tau_2$ incurs additional interference and misses its deadline at time $22$.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.9\columnwidth]{figures/example}
-\caption{An illustrative example demonstrating a deadline miss at time $22$ under the period enforcer algorithm.
-\label{fig:example}%
-}
-\end{center}
-\end{figure}
-
-This example shows that there exist sporadic segmented self-suspending task sets that   \textbf{(i)} are schedulable under fixed-priority scheduling without any enforcement, but \textbf{(ii)} are not schedulable under the period enforcer algorithm.
-
-Furthermore, the example also demonstrates that the conversion to single-segment deferrable tasks does incur a loss of generality since it introduces pessimism: if we convert the two computation segments of task $\tau_2$ into two deferrable tasks $\tau_2^1$ and $\tau_2^2$, where  task $\tau_2^1$ never defers its execution and task $\tau_2^2$ defers its execution by at most \emph{$9$} time units,  then it is immediately obvious that the resulting single-segment deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable under a rate-monotonic priority assignment since we can time the release of a job of $\tau_1$ to coincide with the arrival of a job of $\tau_2^2$ after it has maximally deferred its execution, which clearly results in a deadline miss.
-
-\clearpage
-
-\section{Deriving a Corresponding Deferrable Task Set}
-\label{sec:convert}
-
-The example in Section \ref{sec:unschedulable} can be easily converted to a corresponding single-segment deferrable task set.
-However, in general,  \emph{precisely} converting a self-suspending task system into a corresponding single-segment  deferrable task set is not an easy problem.
-We can demonstrate the inherent difficulty by looking at a special case.
-
-Suppose that the system has $k-1$ ordinary sporadic tasks and only one segmented self-suspending task $\tau_k$.
-Converting a computation segment into a deferrable task requires to derive the \emph{worst-case resume time of a computation segment}, denoted as $R_k^j$ for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
-Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$.
-It is not difficult to see that $R_k^1=0$ and $R_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k-1$.
-We therefore need to derive the worst-case response times of the computation segments of task $\tau_k$.
-
-Based on these considerations, it appears that, at least for the simple example,  the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems.
-However, it has been recently shown by Nelissen et al.\ \cite{ecrts15nelissen} that calculating the worst-case response time in the above ``simple'' case is already a very challenging problem.
-In particular, Nelissen et al.\ \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.
-
-In the context of the period enforcer, we consequently observe that the only existing solution to derive the \emph{precise} bound $W_k^{j}$ (and hence $R_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
-Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.
-
 \section{Concluding Remarks}

### 2015-10-08 transformation.tex
file splitting

--- 2015-10-07 untitled.tex
@@
-\section{Introduction} 
-This report revists the \emph{period enforcer} algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle self-suspending real-time tasks.
-The report briefly reviews the period enforcer algorithm, explains its underlying assumptions and limitations, and discusses how it may be analyzed to correctly determine the schedulability of self-suspending sporadic real-time tasks subject to period enforcement.
-
-The main contributions of this report are two observations that have not been previously reported in the real-time literature:
-
-\begin{enumerate}
-	\item period enforcement can render self-suspending tasks sets unschedulable that are schedulable otherwise (i.e., period enforcement can induce deadline misses); and
-	\item with current techniques, schedulability analysis of the period enforcer algorithm requires a task set transformation that is subject to exponential time complexity.
-\end{enumerate}
-
-\subsection{Preliminaries}
-
-To date, the real-time literature on self-suspending tasks has focused on two task models: the \emph{dynamic} and the \emph{segmented} (or \emph{multi-segment}) self-suspension model. 
-The dynamic self-suspension sporadic task model characterizes each
-task $\tau_i$ as a $4$-tuple $(C_i,S_i,T_i,D_i)$: 
-$C_i$ denotes the upper bound on the total execution time of any job of $\tau_i$,
-$S_i$ denotes the upper bound on the total self-suspension time of any job of $\tau_i$,
-$T_i$ denotes the minimum inter-arrival time (or period) of $\tau_i$, and $D_i$ is the relative deadline.
-The dynamic self-suspension model does not impose a bound on the maximum number of self-suspensions, nor does it make any assumptions as to where during a job's execution self-suspensions occur.
-
-In contrast, the segmented sporadic task model extends the above $4$-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals.
-These intervals are represented as a tuple
-$(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, which is composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals.
-For the simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment.
-The arguments can be easily extended to handle tasks that start with self-suspensions.
-
-The advantage of the dynamic model is that it is more flexible since it does not impose any assumptions on control flow.
-The advantage of the segmented model is that is allows for more accurate analysis.
-The period enforcer algorithm and its analysis fundamentally applies (only) to the segmented model.
-
-The central notation in Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{deferrable task}, which matches our notion of multi-segment tasks.  Specifically, Rajkumar states that:
-\begin{quote}
-With deferred execution, a task $\tau_i$ can execute its $C_i$ units of execution in discrete amounts $C_i^1, C_i^2$, $\ldots$ with suspension in between $C_i^j$ and $C_i^{j+1}$.
-\cite[Section 3]{Raj:suspension1991}\footnote{The notation has been altered here for the sake of consistency.}
-\end{quote}
-%
-Central to Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{task set transformation} that splits each deferrable task with multiple segments into a corresponding number of single-segment deferrable tasks.
-In the words of Rajkumar~\cite[Section 3]{Raj:suspension1991}:
-
-\begin{quote}
-	 Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
-That is, a task $\tau_i$ executes for $C_i$ units with no suspensions once it begins execution.
-Any task that does suspend after it executes for a while can be considered to be two or more tasks each with its own worst-case execution time.
-The only difference is that if a task $\tau_i$ is split into two tasks $\tau_i'$ followed by $\tau_i''$, then $\tau_i''$ has the same deadlines as $\tau_i{{'}}$.
-\end{quote}
-%
-In other words, the transformation can be understood as splitting each self-suspending task into a matching number of non-self-suspending sporadic tasks subject to \emph{release jitter}, which can be easily analyzed with classic fixed-priority response-time analysis~\cite{ABRTW:93}.
-
-It is well known that uncontrolled deferred execution (i.e, release jitter) can impose a scheduling penalty because of the potential for ``back-to-back'' execution~\cite{ABRTW:93}.
-That is, if a job of a deferrable task that maximally defers its execution is directly followed by a job that executes immediately without deferring its execution, then lower-priority tasks may suffer increased interference.
-
-The purpose of the period enforcer algorithm is to reduce such penalties for lower-priority tasks without detrimentally affecting the schedulability of self-suspending, higher-priority tasks.
-The latter aspect --- no detrimental effects for self-suspending tasks --- is captured concisely by Theorem 5 in the original analysis of the period enforcer algorithm \cite{Raj:suspension1991}.
-\begin{quote}
-{\bf Theorem 5}: A deferrable task that is schedulable under its worst-case conditions is also schedulable under the period enforcer algorithm.  \cite{Raj:suspension1991}
-\end{quote}
-
-\subsection{Questions answered in this report}
-
-Theorem 5 in \cite{Raj:suspension1991} provides a very positive result for the analysis of deferrable task sets subject to period enforcement: if the corresponding transformed task set can be shown to be schedulable under fixed-priority scheduling using \emph{any} applicable analysis, then the period enforcer algorithm also yields a correct schedule.
-
-However, note that Theorem 5 in \cite{Raj:suspension1991} applies only to the \emph{transformed} task set: recall that, in the original analysis~\cite{Raj:suspension1991}, deferrable tasks are assumed to defer their entire execution time either completely or not at all (but not parts of it).
-Therefore, if we would like to use the period enforcer algorithm to handle segmented self-suspending task sets, we first have to answer the following question: ``\emph{Given a set of sporadic segmented self-suspending tasks, what is the corresponding set of (single-segment) deferrable tasks?}'' That is, how do we convert given self-suspension segments into equivalent bounds on release jitter such that we may apply Theorem 5 to conclude that the system remains schedulable despite period enforcement?
-
-In this report, which is motivated by the fact that the original proposal \cite{Raj:suspension1991} does not provide an answer to this central question, 
-we make two pertinent observations.
-\begin{enumerate}
-	\item There exist sporadic segmented self-suspending task sets that are schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using the period enforcer algorithm is not feasible.
-This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, and not in the context of the original segmented self-suspending task set.
-
-	\item Deriving a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem.
-Recent findings by Nelissen et al.
-\cite{ecrts15nelissen} can be applied, but their method takes exponential time.
-\end{enumerate}
-
-\section{Period Enforcement Can Induce Deadline Misses}
-\label{sec:unschedulable}
-
-In this section, we demonstrate with an example that there exist sporadic  segmented self-suspending task sets that both \textbf{(i)} are schedulable \emph{without} period enforcement and \textbf{(ii)} are not schedulable with period enforcement.
-
-Consider a task system with $2$ tasks.
-Let $\tau_1$ denote a sporadic task without self-suspensions and parameters $C_1 = 2$ and $T_1=D_1=10$, and let $\tau_2$ denote a self-suspending task consisting of two segments with parameters  $C_{2,1} = 1$,  $S_{2,1} = 6$, $C_{2,2}=1$, and $ T_2=D_2=11$.
-Suppose that we use the rate-monotonic priority assignment, i.e., $\tau_1$ has higher priority than $\tau_2$.
-This task set is schedulable without any enforcement since at most one computation segment of a job of $\tau_2$ can be delayed by $\tau_1$:
-\begin{itemize}
-	\item if the first segment of a job of $\tau_2$ is interfered with by $\tau_1$, then the second segment resumes at most after $9$ time units after the release of the job and the response time of task $\tau_2$ is hence $10$; otherwise,
-	\item  if the first segment of a job of $\tau_2$ is not interfered with by $\tau_1$, then the second segment resumes at most $7$ time units after the release of the job and hence the  response time of task $\tau_2$ is at most $10$ even if the second segment is interfered with by $\tau_1$.
-\end{itemize}
-
-Let's release these two tasks periodically, starting from time $0$. The fixed-priority schedule is presented in Figure \ref{fig:example-original}.
-The period enforcer algorithm determines for each segment (i.e., for each converted single-segment deferrable task) an \emph{eligibility time}.
-If a segment resumes (i.e., a job of a single-segment deferrable task arrives) before its eligibility time, the execution of the segment (i.e., the single-segment job) is delayed until the eligibility time is reached.
-
-Let $ET_{2,j}^1$ and $ET_{2,j}^2$ denote the eligibility times of the first and second segment, respectively, of the $j^{\mathrm{th}}$ job of $\tau_2$.
-Further, let $a^1_{2,j}$ and $a^2_{2,j}$ denote the arrival times\footnote{A segment \emph{arrives} when it becomes available for execution.
-The first segment arrives immediately when the job is released; the segment arrives when the job resumes from its self-suspension.} of the job's first and second segment, respectively.
-Finally, let $\mathit{busy}(\tau_2, t')$ denote the last time that a level-2 busy interval began on or prior to time $t'$.
-According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm then defines the segment eligibility times as
-\begin{align*}
-	ET_{2,j}^1 & = \max\left(ET_{2,j-1}^1 + T_2,\ \mathit{busy}(\tau_2, a^1_{2,j})\right)  \text{ and }
-\\
-	ET_{2,j}^2 & = \max\left(ET_{2,j-1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,j}) \right),
-\end{align*}
-where $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$.
-
-Figure \ref{fig:example} shows the resulting schedule assuming a periodic release pattern.
-The first job of task $\tau_2$ (which arrives at time $a^1_{2,1} =  0$) is executed as if there is no period enforcement since the definition $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$ ensures that both segments are immediately eligible.
-Note that the first segment of $\tau_2$'s first job is delayed due to interference from $\tau_1$.
-As a result, the second segment of $\tau_2$'s first job does not resume until time $a^2_{2,1} = 9$.
-Thus, we have
-\begin{align*}
-	ET_{2,1}^1 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 0)\right) = 0  \text{ and }
-\\
-	ET_{2,1}^2 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 9\right) ) = 9.
-\end{align*}
-
-In contrast to the first job, the second job of task $\tau_2$ (which is released at time $11$) is affected by period enforcement.
-The first segment of the second job arrives at time $a^1_{2,2} = 11$, is interfered with for one time unit, and suspends at time $13$.
-The segment second of the second job hence resumes only at time $a^2_{2,2} = 19$.
-Thus, we have
-\begin{align*}
-	ET_{2,2}^1 & = \max\left(0 + 11,\ \mathit{busy}(\tau_2, 11)\right) = 11  \text{ and }
-\\
-	ET_{2,2}^2 & = \max\left(9 + 11,\ \mathit{busy}(\tau_2, 19\right) ) = 20.
-\end{align*}
-According to the rules of the period enforcer algorithm, the processor therefore remains idle at time $19$ because the segment is not eligible to execute until time $ET_{2,2}^2 = 20$.
-However, at time $20$, the third job of $\tau_1$ is released.
-As a result, the second job of $\tau_2$ incurs additional interference and misses its deadline at time $22$.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.9\columnwidth]{figures/example}
-\caption{An illustrative example demonstrating a deadline miss at time $22$ under the period enforcer algorithm.
-\label{fig:example}%
-}
-\end{center}
-\end{figure}
-
-This example shows that there exist sporadic segmented self-suspending task sets that   \textbf{(i)} are schedulable under fixed-priority scheduling without any enforcement, but \textbf{(ii)} are not schedulable under the period enforcer algorithm.
-
-Furthermore, the example also demonstrates that the conversion to single-segment deferrable tasks does incur a loss of generality since it introduces pessimism: if we convert the two computation segments of task $\tau_2$ into two deferrable tasks $\tau_2^1$ and $\tau_2^2$, where  task $\tau_2^1$ never defers its execution and task $\tau_2^2$ defers its execution by at most \emph{$9$} time units,  then it is immediately obvious that the resulting single-segment deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable under a rate-monotonic priority assignment since we can time the release of a job of $\tau_1$ to coincide with the arrival of a job of $\tau_2^2$ after it has maximally deferred its execution, which clearly results in a deadline miss.
-
-\clearpage
-
 \section{Deriving a Corresponding Deferrable Task Set}
@@
 Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.
-
-\section{Concluding Remarks}
-
-We have revisited the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle segmented self-suspending real-time tasks.
-One key assumption in the original report \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
-This creates some mismatches between the original self-suspending task set and the corresponding deferrable task set, which we have demonstrated with an example that shows that Theorem 5 in \cite{Raj:suspension1991} does not reflect the schedulability of the original self-suspending task system.
-
-Furthermore, the original report \cite{Raj:suspension1991} left open the question of how to convert a segmented self-suspending task system to a corresponding deferrable task system.
-Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a task set transformation is non-trivial in the general case.
-
-Nevertheless, Theorem 5 in \cite{Raj:suspension1991} can be useful for handling self-suspending task systems if there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm.
-However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.

### 2015-10-08 untitled.tex
file splitting

@@
 
-\clearpage
-
-\section{Deriving a Corresponding Deferrable Task Set}
-\label{sec:convert}
-
-The example in Section \ref{sec:unschedulable} can be easily converted to a corresponding single-segment deferrable task set.
-However, in general,  \emph{precisely} converting a self-suspending task system into a corresponding single-segment  deferrable task set is not an easy problem.
-We can demonstrate the inherent difficulty by looking at a special case.
-
-Suppose that the system has $k-1$ ordinary sporadic tasks and only one segmented self-suspending task $\tau_k$.
-Converting a computation segment into a deferrable task requires to derive the \emph{worst-case resume time of a computation segment}, denoted as $R_k^j$ for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
-Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$.
-It is not difficult to see that $R_k^1=0$ and $R_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k-1$.
-We therefore need to derive the worst-case response times of the computation segments of task $\tau_k$.
-
-Based on these considerations, it appears that, at least for the simple example,  the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems.
-However, it has been recently shown by Nelissen et al.\ \cite{ecrts15nelissen} that calculating the worst-case response time in the above ``simple'' case is already a very challenging problem.
-In particular, Nelissen et al.\ \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.
-
-In the context of the period enforcer, we consequently observe that the only existing solution to derive the \emph{precise} bound $W_k^{j}$ (and hence $R_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
-Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.
-
-\section{Concluding Remarks}
-
-We have revisited the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle segmented self-suspending real-time tasks.
-One key assumption in the original report \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
-This creates some mismatches between the original self-suspending task set and the corresponding deferrable task set, which we have demonstrated with an example that shows that Theorem 5 in \cite{Raj:suspension1991} does not reflect the schedulability of the original self-suspending task system.
-
-Furthermore, the original report \cite{Raj:suspension1991} left open the question of how to convert a segmented self-suspending task system to a corresponding deferrable task system.
-Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a task set transformation is non-trivial in the general case.
-
-Nevertheless, Theorem 5 in \cite{Raj:suspension1991} can be useful for handling self-suspending task systems if there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm.
-However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.
+%\clearpage

### 2015-10-08 LITES/LITES-Paper.tex
LITES format
add/remove stuffs for LITES.

\input{../preamble}

\input{../header.tex}

\title{A Note on the Period Enforcer Algorithm  for Self-Suspending Tasks}

\author[1]{Jian-Jia Chen}
\author[2]{Bj\"orn Brandenburg}
\affil[1]{TU Dortmund University, Germany\\
  \texttt{jian-jia.chen@cs.uni-dortmund.de}}
\affil[2]{Max Planck Institute for Software Systems (MPI-SWS)\\
  \texttt{bbb@mpi-sws.org}}
\authorrunning{Chen and Brandenburg} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

\Copyright{Chen and Brandenburg}%mandatory. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

%\subjclass{MANDATORY:  Please refer to \url{www.acm.org/about/class/2012}}% mandatory: Please choose 2012 ACM classifications from http://www.acm.org/about/class/2012 . E.g., cite as "Integrated circuits". 
%\keywords{MANDATORY: Please provide 1--5 keywords as a comma-separated list}% mandatory: Please provide 1-5 keywords
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%EditorialOffice macros (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Volume{XXX}
\Issue{YYY}
\DateSubmission{Date of submission}
\DateAcceptance{Date of acceptance}
\DatePublished{Date of publishing}
\SectionAreaEditor{LITES section area editor}
\DOI{10.4230/LITES.xxx.yyy.p}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle
\input{../abstract.tex}
\input{../untitled.tex}
\begin{figure}[t]
  \centering
  \input{../figures/example-original/example-body.tex}
  \caption{An illustrative example of the original self-suspending task schedule.}
  \label{fig:example-original}
  \end{figure}

\begin{figure}[t]
  \centering
  \input{../figures/example/example-body.tex}
  \caption{An illustrative example demonstrating a deadline miss at time 22
    under the period enforcer algorithm.}
\label{fig:example}  
\end{figure}
%\input{../figures/example-original/example.png}
%\input{../figures/example/example.png}
\input{../transformation.tex}
\input{../conclusion.tex}

\bibliography{../bibliography/biblio}{}

\end{document}

### 2015-10-08 figures/example/example-body.tex
LITES format
add/remove stuffs for LITES.

		\renewcommand{\arraystretch}{1.3}

		\def\ux{0.4cm}\def\uy{0.75cm} 
		\begin{tikzpicture}[y=\uy, font=\sffamily,thick]  
		\tikzset{
			task/.style={fill=#1,  rectangle, text height=.3cm},
			task1a/.style={task=green!30},
			task1b/.style={task=green},
			task2a/.style={task=orange!30},
			task2b/.style={task=orange, minimum width=1mm},
			task3a/.style={task=pink, minimum width=1mm},
			task3b/.style={task=pink!80, minimum width=1mm},
			task4a/.style={task=cyan, minimum width=1mm},
			task4b/.style={task=cyan!50},
			task5/.style={task=blue},
			task6/.style={task=purple},
			task7/.style={minimum height=\uy,draw},
			task8/.style={minimum height=\uy,draw,thick},
			task9/.style={task=gray,minimum height=0.7cm,draw},
		}
		\tikzstyle{jobs}=[ fill=black!50];

		\begin{scope}[shift={(0,0)}]

		\draw[<-](0,2) -- (0,0);
		\draw[<->](5,2) -- (5,0);
		\draw[<->](10,2) -- (10,0);

		\draw[->] (0,0) node[anchor=east] {$\tau_1$}-- coordinate (xaxis) (12.5,0);

		\node[task7,minimum width=2.5*\ux, 
                anchor=south west] at (0,0){};
		\node[task7, minimum width=2.5*\ux,
		anchor=south west]at (5, 0){};
		\node[task7, minimum width=2.5*\ux,
		anchor=south west]at (10, 0){};
		\end{scope}

		\begin{scope}[shift={(0,-2.5)}]
		%%timeline

		\draw[<-](0,2) -- (0,0);
		\draw[<->](5.5,2) -- (5.5,0);
		\draw[<->](11,2) -- (11,0);

		\node[task7, minimum width=1.25*\ux,
		anchor=south west] at (1, 0){\tiny $C_2^1$};
		\draw[(-), thin] (1.5, 0.3) -- (3, 0.3)
                node[anchor=south]{\scriptsize suspension} -- (4.5, 0.3);
		\node[task7, minimum width=1.25*\ux,
		anchor=south west] at (4.5, 0){\tiny $C_2^2$};

		\node[task7, minimum width=1.25*\ux,
		anchor=south west] at (6, 0){\tiny $C_2^1$};
		\draw[(-), thin] (6.5, 0.3) -- (8, 0.3)
                node[anchor=south]{\scriptsize suspension} -- (9.5, 0.3);

		\draw[->] (0,0)node[anchor=east] {$\tau_2$} -- coordinate (xaxis) (12.5,0);

		\node[task7, minimum width=1.25*\ux,
		anchor=south west] at (11, 0){\tiny $C_2^2$};
		\node[task7, minimum width=1.25*\ux,
		anchor=south west] at (11.5, 0){\tiny $C_2^1$};

               \foreach \x in {0,...,24}{
                        \draw[-](.5*\x,0.1) -- (.5*\x,-0.1)
                        node[below] {\small $\x$};

                }	
                \draw (11,1.1) node[anchor=south] {\scriptsize\textcolor{red}{deadline miss}};
		\end{scope}

		\end{tikzpicture}

### 2015-10-08 untitled.tex
Some minor polishing and shuffling around of the figures

@@
 \section{Introduction} 
-This report revists the \emph{period enforcer} algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle self-suspending real-time tasks.
+This report revisits the \emph{period enforcer} algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle self-suspending real-time tasks.
 The report briefly reviews the period enforcer algorithm, explains its underlying assumptions and limitations, and discusses how it may be analyzed to correctly determine the schedulability of self-suspending sporadic real-time tasks subject to period enforcement.
@@
 \end{itemize}
+Figure \ref{fig:example-original} depicts an example schedule of the task set assuming periodic job arrivals.
 
-Let's release these two tasks periodically, starting from time $0$. The fixed-priority schedule is presented in Figure \ref{fig:example-original}.
+\begin{figure}[t]
+  \centering
+  \input{../figures/example-original/example-body.tex}
+  \caption{An illustrative example of the original self-suspending task set (without period enforcement) assuming periodic job arrivals.}
+  \label{fig:example-original}
+  \end{figure}
+
+Next, let us consider the same task set under control of the period enforcer algorithm.
 The period enforcer algorithm determines for each segment (i.e., for each converted single-segment deferrable task) an \emph{eligibility time}.
@@
 
+\begin{figure}[t]
+  \centering
+  \input{../figures/example/example-body.tex}
+  \caption{An illustrative example demonstrating a deadline miss at time 22
+    under the period enforcer algorithm.}
+\label{fig:example}  
+\end{figure}
+
 This example shows that there exist sporadic segmented self-suspending task sets that   \textbf{(i)} are schedulable under fixed-priority scheduling without any enforcement, but \textbf{(ii)} are not schedulable under the period enforcer algorithm.

### 2015-10-08 untitled.tex
light editing of the note on "idle reset"

@@
 
-One may consider to enrich the period enforcer with the following power: when the processor becomes idle, the task becomes ready to run right away regardless of the settings of eligibility time.
-Even with this, the above example remains valid by introducing one additional lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from $3$ to $9$ and $13$ to $20$) and $T_3=D_3=100$.
-With task $\tau_3$, the processor is always busy from time $0$ to time $23$.
+One may consider to enrich the period enforcer with the following power: when the processor becomes idle, a task immediately becomes eligible to execute regardless of its eligibility time.
+However, even with this exension, the above example remains valid by introducing one additional lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from time $3$ to time $9$ and time $13$ to time $20$) and $T_3=D_3=100$.
+With task $\tau_3$, the processor is always busy from time $0$ to time $23$ and consequently $\tau_2$ still misses its deadline at time $22$.
 

### 2015-10-08 conclusion.tex
proofreading fixes and tweaks

@@
 We have revisited the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle segmented self-suspending real-time tasks.
-One key assumption in the original report \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
+One key assumption in the original proposal \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
 This creates some mismatches between the original self-suspending task set and the corresponding deferrable task set, which we have demonstrated with an example that shows that Theorem 5 in \cite{Raj:suspension1991} does not reflect the schedulability of the original self-suspending task system.
 
-Furthermore, the original report \cite{Raj:suspension1991} left open the question of how to convert a segmented self-suspending task system to a corresponding deferrable task system.
+Furthermore, the original proposal \cite{Raj:suspension1991} left open the question of how to convert a segmented self-suspending task system to a corresponding deferrable task system.
 Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a task set transformation is non-trivial in the general case.

### 2015-10-08 untitled.tex
proofreading fixes and tweaks

@@
 This report revisits the \emph{period enforcer} algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle self-suspending real-time tasks.
-The report briefly reviews the period enforcer algorithm, explains its underlying assumptions and limitations, and discusses how it may be analyzed to correctly determine the schedulability of self-suspending sporadic real-time tasks subject to period enforcement.
+We briefly review the period enforcer algorithm, explain its underlying assumptions and limitations, and discuss how it may be analyzed to correctly determine the schedulability of self-suspending sporadic real-time tasks subject to period enforcement.
 
-The main contributions of this report are two observations that have not been previously reported in the real-time literature:
+The main contributions of this paper are two observations that have not been previously reported in the real-time literature:
 
@@
 
-To date, the real-time literature on self-suspending tasks has focused on two task models: the \emph{dynamic} and the \emph{segmented} (or \emph{multi-segment}) self-suspension model. 
+To date, the real-time literature on self-suspensions has focused on two task models: the \emph{dynamic} and the \emph{segmented} (or \emph{multi-segment}) self-suspension model. 
 The dynamic self-suspension sporadic task model characterizes each
 task $\tau_i$ as a $4$-tuple $(C_i,S_i,T_i,D_i)$: 
-$C_i$ denotes the upper bound on the total execution time of any job of $\tau_i$,
-$S_i$ denotes the upper bound on the total self-suspension time of any job of $\tau_i$,
+$C_i$ denotes an upper bound on the total execution time of any job of $\tau_i$,
+$S_i$ denotes an upper bound on the total self-suspension time of any job of $\tau_i$,
 $T_i$ denotes the minimum inter-arrival time (or period) of $\tau_i$, and $D_i$ is the relative deadline.
@@
 For the simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment.
-The arguments can be easily extended to handle tasks that start with self-suspensions.
+The arguments can be easily extended to handle tasks that start with a self-suspension.
 
@@
 
-\subsection{Questions answered in this report}
+\subsection{Questions answered in this paper}
 
@@
 
-In this report, which is motivated by the fact that the original proposal \cite{Raj:suspension1991} does not provide an answer to this central question, 
+In this paper, which is motivated by the fact that the original proposal \cite{Raj:suspension1991} does not provide an answer to this central question, 
 we make two pertinent observations.
@@
 
-Consider a task system with $2$ tasks.
+To this end, consider a task system consisting of $2$ tasks.
 Let $\tau_1$ denote a sporadic task without self-suspensions and parameters $C_1 = 2$ and $T_1=D_1=10$, and let $\tau_2$ denote a self-suspending task consisting of two segments with parameters  $C_{2,1} = 1$,  $S_{2,1} = 6$, $C_{2,2}=1$, and $ T_2=D_2=11$.

### 2015-10-13 locking.tex
add discussion of first locking counterexample

\section{Incompatibility with Suspension-Based Locking Protocols}
\label{sec:locking}

A common source of self-suspensions in multiprocessor real-time systems are \emph{binary semaphores}, i.e., suspension-based locks used to realize mutually exclusive access to shared resources. When a task tries to use a resource that has already been locked, it self-suspends until the resource becomes available. Such self-suspensions due to lock contention, just like any other self-suspension, result in deferred execution and thus can detrimentally affect the interference of lower-priority tasks. It may thus seem natural to apply the period enforcer algorithm to control the negative effects of locking-induced self-suspensions.\footnote{The combination of suspension-based locks and period enforcement has indeed been assumed in prior work~\cite{Raj:91} and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.} However, as we demonstrate with two examples, it is actually not safe to combine the period enforcer algorithm with suspension-based locks.

\subsection{Combining Period Enforcement and Suspension-Based Locks}

Whenever a task attempts to lock a shared resource, it may potentially block if the resource has already been  locked by another task. In the context of the multi-segmented self-suspending task model, each lock operations hence marks the beginning of a new segment.

The period enforcer algorithm may therefore be applied to determine the eligibility time of each such segment (which, again, starts with a critical section). There is, however, one complication: when does the task actually \emph{acquire} the lock? That is, if a task's execution is postponed due to the rules of the period enforcement algorithm, at which point is the lock request processed, with the consequence that the resource become unavailable to other tasks?

There are two possible interpretations.
\begin{enumerate}
	\item When a task requires a shared resource, which implies the beginning of a new segment, its lock request is processed only when its new segment becomes eligible for execution (as determined by the period enforcer algorithm). As a consequence, a task may find a required shared resource unavailable when its new segment becomes eligible for execution even though it was available when the prior segment finished.
	\item Alternatively, when a task requires a shared resource, its request may be processed \emph{immediately}. As a consequence, a shared resource may be locked by a task that is currently ineligible to execute.
\end{enumerate}
We believe that the first interpretation is the more natural one. However, for the sake of completeness, we show that either interpretation leads to potentially unbounded worst-case response times.

\subsection{Alternative 1: Locking Takes Effect at Segment Eligibility Time}
Consider a simple task set consisting of two tasks on two processors that share one resource. Task $\tau_1$, on processor~1, has a total execution cost of $C_1 = 4$ and a period and deadline of $T_1 = D_i = 8$. After one time unit of execution, jobs of $\tau_1$ require the shared resource for two time units. $\tau_1$ thus consists of two segments with cost $C_{1,1} = 1$ and $C_{1,2} = 3$. Task $\tau_2$, on processor~2, has the same overall WCET ($C_2 = 4$), a slightly shorter period ($T_2 = D_2 = 7$), and requires the shared resource for one time unit after \emph{two} time units of execution ($C_{2,1} = 2$ and $C_{2,2} = 2$).

A schedule of the two tasks assuming period job arrivals is depicted in Figure~\ref{XXX}. We focus on the eligibility times $ET_{2,1}^2,ET_{2,2}^2,ET_{2,3}^2,\ldots$ of the second segment of $\tau_2$.

Since $\tau_2$'s first job requests the shared resource only after two time units of execution, it is blocked by $\tau_1$'s critical section, which commenced at time $1$. At time $3$, $\tau_1$ releases the shared resource and $\tau_2$ consequently resumes (i.e., $a^2_{2,1} = 3$). According to the period enforcer rules~\cite{Raj:suspension1991}, the second segment is immediately eligible because
\begin{align*}
	ET_{2,1}^2 & = \max\left(ET_{2,0}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,1})\right) =\max(-T_2 + T_2,\ 3) = 3.
\end{align*}
(Recall that $ET_{2,0}^2 = -T_2$, and interpret $\mathit{busy}(\tau_2, a^2_{2,1})$ with respect to $\tau_2$'s processor.)

At time $7$, the second job of $\tau_2$ is released. Its first segment ends at time $9$. However, its second segment is not eligible to be scheduled before time $10$ since $ET_{2,2}^2 \geq ET_{2,1}^2 + T_2 = 3 + 7 = 10$. At time $9$, the second job of $\tau_1$, released at time $8$, can thus lock the shared resource without contention. Consequently, when $\tau_2$'s request for the shared resource takes effect at time $10$, the resource is no longer available and $\tau_2$ must wait until time $a^2_{2,2} = 11$ before it can proceed in its execution. We thus have
\begin{align*}
	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,2})\right) =\max(10,\ 11) = 11.
\end{align*}

The third job of $\tau_2$ is released at time $14$. Its first segment ends at time $16$, but since $ET_{2,3}^2 \geq ET_{2,2}^2 + T_2 = 11 + 7 = 18$, the segment segment may not commence execution until time $18$ and the shared resource remains available to other tasks in the meantime. The third job of $\tau_1$ is released at time $16$ and acquires the uncontested shared resource at time $17$. Thus, the segment of $\tau_2$ cannot resume execution before time $a^2_{2,3} = 19$. Therefore
\begin{align*}
	ET_{2,3}^2 & = \max\left(ET_{2,2}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,3})\right) =\max(18,\ 19) = 19.
\end{align*}

The same pattern repeats for the fourth job of $\tau_2$, released at time $21$: when its first segment ends at time $23$, the second segment is not eligible to commence execution before time $26$ since $ET_{2,4}^2 \geq ET_{2,3}^2 + T_2 = 19 + 7 = 26$. By then, however, $\tau_1$ has already locked the shared semaphore again, and the second segment of the fourth job of $\tau_2$ cannot resume before time
\begin{align*}
	ET_{2,4}^2 & = \max\left(ET_{2,3}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,4})\right) =\max(26,\ 27) = 27.
\end{align*}
However, this leaves insufficient time to meet the job's deadline: the second segment of $\tau_2$ requires $C_{2,2} = 2$ time units to complete, the deadline at time $28$ is thus missed.

By construction, this example does not depend on a specific locking protocol; the effect occurs with both the MPCP~\cite{Ra:90} (based on priority queues) and the FMLP~\cite{BLBA:07,BA:08} (based on FIFO queues). However,  the response-time analyses for both the MPCP~\cite{LNR:09,Br:13} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for task $\tau_2$ (i.e., four time units of execution, and at most two time units of blocking due to the critical section of $\tau_1$). This demonstrates that adding period enforcement to suspension-based locks invalidates all existing blocking analyses. Furthermore, it is clear that the devised repeating pattern can be used to construct schedules in which the response time of $\tau_2$  grows beyond any given implicit or constrained deadline.

Next, we show that the alternative interpretation of when a lock request takes effect also leads to unbounded response time growth.

\subsection{Alternative 2: Locking Takes Effect Immediately}

### 2015-10-13 untitled.tex
add discussion of first locking counterexample

@@
 
-However, note that Theorem 5 in \cite{Raj:suspension1991} applies only to the \emph{transformed} task set: recall that, in the original analysis~\cite{Raj:suspension1991}, deferrable tasks are assumed to defer their entire execution time either completely or not at all (but not parts of it).
+However, we note that Theorem 5 in \cite{Raj:suspension1991} applies only to the \emph{transformed} task set: recall that, in the original analysis~\cite{Raj:suspension1991}, deferrable tasks are assumed to defer their entire execution time either completely or not at all (but not parts of it).
 Therefore, if we would like to use the period enforcer algorithm to handle segmented self-suspending task sets, we first have to answer the following question: ``\emph{Given a set of sporadic segmented self-suspending tasks, what is the corresponding set of (single-segment) deferrable tasks?}'' That is, how do we convert given self-suspension segments into equivalent bounds on release jitter such that we may apply Theorem 5 to conclude that the system remains schedulable despite period enforcement?

### 2015-10-13 locking.tex
add discussion of second example

@@
 \begin{enumerate}
-	\item When a task requires a shared resource, which implies the beginning of a new segment, its lock request is processed only when its new segment becomes eligible for execution (as determined by the period enforcer algorithm).
-As a consequence, a task may find a required shared resource unavailable when its new segment becomes eligible for execution even though it was available when the prior segment finished.
+	\item When a task requires a shared resource, which implies the beginning of a new segment, its lock request is processed \emph{only when its new segment would be eligible for execution}, as determined by the period enforcer algorithm.
 	\item Alternatively, when a task requires a shared resource, its request may be processed \emph{immediately}.
-As a consequence, a shared resource may be locked by a task that is currently ineligible to execute.
 \end{enumerate}
-We believe that the first interpretation is the more natural one. However, for the sake of completeness, we show that either interpretation leads to potentially unbounded worst-case response times.
+
+As a consequence of the first rule, a task may find a required shared resource unavailable when its new segment becomes eligible for execution even though it was available when the prior segment finished.
+As a consequence of the second rule, a shared resource may be locked by a task that cannot use it for the time being because it is still ineligible to execute.
+
+We believe that the first interpretation is the more natural one, as it does not make much sense to award resources to tasks that cannot yet use them.
+However, for the sake of completeness, we show that either interpretation leads to potentially unbounded worst-case response times.
 
 \subsection{Alternative 1: Locking Takes Effect at Segment Eligibility Time}
-Consider a simple task set consisting of two tasks on two processors that share one resource.
+First, assume that the processing of lock requests is delayed until the point when a resuming segment would no longer be subject to any delay due to period enforcement).
+We show that this interpretation leads to a deadline miss in a task set that would otherwise be trivially schedulable.
+
+To this end, consider a simple task set consisting of two tasks on two processors that share one resource.
 Task $\tau_1$, on processor~1, has a total execution cost of $C_1 = 4$ and a period and deadline of $T_1 = D_i = 8$.
@@
 At time $3$, $\tau_1$ releases the shared resource and $\tau_2$ consequently resumes (i.e., $a^2_{2,1} = 3$).
-According to the period enforcer rules~\cite{Raj:suspension1991}, the second segment is immediately eligible because
+According to the period enforcer rules~\cite{Raj:suspension1991}, the second segment is immediately eligible because, according to Equation~\ref{eq:ET-def},
 \begin{align*}
@@
 
-Next, we show that the alternative interpretation of when a lock request takes effect also leads to unbounded response time growth.
+Next, we show that the alternative interpretation of when a lock request takes effect also leads to unbounded response-time growth.
 
 \subsection{Alternative 2: Locking Takes Effect Immediately}
+From now on, assume that all lock requests are processed immediately when they are made, even if this causes the shared resource to be locked by a task that is not yet eligible to execute according to  the rules of the period enforcer algorithm.
+We construct an example that demonstrates unbounded response-time growth.
+
+To this end, consider two identical tasks hosted on two processors.
+Task $\tau_1$ is hosted on processor~1; $\tau_2$ is hosted on processor~2.
+Both tasks have a period of $T_1 = T_2 = D_1 = D_2 = 8$ and a WCET of $C_1 = C_2 = 4$.
+They both access a single shared resource for two time units each per job.
+Both tasks request the shared resource after executing for \emph{at most} one time unit.
+Both tasks thus have two segments each with parameters $C_{1,1} = C_{2,1} = 1$ and $C_{1,2} = C_{2,2} = 3$.
+The example exploits that jobs may require \emph{less} service than their specified worst-case costs.
+
+Figure~\ref{xxx} shows an example schedule assuming periodic job arrivals and that, for an arbitrary positive $\epsilon <1$,
+\begin{itemize}
+	\item the first segment of the first job of $\tau_2$ executes for only $1-\epsilon$ time units,
+	\item the first segment of the second job of $\tau_1$ executes for only $1-\epsilon$ time units,
+	\item the first segment of the third job  of $\tau_2$ executes for only $1-\epsilon$ time units, and that
+	\item all other segments must execute for their specified worst-case costs.
+\end{itemize}
+
+At time $1-\epsilon$, the first job of $\tau_2$ acquires the shared resource because $\tau_1$ does not issue its request until time $1$.
+Consequently, $\tau_1$ is blocked until time $a^2_{1,1} = 3$, and we have
+\begin{align*}
+	ET_{1,1}^2 & = \max\left(ET_{1,0}^2 + T_1,\ \mathit{busy}(\tau_1, a^2_{1,1})\right) =\max(-T_1 + T_1,\ 3) = 3 
+\\ \intertext{and}
+	ET_{2,1}^2 & = \max\left(ET_{2,0}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,1})\right) =\max(-T_2 + T_2,\ 1 - \epsilon) = 1 - \epsilon.
+\end{align*}
+
+The roles of the second jobs of both tasks are reversed: since the second job of $\tau_1$ locks the shared resource already at time $9-\epsilon$, $\tau_2$ is blocked when it attempts to lock the resource at time $9$.
+However, according to the rules of the period enforcer algorithm, the second segment of the second job of $\tau_1$ is not actually eligible to execute before time $11$ since
+\begin{align*}
+	ET_{1,2}^2 & = \max\left(ET_{1,1}^2 + T_1,\ \mathit{busy}(\tau_1, a^2_{1,2})\right) =\max(3 + 8,\ 8) = 11.
+\end{align*}
+Consequently, even though the lock is granted to $\tau_1$ already  at time $9-\epsilon$, the critical section is executed only starting at time $11$, and $\tau_2$ is thus delayed until time $13$.
+At time $13$, $\tau_2$ is immediately eligible to execute since
+\begin{align*}
+	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,2})\right) =\max(1-\epsilon + 8,\ 13) = 13.
+\end{align*}
+
+The third jobs of both tasks are released at time $16$.
+The roles are swapped again: $\tau_2$ acquires the lock before $\tau_1$ can issue its request because $\tau_2$'s first segment requires only $1-\epsilon$ time units of service.
+However, according to the period enforcer algorithm's eligibility criterium, $\tau_2$ cannot actually continue its execution before time $21$ since
+\begin{align*}
+	ET_{2,3}^2 & = \max\left(ET_{2,2}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,3})\right) =\max(13 + 8,\ 17-\epsilon) = 21.
+\end{align*}
+This, however, means that $\tau_1$ cannot use the shared resource before time $23$, which leaves insufficient time to complete the -segment before $\tau_1$'s deadline at time $24$.
+However, as before,  the response-time analyses for both the MPCP~\cite{Br:13,LNR:09} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for both tasks (i.e., four time units of execution, and at most two time units of blocking).
+
+\subsection{Discussion}
+
+While it is intuitively appealing to combine period enforcement with suspension-based locking protocols, we observe that this causes non-trivial difficulties.
+In particular, our examples show that the addition of the period enforcer algorithm invalidates all existing blocking analyses.
+They also suggest that devising a correct blocking analysis would be a substantial challenge due to the demonstrated feedback cycle between the period enforcer rules and blocking durations.
+Worse, such an analysis carries the risk of being an exercise in futility, as the examples show that substantial response time increases are possible---however, at which point is the period enforcer's sole benefit, namely reduced interference on lower-priority tasks, outweighed by its cost, namely the (possibly large) increases in response times of lock-using higher-priority tasks?
+
+Fundamentally, the period enforcer algorithm's design implicitly rests on the assumption that a segment \emph{can} execute as soon as it is eligible to do so.
+In the presence of locks, however, this assumption is invalidated.
+As demonstrated, the result can be a successive growth of self-suspension times until a deadline is missed.
+In a sense, the period enforcer algorithm, at least as defined in the existing literature, is therefore incompatible with suspension-based locking protocols.

### 2015-10-13 untitled.tex
add discussion of second example

@@
 
-Let $ET_{2,j}^1$ and $ET_{2,j}^2$ denote the eligibility times of the first and second segment, respectively, of the $j^{\mathrm{th}}$ job of $\tau_2$.
-Further, let $a^1_{2,j}$ and $a^2_{2,j}$ denote the arrival times\footnote{A segment \emph{arrives} when it becomes available for execution.
-The first segment arrives immediately when the job is released; the segment arrives when the job resumes from its self-suspension.} of the job's first and second segment, respectively.
-Finally, let $\mathit{busy}(\tau_2, t')$ denote the last time that a level-2 busy interval began on or prior to time $t'$.
-According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm then defines the segment eligibility times as
-\begin{align*}
-	ET_{2,j}^1 & = \max\left(ET_{2,j-1}^1 + T_2,\ \mathit{busy}(\tau_2, a^1_{2,j})\right)  \text{ and }
-\\
-	ET_{2,j}^2 & = \max\left(ET_{2,j-1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,j}) \right),
-\end{align*}
-where $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$.
+A segment's eligibility time is determined according to the following rule.
+Let $ET_{i,j}^k$ denote the eligibility time of the $k$\xth segment of the $j$\xth job of task $\tau_i$.
+Further, let $a^k_{i,j}$ denote the segment's arrival time.\footnote{A segment \emph{arrives} when it becomes available for execution.
+The first segment arrives immediately when the job is released; the segment arrives when the job resumes from its self-suspension.} Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$.
+According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
+\begin{align}\label{eq:ET-def}
+	ET_{i,j}^k & = \max\left(ET_{i,j-1}^k + T_i,\ \mathit{busy}(\tau_i, a^k_{i,j})\right),
+\end{align}
+where $ET_{i,0}^k = -T_i$.
 
-Figure \ref{fig:example} shows the resulting schedule assuming a periodic release pattern.
+Assuming this definition, Figure \ref{fig:example} shows the resulting schedule for a periodic release pattern.
 The first job of task $\tau_2$ (which arrives at time $a^1_{2,1} =  0$) is executed as if there is no period enforcement since the definition $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$ ensures that both segments are immediately eligible.

### 2015-10-14 locking.tex
add figures and update times in alternative 2 example

@@
 
-A schedule of the two tasks assuming period job arrivals is depicted in Figure~\ref{XXX}. We focus on the eligibility times $ET_{2,1}^2,ET_{2,2}^2,ET_{2,3}^2,\ldots$ of the second segment of $\tau_2$.
+A schedule of the two tasks assuming period job arrivals is depicted in Figure~\ref{fig:locking-alt1}.
+We focus on the eligibility times $ET_{2,1}^2,ET_{2,2}^2,ET_{2,3}^2,\ldots$ of the second segment of $\tau_2$.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=1]{../figures/locking/legend}
+  \includegraphics[scale=1]{../figures/locking/sched-locking-alt1}
+  \caption{Example schedule of two tasks $\tau_1$ and $\tau_2$ on two processors sharing one lock-protected resource.
+The example assumes that lock requests take effect only when the critical section segment  becomes eligible to be scheduled according to the rules of the period enforcer algorithm.
+Under this interpretation, the fourth job of task $\tau_2$ misses its deadline at time $28$.}
+  \label{fig:locking-alt1}
+  \end{figure}
 
@@
 
-Figure~\ref{xxx} shows an example schedule assuming periodic job arrivals and that, for an arbitrary positive $\epsilon <1$,
+Figure~\ref{fig:locking-alt2} shows an example schedule assuming periodic job arrivals and that, for an arbitrarily small, positive $\epsilon <1$,
 \begin{itemize}
@@
 At time $1-\epsilon$, the first job of $\tau_2$ acquires the shared resource because $\tau_1$ does not issue its request until time $1$.
-Consequently, $\tau_1$ is blocked until time $a^2_{1,1} = 3$, and we have
+Consequently, $\tau_1$ is blocked until time $a^2_{1,1} = 3 - \epsilon$, and we have
 \begin{align*}
-	ET_{1,1}^2 & = \max\left(ET_{1,0}^2 + T_1,\ \mathit{busy}(\tau_1, a^2_{1,1})\right) =\max(-T_1 + T_1,\ 3) = 3 
+	ET_{1,1}^2 & = \max\left(ET_{1,0}^2 + T_1,\ \mathit{busy}(\tau_1, a^2_{1,1})\right) =\max(-T_1 + T_1,\ 3 - \epsilon) = 3 - \epsilon
 \\ \intertext{and}
@@
 
-The roles of the second jobs of both tasks are reversed: since the second job of $\tau_1$ locks the shared resource already at time $9-\epsilon$, $\tau_2$ is blocked when it attempts to lock the resource at time $9$.
-However, according to the rules of the period enforcer algorithm, the second segment of the second job of $\tau_1$ is not actually eligible to execute before time $11$ since
+The roles of the second jobs of both tasks are reversed: since the second job of $\tau_1$ locks the shared resource already at time $9-\epsilon$, $\tau_2$ is blocked when it attempts to lock the resource at time~$9$.
+However, according to the rules of the period enforcer algorithm, the second segment of the second job of $\tau_1$ is not actually eligible to execute before time $11 - \epsilon$ since
 \begin{align*}
-	ET_{1,2}^2 & = \max\left(ET_{1,1}^2 + T_1,\ \mathit{busy}(\tau_1, a^2_{1,2})\right) =\max(3 + 8,\ 8) = 11.
+	ET_{1,2}^2 & = \max\left(ET_{1,1}^2 + T_1,\ \mathit{busy}(\tau_1, a^2_{1,2})\right) =\max(3 - \epsilon + 8,\ 8) = 11 - \epsilon.
 \end{align*}
-Consequently, even though the lock is granted to $\tau_1$ already  at time $9-\epsilon$, the critical section is executed only starting at time $11$, and $\tau_2$ is thus delayed until time $13$.
-At time $13$, $\tau_2$ is immediately eligible to execute since
+Consequently, even though the lock is granted to $\tau_1$ already  at time $9-\epsilon$, the critical section is executed only starting at time $11 - \epsilon$, and $\tau_2$ is thus delayed until time $13 - \epsilon$.
+At time $13 - \epsilon$, $\tau_2$ is immediately eligible to execute since
 \begin{align*}
-	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,2})\right) =\max(1-\epsilon + 8,\ 13) = 13.
+	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,2})\right) =\max(1-\epsilon + 8,\ 13 - \epsilon) = 13 - \epsilon.
 \end{align*}
@@
 The third jobs of both tasks are released at time $16$.
-The roles are swapped again: $\tau_2$ acquires the lock before $\tau_1$ can issue its request because $\tau_2$'s first segment requires only $1-\epsilon$ time units of service.
-However, according to the period enforcer algorithm's eligibility criterium, $\tau_2$ cannot actually continue its execution before time $21$ since
+The roles are swapped again: $\tau_2$ acquires the lock at time $a^2_{2,3} = 17 - \epsilon$, before $\tau_1$ can issue its request, because $\tau_2$'s first segment requires only $1-\epsilon$ time units of service.
+However, according to the period enforcer algorithm's eligibility criterium, $\tau_2$ cannot actually continue its execution before time $21- \epsilon$ since
 \begin{align*}
-	ET_{2,3}^2 & = \max\left(ET_{2,2}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,3})\right) =\max(13 + 8,\ 17-\epsilon) = 21.
+	ET_{2,3}^2 & = \max\left(ET_{2,2}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,3})\right) =\max(13- \epsilon + 8,\ 16) = 21- \epsilon.
 \end{align*}
-This, however, means that $\tau_1$ cannot use the shared resource before time $23$, which leaves insufficient time to complete the -segment before $\tau_1$'s deadline at time $24$.
-However, as before,  the response-time analyses for both the MPCP~\cite{Br:13,LNR:09} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for both tasks (i.e., four time units of execution, and at most two time units of blocking).
+This, however, means that $\tau_1$ cannot use the shared resource before time $23 - \epsilon$, which leaves insufficient time to complete the second segment of $\tau_1$'s third job before its deadline at time $24$.
+
+However, as in the previous example,  the response-time analyses for both the MPCP~\cite{Br:13,LNR:09} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for both tasks (i.e., four time units of execution, and at most two time units of blocking).
+The example thus demonstrates that, if lock requests take effect immediately, then the period enforcer is incompatible with existing blocking analyses because it increases the effective lock-holding times.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=1]{../figures/locking/legend}
+  \includegraphics[scale=1]{../figures/locking/sched-locking-alt2}
+  \caption{Example schedule of two tasks $\tau_1$ and $\tau_2$ on two processors sharing one lock-protected resource.
+The example assumes that lock requests take effect immediately, even if the critical section segment is not yet eligible to be scheduled according to the rules of the period enforcer algorithm.
+Under this interpretation, the third job of task $\tau_1$ misses its deadline at time $24$.}
+  \label{fig:locking-alt2}
+  \end{figure}
 

### 2015-10-14 locking.tex
revise discussion in locking section

@@
 
-A common source of self-suspensions in multiprocessor real-time systems are \emph{binary semaphores}, i.e., suspension-based locks used to realize mutually exclusive access to shared resources.
+\emph{Binary semaphores}, i.e., suspension-based locks used to realize mutually exclusive access to shared resources, are a common source of self-suspensions in multiprocessor real-time systems.
 When a task tries to use a resource that has already been locked, it self-suspends until the resource becomes available.
-Such self-suspensions due to lock contention, just like any other self-suspension, result in deferred execution and thus can detrimentally affect the interference of lower-priority tasks.
-It may thus seem natural to apply the period enforcer algorithm to control the negative effects of locking-induced self-suspensions.\footnote{The combination of suspension-based locks and period enforcement has indeed been assumed in prior work~\cite{Raj:91} and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.} However, as we demonstrate with two examples, it is actually not safe to combine the period enforcer algorithm with suspension-based locks.
+Such self-suspensions due to lock contention, just like any other self-suspension, result in deferred execution and thus can detrimentally affect a task's interference on lower-priority tasks.
+It may thus seem natural to apply the period enforcer algorithm to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91} and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.} However, as we demonstrate with two examples, it is actually not safe to use the period enforcer algorithm in the presence of suspension-based locks.
 
@@
 
-Whenever a task attempts to lock a shared resource, it may potentially block if the resource has already been  locked by another task.
-In the context of the multi-segmented self-suspending task model, each lock operations hence marks the beginning of a new segment.
+Whenever a task attempts to lock a shared resource, it may potentially block and self-suspend.
+In the context of the multi-segmented self-suspending task model, each lock request hence marks the beginning of a new segment.
 
-The period enforcer algorithm may therefore be applied to determine the eligibility time of each such segment (which, again, starts with a critical section).
+The period enforcer algorithm may therefore be applied to determine the eligibility time of each such segment (which, again, all start with a critical section).
 There is, however, one complication: when does the task actually \emph{acquire} the lock?
-That is, if a task's execution is postponed due to the rules of the period enforcement algorithm, at which point is the lock request processed, with the consequence that the resource become unavailable to other tasks?
+That is, if a task's execution is postponed due to the rules of the period enforcement algorithm, at which point is the lock request processed, with the consequence that the resource becomes unavailable to other tasks?
 
-There are two possible interpretations.
-\begin{enumerate}
-	\item When a task requires a shared resource, which implies the beginning of a new segment, its lock request is processed \emph{only when its new segment would be eligible for execution}, as determined by the period enforcer algorithm.
-	\item Alternatively, when a task requires a shared resource, its request may be processed \emph{immediately}.
-\end{enumerate}
+There are two obvious interpretations of the combined period enforcer and locking rules.
+In the \textbf{first interpretation}, when a task requires a shared resource, which implies the beginning of a new segment, its lock request is processed \emph{only when its new segment is eligible for execution}, as determined by the period enforcer algorithm.
+Alternatively, in the \textbf{second interpretation}, a task's request is processed \emph{immediately} when it requires a shared resource.
 
 As a consequence of the first rule, a task may find a required shared resource unavailable when its new segment becomes eligible for execution even though it was available when the prior segment finished.
-As a consequence of the second rule, a shared resource may be locked by a task that cannot use it for the time being because it is still ineligible to execute.
+As a consequence of the second rule, a shared resource may be locked by a task that cannot currently use it  because it is still ineligible to execute.
 
-We believe that the first interpretation is the more natural one, as it does not make much sense to award resources to tasks that cannot yet use them.
+We believe that the first interpretation is the more natural one, as it does not make much sense to allocate resources to tasks that cannot yet use them.
 However, for the sake of completeness, we show that either interpretation leads to potentially unbounded worst-case response times.
 
-\subsection{Alternative 1: Locking Takes Effect at Segment Eligibility Time}
-First, assume that the processing of lock requests is delayed until the point when a resuming segment would no longer be subject to any delay due to period enforcement).
+\subsection{Case 1: Locking Takes Effect at Earliest Segment Eligibility Time}
+In the following example, we assume the first interpretation, i.e., that the processing of lock requests is delayed until the point when a resuming segment would no longer be subject to any delay due to period enforcement.
 We show that this interpretation leads to a deadline miss in a task set that would otherwise be trivially schedulable.
@@
 After one time unit of execution, jobs of $\tau_1$ require the shared resource for two time units.
-$\tau_1$ thus consists of two segments with cost $C_{1,1} = 1$ and $C_{1,2} = 3$.
+$\tau_1$ thus consists of two segments with costs $C_{1,1} = 1$ and $C_{1,2} = 3$.
 Task $\tau_2$, on processor~2, has the same overall WCET ($C_2 = 4$), a slightly shorter period ($T_2 = D_2 = 7$), and requires the shared resource for one time unit after \emph{two} time units of execution ($C_{2,1} = 2$ and $C_{2,2} = 2$).
 
-A schedule of the two tasks assuming period job arrivals is depicted in Figure~\ref{fig:locking-alt1}.
+A schedule of the two tasks assuming periodic job arrivals is depicted in Figure~\ref{fig:locking-alt1}.
 We focus on the eligibility times $ET_{2,1}^2,ET_{2,2}^2,ET_{2,3}^2,\ldots$ of the second segment of $\tau_2$.
@@
 The third job of $\tau_2$ is released at time $14$.
-Its first segment ends at time $16$, but since $ET_{2,3}^2 \geq ET_{2,2}^2 + T_2 = 11 + 7 = 18$, the segment segment may not commence execution until time $18$ and the shared resource remains available to other tasks in the meantime.
+Its first segment ends at time $16$, but since $ET_{2,3}^2 \geq ET_{2,2}^2 + T_2 = 11 + 7 = 18$, the second segment may not commence execution until time~$18$ and the shared resource remains available to other tasks in the meantime.
 The third job of $\tau_1$ is released at time $16$ and acquires the uncontested shared resource at time $17$.
@@
 The same pattern repeats for the fourth job of $\tau_2$, released at time $21$: when its first segment ends at time $23$, the second segment is not eligible to commence execution before time $26$ since $ET_{2,4}^2 \geq ET_{2,3}^2 + T_2 = 19 + 7 = 26$.
-By then, however, $\tau_1$ has already locked the shared semaphore again, and the second segment of the fourth job of $\tau_2$ cannot resume before time
+By then, however, $\tau_1$ has already locked the shared semaphore again, and the second segment of the fourth job of $\tau_2$ cannot resume before time $a^2_{2,4} = 27$, at which point
 \begin{align*}
@@
 \end{align*}
-However, this leaves insufficient time to meet the job's deadline: the second segment of $\tau_2$ requires $C_{2,2} = 2$ time units to complete, the deadline at time $28$ is thus missed.
+However, this leaves insufficient time to meet the job's deadline: as the second segment of $\tau_2$ requires $C_{2,2} = 2$ time units to complete, the job's deadline at time~$28$ is  missed.
 
-By construction, this example does not depend on a specific locking protocol; the effect occurs with both the MPCP~\cite{Ra:90} (based on priority queues) and the FMLP~\cite{BLBA:07,BA:08} (based on FIFO queues).
-However,  the response-time analyses for both the MPCP~\cite{LNR:09,Br:13} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for task $\tau_2$ (i.e., four time units of execution, and at most two time units of blocking due to the critical section of $\tau_1$).
-This demonstrates that adding period enforcement to suspension-based locks invalidates all existing blocking analyses.
+By construction, this example does not depend on a specific locking protocol; for instance, the effect occurs with both the MPCP~\cite{Ra:90} (based on priority queues) and the FMLP~\cite{BLBA:07,BA:08} (based on FIFO queues).
+However,  the response-time analyses for both the MPCP~\cite{Br:13,LNR:09} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for task $\tau_2$ (i.e., four time units of execution, and at most two time units of blocking due to the critical section of $\tau_1$).
+This demonstrates that, under the first interpretation, adding period enforcement to suspension-based locks invalidates existing blocking analyses.
 Furthermore, it is clear that the devised repeating pattern can be used to construct schedules in which the response time of $\tau_2$  grows beyond any given implicit or constrained deadline.
 
-Next, we show that the alternative interpretation of when a lock request takes effect also leads to unbounded response-time growth.
+Next, we show that the second interpretation also leads to unbounded response-time growth.
 
-\subsection{Alternative 2: Locking Takes Effect Immediately}
-From now on, assume that all lock requests are processed immediately when they are made, even if this causes the shared resource to be locked by a task that is not yet eligible to execute according to  the rules of the period enforcer algorithm.
+\subsection{Case 2: Locking Takes Effect Immediately}
+From now on, we assume the second interpretation, i.e, that all lock requests are processed immediately when they are made, even if this causes the shared resource to be locked by a task that is not yet eligible to execute according to  the rules of the period enforcer algorithm.
 We construct an example that demonstrates unbounded response-time growth.
 
-To this end, consider two identical tasks hosted on two processors.
-Task $\tau_1$ is hosted on processor~1; $\tau_2$ is hosted on processor~2.
-Both tasks have a period of $T_1 = T_2 = D_1 = D_2 = 8$ and a WCET of $C_1 = C_2 = 4$.
+To this end, consider two tasks with identical parameters hosted on two processors.
+Task $\tau_1$ is hosted on processor~1; task $\tau_2$ is hosted on processor~2.
+Both tasks have a period and relative deadline of $T_1 = T_2 = D_1 = D_2 = 8$ and a WCET of $C_1 = C_2 = 4$.
 They both access a single shared resource for two time units each per job.
 Both tasks request the shared resource after executing for \emph{at most} one time unit.
-Both tasks thus have two segments each with parameters $C_{1,1} = C_{2,1} = 1$ and $C_{1,2} = C_{2,2} = 3$.
+They both thus have two segments each with parameters $C_{1,1} = C_{2,1} = 1$ and $C_{1,2} = C_{2,2} = 3$.
 The example exploits that jobs may require \emph{less} service than their specified worst-case costs.
@@
 	\item the first segment of the third job  of $\tau_2$ executes for only $1-\epsilon$ time units, and that
-	\item all other segments must execute for their specified worst-case costs.
+	\item all other segments execute for their specified worst-case costs.
 \end{itemize}
@@
 \\ \intertext{and}
-	ET_{2,1}^2 & = \max\left(ET_{2,0}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,1})\right) =\max(-T_2 + T_2,\ 1 - \epsilon) = 1 - \epsilon.
+	ET_{2,1}^2 & = \max\left(ET_{2,0}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,1})\right) =\max(-T_2 + T_2,\ 0) = 0.
 \end{align*}
@@
 \begin{align*}
-	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,2})\right) =\max(1-\epsilon + 8,\ 13 - \epsilon) = 13 - \epsilon.
+	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,2})\right) =\max(0 + 8,\ 13 - \epsilon) = 13 - \epsilon.
 \end{align*}
@@
 The third jobs of both tasks are released at time $16$.
-The roles are swapped again: $\tau_2$ acquires the lock at time $a^2_{2,3} = 17 - \epsilon$, before $\tau_1$ can issue its request, because $\tau_2$'s first segment requires only $1-\epsilon$ time units of service.
+The roles are swapped again: because $\tau_2$'s first segment requires only $1-\epsilon$ time units of service, it acquires the lock at time $a^2_{2,3} = 17 - \epsilon$, before $\tau_1$ issues its request at time~$17$.
 However, according to the period enforcer algorithm's eligibility criterium, $\tau_2$ cannot actually continue its execution before time $21- \epsilon$ since
@@
 
-However, as in the previous example,  the response-time analyses for both the MPCP~\cite{Br:13,LNR:09} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for both tasks (i.e., four time units of execution, and at most two time units of blocking).
-The example thus demonstrates that, if lock requests take effect immediately, then the period enforcer is incompatible with existing blocking analyses because it increases the effective lock-holding times.
+As in the previous example,  the response-time analyses for both the MPCP~\cite{Br:13,LNR:09} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for both tasks (i.e., four time units of execution, and at most two time units of blocking).
+The example thus demonstrates that, if lock requests take effect immediately, then the period enforcer is incompatible with existing blocking analyses because, under the second interpretation, it increases the effective lock-holding times.
 
@@
 While it is intuitively appealing to combine period enforcement with suspension-based locking protocols, we observe that this causes non-trivial difficulties.
-In particular, our examples show that the addition of the period enforcer algorithm invalidates all existing blocking analyses.
+In particular, our examples show that the addition of period enforcement invalidates all existing blocking analyses.
 They also suggest that devising a correct blocking analysis would be a substantial challenge due to the demonstrated feedback cycle between the period enforcer rules and blocking durations.
-Worse, such an analysis carries the risk of being an exercise in futility, as the examples show that substantial response time increases are possible---however, at which point is the period enforcer's sole benefit, namely reduced interference on lower-priority tasks, outweighed by its cost, namely the (possibly large) increases in response times of lock-using higher-priority tasks?
 
-Fundamentally, the period enforcer algorithm's design implicitly rests on the assumption that a segment \emph{can} execute as soon as it is eligible to do so.
+Fundamentally, the design of the period enforcer algorithm implicitly rests on the assumption that a segment \emph{can} execute as soon as it is eligible to do so.
 In the presence of locks, however, this assumption is invalidated.
-As demonstrated, the result can be a successive growth of self-suspension times until a deadline is missed.
-In a sense, the period enforcer algorithm, at least as defined in the existing literature, is therefore incompatible with suspension-based locking protocols.
+As demonstrated, the result can be a successive growth of self-suspension times that proceeds until a deadline is missed.
+The period enforcer algorithm, at least as defined and used in the literature to date~\cite{Raj:suspension1991,Raj:91}, is therefore incompatible with the existing literature on suspension-based real-time locking protocols (e.g., \cite{Raj:91,Lak:11,LNR:09,BLBA:07,Br:13}).

### 2015-10-14 locking.tex
disambiguate 'it'

@@
 
-There are two obvious interpretations of the combined period enforcer and locking rules.
-In the \textbf{first interpretation}, when a task requires a shared resource, which implies the beginning of a new segment, its lock request is processed \emph{only when its new segment is eligible for execution}, as determined by the period enforcer algorithm.
-Alternatively, in the \textbf{second interpretation}, a task's request is processed \emph{immediately} when it requires a shared resource.
+There are two possible interpretations of how the period enforcer and locking rules may interact.
+Under the \textbf{first interpretation}, when a task requires a shared resource, which implies the beginning of a new segment, its lock request is processed \emph{only when its new segment is eligible for execution}, as determined by the period enforcer algorithm.
+Alternatively, under the \textbf{second interpretation}, a task's request is processed \emph{immediately} when it requires a shared resource.
 
-As a consequence of the first rule, a task may find a required shared resource unavailable when its new segment becomes eligible for execution even though it was available when the prior segment finished.
-As a consequence of the second rule, a shared resource may be locked by a task that cannot currently use it  because it is still ineligible to execute.
+As a consequence of the first rule, a task may find a required shared resource unavailable when its new segment becomes eligible for execution even though the resource was available when the prior segment finished.
+As a consequence of the second rule, a shared resource may be locked by a task that cannot currently use the resource  because the task is still ineligible to execute.
 

### 2015-10-16 untitled.tex
add intro and many citations

@@
 \section{Introduction} 
-This report revisits the \emph{period enforcer} algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle self-suspending real-time tasks.
-We briefly review the period enforcer algorithm, explain its underlying assumptions and limitations, and discuss how it may be analyzed to correctly determine the schedulability of self-suspending sporadic real-time tasks subject to period enforcement.
+When real-time task self-suspend (due to blocking I/O, lock contention, etc.), they defer a part of their execution to be processed at a later time.
+A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}.
+This penalty, which is maximized when a task defers the completion of one job just until the release of the next job, can manifest as response-time increases and thus may lead to deadline misses.
 
-The main contributions of this paper are two observations that have not been previously reported in the real-time literature:
+To avoid such detrimental effects,  Rajkumar \cite{Raj:suspension1991} proposed the \emph{period enforcer} algorithm~\cite{Raj:suspension1991},  a technique to control (or shape) the processor demand of self-suspending tasks.
+In a nutshell, the period enforcer algorithm artificially increases the length of certain suspensions whenever a task's activation pattern carries the risk of inducing undue interference in lower-priority tasks.
 
+The period enforcer algorithm is worth a second look for a number of reasons.
+First, in the words of Rajkumar, it ``forces tasks to behave like ideal periodic tasks from the scheduling point of view with no associated scheduling penalties''~\cite{Raj:suspension1991}, which is obviously highly desirable in many practical applications in which self-suspensions are inevitable (e.g., when offloading computations to co-processors such as GPUs or DSPs).
+Second, the later-proposed, but more widely-known \emph{released guard} algorithm~\cite{SL:96} uses a technique quite similar to period enforcement to control scheduling penalties due to release jitter in distributed systems.
+The period enforcer algorithm has also attracted growing attention  and has been mentioned in several recent works  (e.g.,~\cite{LNR:09,LR:10,Lak:11,LC:14,KANR:13,HY:11,CA:09,CA:10,CA:10b}), at times controversially~\cite{BA:08a}.
+And last but not least, the period enforcer algorithm plays a significant role in Rajkumar's seminal book on   real-time  synchronization~\cite{Raj:91}.
+
+In this note, we revisit the period enforcer \cite{Raj:suspension1991} to carefully re-examine and explain its underlying assumptions and limitations, and to point out potential misconceptions.
+The main contributions are three observations that, to the best of our knowledge, have not been previously reported in the real-time literature:
 \begin{enumerate}
-	\item period enforcement can render self-suspending tasks sets unschedulable that are schedulable otherwise (i.e., period enforcement can induce deadline misses); and
-	\item with current techniques, schedulability analysis of the period enforcer algorithm requires a task set transformation that is subject to exponential time complexity.
+	\item period enforcement can be a cause of deadline misses in self-suspending tasks sets that are otherwise schedulable (Section~\ref{sec:unschedulable}); 
+	\item with current techniques, schedulability analysis of the period enforcer algorithm requires a task set transformation that is subject to exponential time complexity (Section~\ref{sec:convert}); and
+	\item the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed (Section~\ref{sec:locking}).
 \end{enumerate}
 
+We briefly introduce the needed background in Section~\ref{sec:prelim}, restate our contributions more precisely in Section~\ref{sec:questions}, and then explain the three above  observations in detail in Sections \ref{sec:unschedulable}--\ref{sec:locking} before concluding in Section \ref{sec:conclusion}.
+
 \subsection{Preliminaries}
+\label{sec:prelim}
 
@@
 
-It is well known that uncontrolled deferred execution (i.e, release jitter) can impose a scheduling penalty because of the potential for ``back-to-back'' execution~\cite{ABRTW:93}.
+It is well known that uncontrolled deferred execution (i.e, release jitter) can impose a scheduling penalty because of the potential for ``back-to-back'' execution~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}.
 That is, if a job of a deferrable task that maximally defers its execution is directly followed by a job that executes immediately without deferring its execution, then lower-priority tasks may suffer increased interference.
@@
 \subsection{Questions answered in this paper}
+\label{sec:questions}
 

### 2015-10-16 locking.tex

@@
 We believe that the first interpretation is the more natural one, as it does not make much sense to allocate resources to tasks that cannot yet use them.
-However, for the sake of completeness, we show that either interpretation leads to potentially unbounded worst-case response times.
+However, for the sake of completeness, we show that either interpretation leads to potentially deadline misses even if the original task systems can be feasibly scheduled without any enforcement.
 
@@
 
-To this end, consider a simple task set consisting of two tasks on two processors that share one resource.
-Task $\tau_1$, on processor~1, has a total execution cost of $C_1 = 4$ and a period and deadline of $T_1 = D_i = 8$.
+Towards this, consider a simple task set consisting of two tasks on two processors that share one resource.
+Task $\tau_1$, on processor~1, has a total execution cost of $C_1 = 4$ and a period and deadline of $T_1 = D_1 = 8$.
 After one time unit of execution, jobs of $\tau_1$ require the shared resource for two time units.
@@
 Task $\tau_2$, on processor~2, has the same overall WCET ($C_2 = 4$), a slightly shorter period ($T_2 = D_2 = 7$), and requires the shared resource for one time unit after \emph{two} time units of execution ($C_{2,1} = 2$ and $C_{2,2} = 2$).
+We will explain later why these two tasks can be feasibly scheduled without any enforcement.
 
-A schedule of the two tasks assuming periodic job arrivals is depicted in Figure~\ref{fig:locking-alt1}.
+A schedule of the two tasks assuming periodic job arrivals and using the period enforcer is depicted in Figure~\ref{fig:locking-alt1}.
 We focus on the eligibility times $ET_{2,1}^2,ET_{2,2}^2,ET_{2,3}^2,\ldots$ of the second segment of $\tau_2$.
@@
 By construction, this example does not depend on a specific locking protocol; for instance, the effect occurs with both the MPCP~\cite{Ra:90} (based on priority queues) and the FMLP~\cite{BLBA:07,BA:08} (based on FIFO queues).
-However,  the response-time analyses for both the MPCP~\cite{Br:13,LNR:09} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for task $\tau_2$ (i.e., four time units of execution, and at most two time units of blocking due to the critical section of $\tau_1$).
+The response-time analyses for both the MPCP~\cite{Br:13,LNR:09} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for task $\tau_2$ (i.e., four time units of execution, and at most two time units of blocking due to the critical section of $\tau_1$).
 This demonstrates that, under the first interpretation, adding period enforcement to suspension-based locks invalidates existing blocking analyses.
@@
 
-Next, we show that the second interpretation also leads to unbounded response-time growth.
-
 \subsection{Case 2: Locking Takes Effect Immediately}
-From now on, we assume the second interpretation, i.e, that all lock requests are processed immediately when they are made, even if this causes the shared resource to be locked by a task that is not yet eligible to execute according to  the rules of the period enforcer algorithm.
+Next, we show that the second interpretation may lead to unbounded response-time growth.
+ From now on, we assume the second interpretation, i.e., all lock requests are processed immediately when they are made, even if this causes the shared resource to be locked by a task that is not yet eligible to execute according to  the rules of the period enforcer algorithm.
 We construct an example that demonstrates unbounded response-time growth.
 
-To this end, consider two tasks with identical parameters hosted on two processors.
+Towards this, consider two tasks with identical parameters hosted on two processors.
 Task $\tau_1$ is hosted on processor~1; task $\tau_2$ is hosted on processor~2.
-Both tasks have a period and relative deadline of $T_1 = T_2 = D_1 = D_2 = 8$ and a WCET of $C_1 = C_2 = 4$.
+Both tasks have the same period and relative deadline $T_1 = T_2 = D_1 = D_2 = 8$ and the same WCET of $C_1 = C_2 = 4$.
 They both access a single shared resource for two time units each per job.
@@
 They both thus have two segments each with parameters $C_{1,1} = C_{2,1} = 1$ and $C_{1,2} = C_{2,2} = 3$.
-The example exploits that jobs may require \emph{less} service than their specified worst-case costs.
 
+The example exploits that the jobs may require \emph{less} service than their specified worst-case execution times.
+We will use a deterministic pattern of the actual execution times to ensure the ``deterministic behaviour'' for accessing and being granted to use the shared resources.
+Let $\epsilon$ be an arbitrarily small and positive real number with $\epsilon <1$.
+\begin{itemize}
+	\item The first segment of the 1st, 3rd, 5th, $\ldots$, $(2i-1)$-th jobs of $\tau_2$ executes for only $1-\epsilon$ time units, $\forall i=1,2,\ldots$.
+	\item The first segment of the 2nd, 4th, 6th, $\ldots$, $(2i)$-th jobs of $\tau_2$ executes for only $1-\epsilon$ time units, $\forall i=1,2,\ldots$.
+	\item All the other segments execute for their specified worst-case costs.
+\end{itemize}
 Figure~\ref{fig:locking-alt2} shows an example schedule assuming periodic job arrivals and that, for an arbitrarily small, positive $\epsilon <1$,
-\begin{itemize}
-	\item the first segment of the first job of $\tau_2$ executes for only $1-\epsilon$ time units,
-	\item the first segment of the second job of $\tau_1$ executes for only $1-\epsilon$ time units,
-	\item the first segment of the third job  of $\tau_2$ executes for only $1-\epsilon$ time units, and that
-	\item all other segments execute for their specified worst-case costs.
-\end{itemize}
 
@@
 This, however, means that $\tau_1$ cannot use the shared resource before time $23 - \epsilon$, which leaves insufficient time to complete the second segment of $\tau_1$'s third job before its deadline at time $24$.
+If we keep the above execution pattern, the period enforcer continues to further increase the response time of these two tasks. This, at the end, leads to unbounded worst-case response time.
 

### 2015-10-16 untitled.tex
update intro

@@
 However, we note that Theorem 5 in \cite{Raj:suspension1991} applies only to the \emph{transformed} task set: recall that, in the original analysis~\cite{Raj:suspension1991}, deferrable tasks are assumed to defer their entire execution time either completely or not at all (but not parts of it).
-Therefore, if we would like to use the period enforcer algorithm to handle segmented self-suspending task sets, we first have to answer the following question: ``\emph{Given a set of sporadic segmented self-suspending tasks, what is the corresponding set of (single-segment) deferrable tasks?}'' That is, how do we convert given self-suspension segments into equivalent bounds on release jitter such that we may apply Theorem 5 to conclude that the system remains schedulable despite period enforcement?
+Therefore, if we would like to apply the period enforcer to segmented self-suspending task sets, we first have to answer the following question: ``\emph{Given a set of sporadic segmented self-suspending tasks, what is the corresponding set of (single-segment) deferrable tasks?}'' That is, how do we convert given self-suspension segments into equivalent bounds on release jitter such that we may apply Theorem 5 to conclude that the system remains schedulable despite period enforcement?
 
@@
 \cite{ecrts15nelissen} can be applied, but their method takes exponential time.
+\end{enumerate}
+%
+Finally, we make a third observation concerning the use of the period enforcer in conjunction with suspension-based multiprocessor locking protocols (such as the MPCP~\cite{LNR:09,Ra:90} or the FMLP~\cite{BLBA:07,BA:08}): while it is certainly tempting to apply period enforcement with the intention of avoiding the negative effects of deferred execution due to lock contention, it is actually unsafe to do so.
+\begin{enumerate}
+\setcounter{enumi}{2}
+	\item The period enforcer algorithm invalidates all existing blocking analyses as there exist non-trivial feedback cycles between the period enforcer rules and blocking durations.
 \end{enumerate}

### 2015-10-16 conclusion.tex
add intro and many citations
revise conclusion to mention locking observation

@@
 \section{Concluding Remarks}
+\label{sec:conclusion}
 
-We have revisited the period enforcer algorithm proposed by Rajkumar \cite{Raj:suspension1991} to handle segmented self-suspending real-time tasks.
+We have revisited the underlying assumptions and limitations of the period enforcer algorithm, which Rajkumar \cite{Raj:suspension1991} introduced to handle segmented self-suspending real-time tasks.
+
 One key assumption in the original proposal \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
@@
 
-Furthermore, the original proposal \cite{Raj:suspension1991} left open the question of how to convert a segmented self-suspending task system to a corresponding deferrable task system.
+Furthermore, the original proposal \cite{Raj:suspension1991} left open the question of how to convert a segmented self-suspending task set to a corresponding set of deferrable tasks.
 Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a task set transformation is non-trivial in the general case.
 
-Nevertheless, Theorem 5 in \cite{Raj:suspension1991} can be useful for handling self-suspending task systems if there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm.
+Finally, we have demonstrated that substantial difficulties arise if one attempts to combine suspension-based locks with period enforcement.
+These difficulties stem from the fact that period enforcement can increase contention, which increases self-suspension times, which then in turn feeds back into the period enforcer's minimum suspension lengths.
+As a consequence, period enforcement invalidates all existing blocking analyses.
+
+Nevertheless, Theorem 5 in \cite{Raj:suspension1991} could be useful for handling self-suspending tasks (that do not use suspension-based locks) if there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm.
 However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.

### 2015-10-16 abstract.tex
edits and tweaks

\begin{abstract}
  To be written...
\end{abstract}

### 2015-10-16 locking.tex
edits and tweaks

--- 2015-10-14 locking.tex
@@
 Such self-suspensions due to lock contention, just like any other self-suspension, result in deferred execution and thus can detrimentally affect a task's interference on lower-priority tasks.
-It may thus seem natural to apply the period enforcer algorithm to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91} and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.} However, as we demonstrate with two examples, it is actually not safe to use the period enforcer algorithm in the presence of suspension-based locks.
+It may thus seem natural to apply the period enforcer  to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91} and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.} However, as we demonstrate with two examples, it is actually not safe to use the period enforcer in the presence of suspension-based locks.
 
@@
 The period enforcer algorithm may therefore be applied to determine the eligibility time of each such segment (which, again, all start with a critical section).
-There is, however, one complication: when does the task actually \emph{acquire} the lock?
-That is, if a task's execution is postponed due to the rules of the period enforcement algorithm, at which point is the lock request processed, with the consequence that the resource becomes unavailable to other tasks?
+There is, however, one complication: when does a task actually \emph{acquire} a lock?
+That is, if a task's execution is postponed due to the period enforcement rules, at which point is the lock request processed, with the consequence that the resource becomes unavailable to other tasks?
 
-There are two possible interpretations of how the period enforcer and locking rules may interact.
+There are two possible interpretations of how period enforcement and locking rules may interact.
 Under the \textbf{first interpretation}, when a task requires a shared resource, which implies the beginning of a new segment, its lock request is processed \emph{only when its new segment is eligible for execution}, as determined by the period enforcer algorithm.
@@
 We believe that the first interpretation is the more natural one, as it does not make much sense to allocate resources to tasks that cannot yet use them.
-However, for the sake of completeness, we show that either interpretation leads to potentially unbounded worst-case response times.
+However, for the sake of completeness, we show that either interpretation can lead to deadline misses even if the task set trivially schedulable without any enforcement.
 
@@
 
-To this end, consider a simple task set consisting of two tasks on two processors that share one resource.
-Task $\tau_1$, on processor~1, has a total execution cost of $C_1 = 4$ and a period and deadline of $T_1 = D_i = 8$.
+Consider the following simple task set consisting of two tasks on two processors that share one resource.
+Task $\tau_1$, on processor~1, has a total execution cost of $C_1 = 4$ and a period and deadline of $T_1 = D_1 = 8$.
 After one time unit of execution, jobs of $\tau_1$ require the shared resource for two time units.
@@
 Task $\tau_2$, on processor~2, has the same overall WCET ($C_2 = 4$), a slightly shorter period ($T_2 = D_2 = 7$), and requires the shared resource for one time unit after \emph{two} time units of execution ($C_{2,1} = 2$ and $C_{2,2} = 2$).
+Without period enforcement (and under any reasonable locking protocol), the task set is trivially schedulable because, by construction, any job of $\tau_1$ incurs at most one time unit of blocking, and any job of $\tau_2$ incurs at most two time units of blocking.
 
-A schedule of the two tasks assuming periodic job arrivals is depicted in Figure~\ref{fig:locking-alt1}.
+In contrast, with period enforcement, deadline misses are possible.
+Figure~\ref{fig:locking-alt1} depicts a schedule of the two tasks assuming periodic job arrivals and use of the period enforcer algorithm.
 We focus on the eligibility times $ET_{2,1}^2,ET_{2,2}^2,ET_{2,3}^2,\ldots$ of the second segment of $\tau_2$.
@@
 By construction, this example does not depend on a specific locking protocol; for instance, the effect occurs with both the MPCP~\cite{Ra:90} (based on priority queues) and the FMLP~\cite{BLBA:07,BA:08} (based on FIFO queues).
-However,  the response-time analyses for both the MPCP~\cite{Br:13,LNR:09} and the   FMLP~\cite{Br:13} predict a worst-case response time of $6$ for task $\tau_2$ (i.e., four time units of execution, and at most two time units of blocking due to the critical section of $\tau_1$).
+The corresponding response-time analyses for both protocols~\cite{Br:13,LNR:09} predict a worst-case response time of $6$ for task $\tau_2$ (i.e., four time units of execution, and at most two time units of blocking due to the critical section of $\tau_1$).
 This demonstrates that, under the first interpretation, adding period enforcement to suspension-based locks invalidates existing blocking analyses.
@@
 
-Next, we show that the second interpretation also leads to unbounded response-time growth.
+Next, we show that the second interpretation can also lead to deadline misses in otherwise trivially schedulable task sets.
 
 \subsection{Case 2: Locking Takes Effect Immediately}
-From now on, we assume the second interpretation, i.e, that all lock requests are processed immediately when they are made, even if this causes the shared resource to be locked by a task that is not yet eligible to execute according to  the rules of the period enforcer algorithm.
-We construct an example that demonstrates unbounded response-time growth.
+ From now on, we assume the second interpretation: all lock requests are processed immediately when they are made, even if this causes the shared resource to be locked by a task that is not yet eligible to execute according to  the rules of the period enforcer algorithm.
+We construct an example that demonstrates \emph{unbounded} response-time growth.
 
@@
 Task $\tau_1$ is hosted on processor~1; task $\tau_2$ is hosted on processor~2.
-Both tasks have a period and relative deadline of $T_1 = T_2 = D_1 = D_2 = 8$ and a WCET of $C_1 = C_2 = 4$.
+Both tasks have the same period and relative deadline $T_1 = T_2 = D_1 = D_2 = 8$ and the same WCET of $C_1 = C_2 = 4$.
 They both access a single shared resource for two time units each per job.
@@
 They both thus have two segments each with parameters $C_{1,1} = C_{2,1} = 1$ and $C_{1,2} = C_{2,2} = 3$.
-The example exploits that jobs may require \emph{less} service than their specified worst-case costs.
 
-Figure~\ref{fig:locking-alt2} shows an example schedule assuming periodic job arrivals and that, for an arbitrarily small, positive $\epsilon <1$,
+The example exploits that a job may require \emph{less} service than its task's specified WCET.
+To ensure that the shared resource is acquired in a certain order, we assume the following deterministic pattern of the actual execution times.
+Let $\epsilon$ be an arbitrarily small, positive real number with $\epsilon <1$.
 \begin{itemize}
-	\item the first segment of the first job of $\tau_2$ executes for only $1-\epsilon$ time units,
-	\item the first segment of the second job of $\tau_1$ executes for only $1-\epsilon$ time units,
-	\item the first segment of the third job  of $\tau_2$ executes for only $1-\epsilon$ time units, and that
-	\item all other segments execute for their specified worst-case costs.
+	\item The first segment of even-numbered jobs  of $\tau_1$ executes for only $1-\epsilon$ time units.
+	\item The first segment of odd-numbered jobs of $\tau_2$ executes for only $1-\epsilon$ time units.
+%	the \nth{1}, \nth{3}, \nth{5}, \ldots, $(2i-1)$\xth job of , $\forall i=1,2,\ldots$.
+%	the \nth{2}, \nth{4}, \nth{6}, \ldots, $(2i)$\xth jobs, $\forall i=1,2,\ldots$.
+	\item All other segments execute for their specified worst-case costs.
 \end{itemize}
+Figure~\ref{fig:locking-alt2} shows an example schedule assuming periodic job arrivals.
 
@@
 This, however, means that $\tau_1$ cannot use the shared resource before time $23 - \epsilon$, which leaves insufficient time to complete the second segment of $\tau_1$'s third job before its deadline at time $24$.
+Furthermore, if both tasks continue the illustrated execution pattern, the period enforcer continues to increase their response times. As a result,  unbounded worst-case response times may arise.
 
@@
 The period enforcer algorithm, at least as defined and used in the literature to date~\cite{Raj:suspension1991,Raj:91}, is therefore incompatible with the existing literature on suspension-based real-time locking protocols (e.g., \cite{Raj:91,Lak:11,LNR:09,BLBA:07,Br:13}).
+
+Finally, it is worth noting that our examples can be trivially extended with lower-priority tasks to ensure that no processor idles before the described deadline misses occur.
+It is also not difficult to extend the second example with a task on a third processor such that all segments of $\tau_1$ and $\tau_2$ are separated by a non-zero self-suspension.

### 2015-10-16 locking.tex
minor fixes

@@
 We believe that the first interpretation is the more natural one, as it does not make much sense to allocate resources to tasks that cannot yet use them.
-However, for the sake of completeness, we show that either interpretation can lead to deadline misses even if the task set trivially schedulable without any enforcement.
+However, for the sake of completeness, we show that either interpretation can lead to deadline misses even if the task set is trivially schedulable without any enforcement.
 

### 2015-10-16 untitled.tex
minor fixes

@@
 Second, the later-proposed, but more widely-known \emph{released guard} algorithm~\cite{SL:96} uses a technique quite similar to period enforcement to control scheduling penalties due to release jitter in distributed systems.
-The period enforcer algorithm has also attracted growing attention  and has been mentioned in several recent works  (e.g.,~\cite{LNR:09,LR:10,Lak:11,LC:14,KANR:13,HY:11,CA:09,CA:10,CA:10b}), at times controversially~\cite{BA:08a}.
+The period enforcer algorithm has also attracted renewed attention in recent years and has been discussed in several current works  (e.g.,~\cite{LNR:09,LR:10,Lak:11,LC:14,KANR:13,HY:11,CA:09,CA:10,CA:10b}), at times controversially~\cite{BA:08a}.
 And last but not least, the period enforcer algorithm plays a significant role in Rajkumar's seminal book on   real-time  synchronization~\cite{Raj:91}.
@@
 
-\subsection{Questions answered in this paper}
+\subsection{Questions Answered in This Paper}
 \label{sec:questions}

### 2015-10-16 figures/example/example-body.tex
revise figures
change the units of x axis… (strange setting of my student)

@@
 
-		\def\ux{0.4cm}\def\uy{0.75cm} 
-		\begin{tikzpicture}[y=\uy, font=\sffamily,thick]  
+		\def\ux{1.1cm}\def\uy{0.75cm} 
+		\begin{tikzpicture}[x=\ux,y=\uy, font=\sffamily,thick]  
 		\tikzset{
@@
 
-		\node[task7,minimum width=2.5*\ux, 
+		\node[task7,minimum width=\ux, 
                 anchor=south west] at (0,0){};
-		\node[task7, minimum width=2.5*\ux,
+		\node[task7, minimum width=\ux,
 		anchor=south west]at (5, 0){};
-		\node[task7, minimum width=2.5*\ux,
+		\node[task7, minimum width=\ux,
 		anchor=south west]at (10, 0){};
@@
 
-		\node[task7, minimum width=1.25*\ux,
+		\node[task7, minimum width=.5*\ux,
 		anchor=south west] at (1, 0){\tiny $C_2^1$};
@@
                 node[anchor=south]{\scriptsize suspension} -- (4.5, 0.3);
-		\node[task7, minimum width=1.25*\ux,
+		\node[task7, minimum width=.5*\ux,
 		anchor=south west] at (4.5, 0){\tiny $C_2^2$};
 
-		\node[task7, minimum width=1.25*\ux,
+		\node[task7, minimum width=.5*\ux,
 		anchor=south west] at (6, 0){\tiny $C_2^1$};
@@
 
-		\node[task7, minimum width=1.25*\ux,
+		\node[task7, minimum width=.5*\ux,
 		anchor=south west] at (11, 0){\tiny $C_2^2$};
-		\node[task7, minimum width=1.25*\ux,
+		\node[task7, minimum width=.5*\ux,
 		anchor=south west] at (11.5, 0){\tiny $C_2^1$};

### 2015-10-17 untitled.tex
change pictures
restructure for sync with Authorea
try another way
proofreading fixes

@@
 \section{Introduction} 
-When real-time task self-suspend (due to blocking I/O, lock contention, etc.), they defer a part of their execution to be processed at a later time.
+When real-time tasks self-suspend (due to blocking I/O, lock contention, etc.), they defer a part of their execution to be processed at a later time.
 A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}.
@@
 \begin{enumerate}
-	\item period enforcement can be a cause of deadline misses in self-suspending tasks sets that are otherwise schedulable (Section~\ref{sec:unschedulable}); 
+	\item period enforcement can be a cause of deadline misses in self-suspending task sets that are otherwise schedulable (Section~\ref{sec:unschedulable}); 
 	\item with current techniques, schedulability analysis of the period enforcer algorithm requires a task set transformation that is subject to exponential time complexity (Section~\ref{sec:convert}); and
@@
 
+\ifpaper
 \begin{figure}[t]
@@
   \end{figure}
+\begin{figure}[t]
+  \centering
+  \input{../figures/example/example-body.tex}
+  \caption{An illustrative example demonstrating a deadline miss at time 22
+    under the period enforcer algorithm.}
+\label{fig:example}  
+\end{figure}
+\fi
 
@@
 Further, let $a^k_{i,j}$ denote the segment's arrival time.\footnote{A segment \emph{arrives} when it becomes available for execution.
-The first segment arrives immediately when the job is released; the segment arrives when the job resumes from its self-suspension.} Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$.
+The first segment arrives immediately when the job is released; the second segment arrives when the job resumes from its self-suspension.} Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$.
 According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
@@
 
-\begin{figure}[t]
-  \centering
-  \input{../figures/example/example-body.tex}
-  \caption{An illustrative example demonstrating a deadline miss at time 22
-    under the period enforcer algorithm.}
-\label{fig:example}  
-\end{figure}
-
 This example shows that there exist sporadic segmented self-suspending task sets that   \textbf{(i)} are schedulable under fixed-priority scheduling without any enforcement, but \textbf{(ii)} are not schedulable under the period enforcer algorithm.

### 2015-10-17 abstract.tex
add abstract

@@
 \begin{abstract}
-  To be written...
+The \emph{period enforcer} algorithm for self-suspending real-time tasks is a technique for suppressing the ``back-to-back'' scheduling penalty associated with deferred execution.
+Originally proposed in 1991, the algorithm has attracted renewed interest in recent years.
+This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~with current techniques, schedulability analysis of the period enforcer algorithm requires a task set transformation that is subject to exponential time complexity; and \textbf{(iii)} the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
 \end{abstract}

### 2015-10-17 LITES/LITES-Paper.tex
Some minor polishing and shuffling around of the figures
proofreading fixes and tweaks
add discussion of first locking counterexample
restructure for sync with Authorea
try another way
add subject classification

@@
 
+\papertrue 
 \title{A Note on the Period Enforcer Algorithm  for Self-Suspending Tasks}
@@
 \author[1]{Jian-Jia Chen}
-\author[2]{Bj\"orn Brandenburg}
+\author[2]{Bj\"orn B.\ Brandenburg}
 \affil[1]{TU Dortmund University, Germany\\
@@
 
-%\subjclass{MANDATORY:  Please refer to \url{www.acm.org/about/class/2012}}% mandatory: Please choose 2012 ACM classifications from http://www.acm.org/about/class/2012 .
+\subjclass{Real-time systems, Real-time schedulability, Synchronization}% mandatory: Please choose 2012 ACM classifications from http://www.acm.org/about/class/2012 .
 E.g., cite as "Integrated circuits".
-%\keywords{MANDATORY: Please provide 1--5 keywords as a comma-separated list}% mandatory: Please provide 1-5 keywords
+\keywords{period enforcer, deferred execution, self-suspension, blocking}% mandatory: Please provide 1-5 keywords
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \input{../untitled.tex}
-\begin{figure}[t]
-  \centering
-  \input{../figures/example-original/example-body.tex}
-  \caption{An illustrative example of the original self-suspending task schedule.}
-  \label{fig:example-original}
-  \end{figure}
 
-\begin{figure}[t]
-  \centering
-  \input{../figures/example/example-body.tex}
-  \caption{An illustrative example demonstrating a deadline miss at time 22
-    under the period enforcer algorithm.}
-\label{fig:example}  
-\end{figure}
 %\input{../figures/example-original/example.png}
@@
 \input{../transformation.tex}
+\input{../locking.tex}
 \input{../conclusion.tex}

### 2015-10-17 locking.tex
restructure for sync with Authorea
try another way
integrate to author
force pdflatex to pick up the .pdf files (picks .png otherwise)
don't say unbounded, and don't speculate what happens for arbitrary deadlines D_i > T_i

@@
 
+\ifpaper
 \begin{figure}[t]
   \centering
-  \includegraphics[scale=1]{../figures/locking/legend}
-  \includegraphics[scale=1]{../figures/locking/sched-locking-alt1}
+  \includegraphics[scale=1]{../figures/locking/legend.pdf}
+  \includegraphics[scale=1]{../figures/locking/sched-locking-alt1.pdf}
   \caption{Example schedule of two tasks $\tau_1$ and $\tau_2$ on two processors sharing one lock-protected resource.
@@
   \end{figure}
+\fi
 
@@
  From now on, we assume the second interpretation: all lock requests are processed immediately when they are made, even if this causes the shared resource to be locked by a task that is not yet eligible to execute according to  the rules of the period enforcer algorithm.
-We construct an example that demonstrates \emph{unbounded} response-time growth.
+We construct an example in which a task's response time grows with each job until a deadline is missed.
 
@@
 This, however, means that $\tau_1$ cannot use the shared resource before time $23 - \epsilon$, which leaves insufficient time to complete the second segment of $\tau_1$'s third job before its deadline at time $24$.
-Furthermore, if both tasks continue the illustrated execution pattern, the period enforcer continues to increase their response times. As a result,  unbounded worst-case response times may arise.
+Furthermore, if both tasks continue the illustrated execution pattern, the period enforcer continues to increase their response times.
+As a result, the pattern may be repeated to construct schedules in which an arbitrarily large implicit or constrained deadline is violated.
 
@@
 
+\ifpaper
 \begin{figure}[t]
   \centering
-  \includegraphics[scale=1]{../figures/locking/legend}
-  \includegraphics[scale=1]{../figures/locking/sched-locking-alt2}
+  \includegraphics[scale=1]{../figures/locking/legend.pdf}
+  \includegraphics[scale=1]{../figures/locking/sched-locking-alt2.pdf}
   \caption{Example schedule of two tasks $\tau_1$ and $\tau_2$ on two processors sharing one lock-protected resource.

### 2015-10-17 locking.tex
update notation in sec 4

@@
 After one time unit of execution, jobs of $\tau_1$ require the shared resource for two time units.
-$\tau_1$ thus consists of two segments with costs $C_{1,1} = 1$ and $C_{1,2} = 3$.
-Task $\tau_2$, on processor~2, has the same overall WCET ($C_2 = 4$), a slightly shorter period ($T_2 = D_2 = 7$), and requires the shared resource for one time unit after \emph{two} time units of execution ($C_{2,1} = 2$ and $C_{2,2} = 2$).
+$\tau_1$ thus consists of two segments with costs $C_1^1 = 1$ and $C_1^2 = 3$.
+Task $\tau_2$, on processor~2, has the same overall WCET ($C_2 = 4$), a slightly shorter period ($T_2 = D_2 = 7$), and requires the shared resource for one time unit after \emph{two} time units of execution ($C_2^1 = 2$ and $C_2^2 = 2$).
 Without period enforcement (and under any reasonable locking protocol), the task set is trivially schedulable because, by construction, any job of $\tau_1$ incurs at most one time unit of blocking, and any job of $\tau_2$ incurs at most two time units of blocking.
@@
 \end{align*}
-However, this leaves insufficient time to meet the job's deadline: as the second segment of $\tau_2$ requires $C_{2,2} = 2$ time units to complete, the job's deadline at time~$28$ is  missed.
+However, this leaves insufficient time to meet the job's deadline: as the second segment of $\tau_2$ requires $C_2^2 = 2$ time units to complete, the job's deadline at time~$28$ is  missed.
 
@@
 Both tasks request the shared resource after executing for \emph{at most} one time unit.
-They both thus have two segments each with parameters $C_{1,1} = C_{2,1} = 1$ and $C_{1,2} = C_{2,2} = 3$.
+They both thus have two segments each with parameters $C_1^1 = C_2^1 = 1$ and $C_1^2 = C_2^2 = 3$.
 

### 2015-10-17 abstract.tex

--- 2015-10-16 abstract.tex
@@
 \begin{abstract}
-  To be written...
+The \emph{period enforcer} algorithm for self-suspending real-time tasks is a technique for suppressing the ``back-to-back'' scheduling penalty associated with deferred execution.
+Originally proposed in 1991, the algorithm has attracted renewed interest in recent years.
+This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~with current techniques, schedulability analysis of the period enforcer algorithm requires a proper transformation, that is subject to exponential time complexity, of the task set to satisfy the assumptions; and \textbf{(iii)} the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
 \end{abstract}

### 2015-10-17 figures/example/example-body.tex

@@
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (1, 0){\tiny $C_2^1$};
+		anchor=south west] at (1, 0) {}; \node[anchor=south
+                west]at (1, 1.05*\uy) {\tiny $C_{2,1}$};
 		\draw[(-), thin] (1.5, 0.3) -- (3, 0.3)
@@
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (4.5, 0){\tiny $C_2^2$};
+		anchor=south west] at (4.5, 0) {}; \node[anchor=south
+                west]at (4.5, 1.05*\uy) {\tiny $C_{2,2}$};
 
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (6, 0){\tiny $C_2^1$};
+		anchor=south west] at (6, 0) {}; \node[anchor=south
+                west]at (6, 1.05*\uy) {\tiny $C_{2,1}$};
 		\draw[(-), thin] (6.5, 0.3) -- (8, 0.3)
@@
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (11, 0){\tiny $C_2^2$};
+		anchor=south west] at (11, 0) {}; \node[anchor=south
+                west]at (11, 1.05*\uy) {\tiny $C_{2,2}$};
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (11.5, 0){\tiny $C_2^1$};
+		anchor=south west] at (11.5, 0) {}; \node[anchor=south
+                west]at (11.5, 1.05*\uy) {\tiny $C_{2,1}$};
 
@@
                 }	
-                \draw (11,1.1) node[anchor=south] {\scriptsize\textcolor{red}{deadline miss}};
+                \draw (11,1.7) node[anchor=south] {\scriptsize\textcolor{red}{deadline miss}};
 		\end{scope}

### 2015-10-17 untitled.tex

@@
 \section{Introduction} 
-When real-time tasks self-suspend (due to blocking I/O, lock contention, etc.), they defer a part of their execution to be processed at a later time.
+When real-time tasks suspend themselves (due to blocking I/O, lock contention, etc.), they defer a part of their execution to be processed at a later time.
 A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}.
@@
 Second, the later-proposed, but more widely-known \emph{released guard} algorithm~\cite{SL:96} uses a technique quite similar to period enforcement to control scheduling penalties due to release jitter in distributed systems.
-The period enforcer algorithm has also attracted renewed attention in recent years and has been discussed in several current works  (e.g.,~\cite{LNR:09,LR:10,Lak:11,LC:14,KANR:13,HY:11,CA:09,CA:10,CA:10b}), at times controversially~\cite{BA:08a}.
+The period enforcer algorithm has also attracted renewed attention in recent years and has been discussed in several current works  (e.g.,~\cite{DBLP:conf/rtss/ChenL14,LNR:09,LR:10,Lak:11,LC:14,KANR:13,HY:11,CA:09,CA:10,CA:10b}), at times controversially~\cite{BA:08a}.
 And last but not least, the period enforcer algorithm plays a significant role in Rajkumar's seminal book on   real-time  synchronization~\cite{Raj:91}.
@@
 In this note, we revisit the period enforcer \cite{Raj:suspension1991} to carefully re-examine and explain its underlying assumptions and limitations, and to point out potential misconceptions.
-The main contributions are three observations that, to the best of our knowledge, have not been previously reported in the real-time literature:
+The main contributions are three observations that, to the best of our knowledge, have not been previously reported in the literature of real-time systems:
 \begin{enumerate}
 	\item period enforcement can be a cause of deadline misses in self-suspending task sets that are otherwise schedulable (Section~\ref{sec:unschedulable}); 
-	\item with current techniques, schedulability analysis of the period enforcer algorithm requires a task set transformation that is subject to exponential time complexity (Section~\ref{sec:convert}); and
+	\item with current techniques, schedulability analysis of the period enforcer algorithm requires a proper transformation, that is subject to exponential time complexity, of a self-suspending task set to satisfy the assumption of the period enforcer (Section~\ref{sec:convert}); and
 	\item the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed (Section~\ref{sec:locking}).

### 2015-10-17 conclusion.tex
slightly touch conclusion

@@
 Finally, we have demonstrated that substantial difficulties arise if one attempts to combine suspension-based locks with period enforcement.
-These difficulties stem from the fact that period enforcement can increase contention, which increases self-suspension times, which then in turn feeds back into the period enforcer's minimum suspension lengths.
+These difficulties stem from the fact that period enforcement can increase contention, which increases the lengths of self-suspension intervals, which then in turn feeds back into the period enforcer's minimum suspension lengths.
 As a consequence, period enforcement invalidates all existing blocking analyses.

### 2015-10-17 figures/example/example-body.tex

--- 2015-10-16 figures/example/example-body.tex
@@
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (1, 0){\tiny $C_2^1$};
+		anchor=south west] at (1, 0) {}; \node[anchor=south
+                west]at (1, 1.05*\uy) {\tiny $C_2^1$};
 		\draw[(-), thin] (1.5, 0.3) -- (3, 0.3)
@@
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (4.5, 0){\tiny $C_2^2$};
+		anchor=south west] at (4.5, 0) {}; \node[anchor=south
+                west]at (4.5, 1.05*\uy) {\tiny $C_2^2$};
 
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (6, 0){\tiny $C_2^1$};
+		anchor=south west] at (6, 0) {}; \node[anchor=south
+                west]at (6, 1.05*\uy) {\tiny $C_2^1$};
 		\draw[(-), thin] (6.5, 0.3) -- (8, 0.3)
@@
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (11, 0){\tiny $C_2^2$};
+		anchor=south west] at (11, 0) {}; \node[anchor=south
+                west]at (11, 1.05*\uy) {\tiny $C_2^2$};
 		\node[task7, minimum width=.5*\ux,
-		anchor=south west] at (11.5, 0){\tiny $C_2^1$};
+		anchor=south west] at (11.5, 0) {}; \node[anchor=south
+                west]at (11.5, 1.05*\uy) {\tiny $C_2^1$};
 

### 2015-10-17 untitled.tex

@@
 To this end, consider a task system consisting of $2$ tasks.
-Let $\tau_1$ denote a sporadic task without self-suspensions and parameters $C_1 = 2$ and $T_1=D_1=10$, and let $\tau_2$ denote a self-suspending task consisting of two segments with parameters  $C_{2,1} = 1$,  $S_{2,1} = 6$, $C_{2,2}=1$, and $ T_2=D_2=11$.
+Let $\tau_1$ denote a sporadic task without self-suspensions and parameters $C_1 = 2$ and $T_1=D_1=10$, and let $\tau_2$ denote a self-suspending task consisting of two segments with parameters  $C_2^1 = 1$,  $S_2^1 = 6$, $C_2^2=1$, and $ T_2=D_2=11$.
 Suppose that we use the rate-monotonic priority assignment, i.e., $\tau_1$ has higher priority than $\tau_2$.

### 2015-10-17 abstract.tex
edits

--- 2015-10-16 abstract.tex
@@
 \begin{abstract}
-  To be written...
+The \emph{period enforcer} algorithm for self-suspending real-time tasks is a technique for suppressing the ``back-to-back'' scheduling penalty associated with deferred execution.
+Originally proposed in 1991, the algorithm has attracted renewed interest in recent years.
+This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~to match the assumptions underlying the analysis of the period enforcer, a schedulability analysis of self-suspending tasks subject to period enforcement requires a task set  transformation that, with current techniques, is subject to exponential time complexity; and \textbf{(iii)} the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
 \end{abstract}

### 2015-10-17 conclusion.tex
edits

--- 2015-10-16 conclusion.tex
@@
 Finally, we have demonstrated that substantial difficulties arise if one attempts to combine suspension-based locks with period enforcement.
-These difficulties stem from the fact that period enforcement can increase contention, which increases self-suspension times, which then in turn feeds back into the period enforcer's minimum suspension lengths.
+These difficulties stem from the fact that period enforcement can increase contention or lock-holding times, which increases the lengths of self-suspension intervals, which then in turn feeds back into the period enforcer's minimum suspension lengths.
 As a consequence, period enforcement invalidates all existing blocking analyses.

### 2015-10-17 untitled.tex
edits

@@
 In this note, we revisit the period enforcer \cite{Raj:suspension1991} to carefully re-examine and explain its underlying assumptions and limitations, and to point out potential misconceptions.
-The main contributions are three observations that, to the best of our knowledge, have not been previously reported in the literature of real-time systems:
+The main contributions are three observations that, to the best of our knowledge, have not been previously reported in the literature on real-time systems:
 \begin{enumerate}
 	\item period enforcement can be a cause of deadline misses in self-suspending task sets that are otherwise schedulable (Section~\ref{sec:unschedulable}); 
-	\item with current techniques, schedulability analysis of the period enforcer algorithm requires a proper transformation, that is subject to exponential time complexity, of a self-suspending task set to satisfy the assumption of the period enforcer (Section~\ref{sec:convert}); and
+	\item to match the assumptions underlying the analysis of the period enforcer, a schedulability analysis of self-suspending tasks subject to period enforcement requires a task set  transformation that, with current techniques, is subject to exponential time complexity (Section~\ref{sec:convert}); and
 	\item the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed (Section~\ref{sec:locking}).

### 2015-10-17 abstract.tex
use non-breaking space after (iii)

--- 2015-10-17 abstract.tex
@@
 Originally proposed in 1991, the algorithm has attracted renewed interest in recent years.
-This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~with current techniques, schedulability analysis of the period enforcer algorithm requires a task set transformation that is subject to exponential time complexity; and \textbf{(iii)} the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
+This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~to match the assumptions underlying the analysis of the period enforcer, a schedulability analysis of self-suspending tasks subject to period enforcement requires a task set  transformation that, with current techniques, is subject to exponential time complexity; and \textbf{(iii)}~the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
 \end{abstract}

### 2015-10-18 abstract.tex
some proofreading fixes and improvements

--- 2015-10-17 abstract.tex
@@
 Originally proposed in 1991, the algorithm has attracted renewed interest in recent years.
-This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~with current techniques, schedulability analysis of the period enforcer algorithm requires a proper transformation, that is subject to exponential time complexity, of the task set to satisfy the assumptions; and \textbf{(iii)} the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
+This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior (compared to the base case without enforcement) as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~to match the assumptions underlying the analysis of the period enforcer, a schedulability analysis of self-suspending tasks subject to period enforcement requires a task set  transformation that, with current techniques, is subject to exponential time complexity; and \textbf{(iii)}~the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
 \end{abstract}

### 2015-10-18 locking.tex
some proofreading fixes and improvements

@@
 At time $3$, $\tau_1$ releases the shared resource and $\tau_2$ consequently resumes (i.e., $a^2_{2,1} = 3$).
-According to the period enforcer rules~\cite{Raj:suspension1991}, the second segment is immediately eligible because, according to Equation~\ref{eq:ET-def},
+According to the period enforcer rules~\cite{Raj:suspension1991}, the second segment is immediately eligible because, according to Equation~\ref{eq:ET-def} (in Section~\ref{sec:unschedulable}),
 \begin{align*}
@@
 At time $9$, the second job of $\tau_1$, released at time $8$, can thus lock the shared resource without contention.
-Consequently, when $\tau_2$'s request for the shared resource takes effect at time $10$, the resource is no longer available and $\tau_2$ must wait until time $a^2_{2,2} = 11$ before it can proceed in its execution.
+Consequently, when $\tau_2$'s request for the shared resource takes effect at time $10$, the resource is no longer available and $\tau_2$ must wait until time $a^2_{2,2} = 11$ before it can proceed to execute.
 We thus have

### 2015-10-18 untitled.tex
some proofreading fixes and improvements

@@
 The first segment of the second job arrives at time $a^1_{2,2} = 11$, is interfered with for one time unit, and suspends at time $13$.
-The segment second of the second job hence resumes only at time $a^2_{2,2} = 19$.
+The  second segment of the second job hence resumes only at time $a^2_{2,2} = 19$.
 Thus, we have
@@
 One may consider to enrich the period enforcer with the following power: when the processor becomes idle, a task immediately becomes eligible to execute regardless of its eligibility time.
-However, even with this exension, the above example remains valid by introducing one additional lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from time $3$ to time $9$ and time $13$ to time $20$) and $T_3=D_3=100$.
+However, even with this extension, the above example remains valid by introducing one additional lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from time $3$ to time $9$ and time $13$ to time $20$) and $T_3=D_3=100$.
 With task $\tau_3$, the processor is always busy from time $0$ to time $23$ and consequently $\tau_2$ still misses its deadline at time $22$.

### 2015-12-15 transformation.tex
proofreading fixes and tweaks
Minor fixes addressing Jim's comments

@@
 However, in general,  \emph{precisely} converting a self-suspending task system into a corresponding single-segment  deferrable task set is not an easy problem.
-We can demonstrate the inherent difficulty by looking at a special case.
+We demonstrate the inherent difficulty by focusing on a special case.
 
 Suppose that the system has $k-1$ ordinary sporadic tasks and only one segmented self-suspending task $\tau_k$.
-Converting a computation segment into a deferrable task requires to derive the \emph{worst-case resume time of a computation segment}, denoted as $R_k^j$ for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
+Converting a computation segment into a deferrable task requires deriving the \emph{worst-case resume time of a computation segment}, denoted as $R_k^j$ for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
 Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$.
@@
 
-In the context of the period enforcer, we consequently observe that the only existing solution to derive the \emph{precise} bound $W_k^{j}$ (and hence $R_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
+In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $R_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
 Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.

### 2015-12-15 untitled.tex
Minor fixes addressing Jim's comments

@@
 
-To avoid such detrimental effects,  Rajkumar \cite{Raj:suspension1991} proposed the \emph{period enforcer} algorithm~\cite{Raj:suspension1991},  a technique to control (or shape) the processor demand of self-suspending tasks.
+To avoid such detrimental effects,  Rajkumar \cite{Raj:suspension1991} proposed the \emph{period enforcer} algorithm,  a technique to control (or shape) the processor demand of self-suspending tasks under \emph{fixed-priority} scheduling.
 In a nutshell, the period enforcer algorithm artificially increases the length of certain suspensions whenever a task's activation pattern carries the risk of inducing undue interference in lower-priority tasks.
@@
   \input{../figures/example-original/example-body.tex}
-  \caption{An illustrative example of the original self-suspending task set (without period enforcement) assuming periodic job arrivals.}
+  \caption{An illustrative example of the original self-suspending task set (without period enforcement) assuming periodic job arrivals on a uniprocessor.
+Task $\tau_1$ has higher priority than task $\tau_2$.}
   \label{fig:example-original}
@@
   \caption{An illustrative example demonstrating a deadline miss at time 22
-    under the period enforcer algorithm.}
+    under the period enforcer algorithm. At time~19, $\tau_2$ resumes, but it remains ineligible to execute until time~20 when $\tau_1$ is released.}
 \label{fig:example}  

### 2015-12-15 untitled.tex
more fixes to address Jim's comments
- mention FP and partitioning early on
- spell out small numbers

Also add acknowledgements.

@@
 
-To avoid such detrimental effects,  Rajkumar \cite{Raj:suspension1991} proposed the \emph{period enforcer} algorithm,  a technique to control (or shape) the processor demand of self-suspending tasks under \emph{fixed-priority} scheduling.
+To avoid such detrimental effects,  Rajkumar \cite{Raj:suspension1991} proposed the \emph{period enforcer} algorithm,  a technique to control (or shape) the processor demand of self-suspending tasks on uniprocessors and partitioned multiprocessors under preemptive fixed-priority scheduling.
 In a nutshell, the period enforcer algorithm artificially increases the length of certain suspensions whenever a task's activation pattern carries the risk of inducing undue interference in lower-priority tasks.
@@
 The dynamic self-suspension sporadic task model characterizes each
-task $\tau_i$ as a $4$-tuple $(C_i,S_i,T_i,D_i)$: 
+task $\tau_i$ as a four-tuple $(C_i,S_i,T_i,D_i)$: 
 $C_i$ denotes an upper bound on the total execution time of any job of $\tau_i$,
@@
 
-In contrast, the segmented sporadic task model extends the above $4$-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals.
+In contrast, the segmented sporadic task model extends the above four-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals.
 These intervals are represented as a tuple

### 2015-12-15 untitled.tex
restructure background, clarify definitions

@@
 
-We briefly introduce the needed background in Section~\ref{sec:prelim}, restate our contributions more precisely in Section~\ref{sec:questions}, and then explain the three above  observations in detail in Sections \ref{sec:unschedulable}--\ref{sec:locking} before concluding in Section \ref{sec:conclusion}.
+We briefly introduce the needed background in Section~\ref{sec:prelim}, restate our contributions more precisely in Section~\ref{sec:questions}, and then establish the three above  observations in detail in Sections \ref{sec:unschedulable}--\ref{sec:locking} before concluding in Section \ref{sec:conclusion}.
 
-\subsection{Preliminaries}
+\section{Preliminaries}
 \label{sec:prelim}
+
+The period enforcer algorithm applies to self-suspending tasks on uniprocessors under fixed-priority scheduling, and hence by extension also to multiprocessors under partitioned fixed-priority scheduling (where tasks are statically assigned to processors and each processor is scheduled as a uniprocessor).
+In this section, we review the underlying task model (Section \ref{sec:taskmodel}), introduce the period enforcer algorithm (Section \ref{sec:pe}), summarize its analysis (Section \ref{sec:classic-analysis}), and finally restate our observations (Section \ref{sec:questions}).
+
+\subsection{Self-Suspending Tasks}
+\label{sec:taskmodel}
 
@@
 $(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, which is composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals.
-For the simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment.
+For simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment.
 The arguments can be easily extended to handle tasks that start with a self-suspension.
@@
 The period enforcer algorithm and its analysis fundamentally applies (only) to the segmented model.
+
+We say that a segment \emph{arrives} when it becomes available for execution.
+The first computation segment arrives immediately when the job is released; the second computation segment (if any) arrives when the job resumes from its first self-suspension, \textit{etc.} For simplicity, we assume that tasks are indexed in order of decreasing priority (i.e., $\tau_1$ is the highest-priority task).
+A \emph{level-$i$ busy interval} is a maximal interval during which  the processor executes only segments of tasks with priority $i$ or higher.
+
+\subsection{The Period Enforcer Algorithm}
+\label{sec:pe}
+
+The key idea underlying the period enforcer algorithm is to artificially delay the execution of computation segments if a job resumes ``too soon.'' To this end,  the period enforcer algorithm determines for each computation segment an \emph{eligibility time}.
+If a segment resumes  before its eligibility time, the execution of the segment is delayed until the eligibility time is reached.
+
+A segment's eligibility time is determined according to the following rule.
+Let $ET_{i,j}^k$ denote the eligibility time of the $k$\xth computation segment of the $j$\xth job of task $\tau_i$.
+Further, let $a^k_{i,j}$ denote the segment's arrival time.
+Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$ (i.e., the processor executes only $\tau_i$ or higher-priority tasks throughout the interval $[\mathit{busy}(\tau_i, t'), t']$).
+According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
+\begin{align}\label{eq:ET-def}
+	ET_{i,j}^k & = \max\left(ET_{i,j-1}^k + T_i,\ \mathit{busy}(\tau_i, a^k_{i,j})\right),
+\end{align}
+where $ET_{i,0}^k = -T_i$.
+
+% Assuming this definition,
+
+\subsection{Classic Analysis of the Period Enforcer Algorithm}
+\label{sec:classic-analysis}
 
@@
 %
-Finally, we make a third observation concerning the use of the period enforcer in conjunction with suspension-based multiprocessor locking protocols (such as the MPCP~\cite{LNR:09,Ra:90} or the FMLP~\cite{BLBA:07,BA:08}): while it is certainly tempting to apply period enforcement with the intention of avoiding the negative effects of deferred execution due to lock contention, it is actually unsafe to do so.
+Finally, we make a third observation concerning the use of the period enforcer in conjunction with suspension-based multiprocessor locking protocols for partitioned fixed-priority scheduling (such as the MPCP~\cite{LNR:09,Ra:90} or the FMLP~\cite{BLBA:07,BA:08}): while it is certainly tempting to apply period enforcement with the intention of avoiding the negative effects of deferred execution due to lock contention, it is actually unsafe to do so.
 \begin{enumerate}
@@
 
-Next, let us consider the same task set under control of the period enforcer algorithm.
-The period enforcer algorithm determines for each segment (i.e., for each converted single-segment deferrable task) an \emph{eligibility time}.
-If a segment resumes (i.e., a job of a single-segment deferrable task arrives) before its eligibility time, the execution of the segment (i.e., the single-segment job) is delayed until the eligibility time is reached.
-
-A segment's eligibility time is determined according to the following rule.
-Let $ET_{i,j}^k$ denote the eligibility time of the $k$\xth segment of the $j$\xth job of task $\tau_i$.
-Further, let $a^k_{i,j}$ denote the segment's arrival time.\footnote{A segment \emph{arrives} when it becomes available for execution.
-The first segment arrives immediately when the job is released; the second segment arrives when the job resumes from its self-suspension.} Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$.
-According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
-\begin{align}\label{eq:ET-def}
-	ET_{i,j}^k & = \max\left(ET_{i,j-1}^k + T_i,\ \mathit{busy}(\tau_i, a^k_{i,j})\right),
-\end{align}
-where $ET_{i,0}^k = -T_i$.
-
-Assuming this definition, Figure \ref{fig:example} shows the resulting schedule for a periodic release pattern.
+Next, let us consider the same task set under control of the period enforcer algorithm, as defined in Section \ref{sec:pe}.
+Figure \ref{fig:example} shows the resulting schedule for a periodic release pattern.
 The first job of task $\tau_2$ (which arrives at time $a^1_{2,1} =  0$) is executed as if there is no period enforcement since the definition $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$ ensures that both segments are immediately eligible.

### 2015-12-15 untitled.tex
Add example illustrating the period enforcer

@@
 
-The period enforcer algorithm applies to self-suspending tasks on uniprocessors under fixed-priority scheduling, and hence by extension also to multiprocessors under partitioned fixed-priority scheduling (where tasks are statically assigned to processors and each processor is scheduled as a uniprocessor).
+The period enforcer algorithm~\cite{Raj:suspension1991} applies to self-suspending tasks on uniprocessors under fixed-priority scheduling, and hence by extension also to multiprocessors under partitioned fixed-priority scheduling (where tasks are statically assigned to processors and each processor is scheduled as a uniprocessor).
 In this section, we review the underlying task model (Section \ref{sec:taskmodel}), introduce the period enforcer algorithm (Section \ref{sec:pe}), summarize its analysis (Section \ref{sec:classic-analysis}), and finally restate our observations (Section \ref{sec:questions}).
@@
 
+Recall that the scheduling penalty associated with self-suspensions is maximized when a task defers the completion of one job just until the release of the next job.
+For example, this effect is illustrated in Figure \ref{fig:not-ok-without-period-enforcement}, which shows a case in which the self-suspension of the higher-priority task $\tau_2$  results in a deadline miss of the lower-priority task $\tau_3$.
+The root cause is increased interference due to the ``back-to-back'' execution effect~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}, where two jobs of $\tau_2$ execute in close succession (i.e, separated by less than a period) because the second job self-suspended for a (much) shorter duration than the first job.
+That is, $\tau_3$ misses its deadline because $\tau_2$ resumed ``too soon.''
+
+\ifpaper
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=1]{../figures/not-ok-without-period-enforcer/not-ok.pdf}
+  \caption{Example schedule (\emph{without} period enforcement) of three tasks $\tau_1 = $, $\tau_2$, and $\tau_3$ with periods $T_1 = T_2 = T_3 = 10$.
+Tasks $\tau_1$ and $\tau_3$ consist of a single computation segment ($C_1^1 = C_3^1 = 3$); task $\tau_2$ consists of two computation and one suspension segment ($C_2^1 = 1$, $S_2^1 = 4$, $C_2^2 = 2$).
+Jobs of tasks $\tau_1$ and $\tau_3$ are released just as $\tau_2$ resumes from its self-suspension at time~$5$.
+Without period enforcement, task $\tau_3$ misses a deadline at time~$15$ because the second job of  task $\tau_2$ suspends only briefly (for one time unit rather than four).}
+  \label{fig:not-ok-without-period-enforcement}
+  \end{figure}
+\fi
+
 The key idea underlying the period enforcer algorithm is to artificially delay the execution of computation segments if a job resumes ``too soon.'' To this end,  the period enforcer algorithm determines for each computation segment an \emph{eligibility time}.
@@
 
-% Assuming this definition,
+Figure \ref{fig:ok-with-period-enforcement} illustrates how this definition of eligibility times restores the schedulability of the task set depicted in Figure~\ref{fig:not-ok-without-period-enforcement}.
+Consider the eligibility times of the second segment of task $\tau_2$.
+
+By definition, we have $ET_{2,0}^2 = -T_2 = -10$. At time~5, when the second computation segment of the first job resumes ($a_{2,1}^2 = 5$), we thus have
+\begin{align*}
+	ET_{2,1}^2 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, a_{2,1}^2\right) ) = \max(0, 5) = 5
+\end{align*}
+since the arrival of $\tau_2$'s second segment (and the release of $\tau_1$) starts a new level-2 busy interval at time $a_{2,1}^2 = 5$.
+The second second segment of $\tau_2$'s first job is hence immediately eligible to execute; however, due to the presence of a pending higher-priority job $\tau_2$ is not actually scheduled until time~8 (just as without period enforcement as depicted in Figure \ref{fig:not-ok-without-period-enforcement}).
+
+The second segment of the second job of $\tau_2$ arrives at time $a_{2,2}^2 = 12$. In this case, the segment is \emph{not} immediately eligible to execute since
+\begin{align*}
+	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a_{2,2}^2\right) ) = \max(5 + 10, 12) = 15.
+\end{align*}
+Hence, the execution of $\tau_2$'s second computation segment does not start until time $ET_{2,2}^2 = 15$, which gives $\tau_3$ sufficient time to finish before its deadline at time~$20$.
+
+\ifpaper
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=1]{../figures/ok-with-period-enforcer/ok.pdf}
+  \caption{Example schedule \emph{with} period enforcement assuming the same scenario as depicted in Figure~\ref{fig:not-ok-without-period-enforcement}.
+With period enforcement, task $\tau_3$ does not miss a deadline because task $\tau_2$'s second computation segment is delayed until it no longer imposes undue interference (i.e., it is prevented from resuming ``too soon'').}
+  \label{fig:ok-with-period-enforcement}
+  \end{figure}
+\fi
+
+The examples in Figures \ref{fig:not-ok-without-period-enforcement} and \ref{fig:ok-with-period-enforcement} suggest an intuition for the benefits provided by period enforcement: computation segments of a self-suspending task $\tau_i$ are forced to execute at least $T_i$ time units apart (hence the name), which ensures that it causes no more interference than a regular (non-self-suspending) sporadic task.
+Next, we revisit the original analysis of this technique.
 

### 2015-12-15 untitled.tex
Clarify questions answered

@@
 For simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment.
-The arguments can be easily extended to handle tasks that start with a self-suspension.
+Any suspension before the first computation segment is equivalent to \emph{release jitter}~\cite{ABRTW:93}.
 
@@
 \end{align*}
-Hence, the execution of $\tau_2$'s second computation segment does not start until time $ET_{2,2}^2 = 15$, which gives $\tau_3$ sufficient time to finish before its deadline at time~$20$.
+Hence, the execution of $\tau_2$'s second computation segment does not start until time $ET_{2,2}^2 = 15$, which gives $\tau_3$ sufficient time to finish before its deadline at time~$15$.
 
@@
 %
-In other words, the transformation can be understood as splitting each self-suspending task into a matching number of non-self-suspending sporadic tasks subject to \emph{release jitter}, which can be easily analyzed with classic fixed-priority response-time analysis~\cite{ABRTW:93}.
+In other words, the transformation can be understood as splitting each self-suspending task into a matching number of non-self-suspending sporadic tasks subject to release jitter, which can be easily analyzed with classic fixed-priority response-time analysis~\cite{ABRTW:93}.
 
-It is well known that uncontrolled deferred execution (i.e, release jitter) can impose a scheduling penalty because of the potential for ``back-to-back'' execution~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}.
-That is, if a job of a deferrable task that maximally defers its execution is directly followed by a job that executes immediately without deferring its execution, then lower-priority tasks may suffer increased interference.
+It is well known that uncontrolled deferred execution (i.e, release jitter) can impose  increased interference on lower-priority tasks because of the potential for ``back-to-back'' execution~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}, as illustrated in Figure \ref{fig:not-ok-without-period-enforcement}.
 
@@
 
-Theorem 5 in \cite{Raj:suspension1991} provides a very positive result for the analysis of deferrable task sets subject to period enforcement: if the corresponding transformed task set can be shown to be schedulable under fixed-priority scheduling using \emph{any} applicable analysis, then the period enforcer algorithm also yields a correct schedule.
+Theorem 5 (in \cite{Raj:suspension1991}) is a strong result that seemingly enables a powerful analysis approach: if the corresponding transformed set of non-self-suspending tasks (subject to release jitter, without period enforcement) can be shown to be schedulable under fixed-priority scheduling using \emph{any} applicable analysis (e.g.,~\cite{ABRTW:93}), then the period enforcer algorithm also yields a correct schedule.
 
-However, we note that Theorem 5 in \cite{Raj:suspension1991} applies only to the \emph{transformed} task set: recall that, in the original analysis~\cite{Raj:suspension1991}, deferrable tasks are assumed to defer their entire execution time either completely or not at all (but not parts of it).
-Therefore, if we would like to apply the period enforcer to segmented self-suspending task sets, we first have to answer the following question: ``\emph{Given a set of sporadic segmented self-suspending tasks, what is the corresponding set of (single-segment) deferrable tasks?}'' That is, how do we convert given self-suspension segments into equivalent bounds on release jitter such that we may apply Theorem 5 to conclude that the system remains schedulable despite period enforcement?
+However, recall that, in the original analysis~\cite{Raj:suspension1991}, deferrable tasks are assumed to defer their  execution either completely or not at all (but not parts of it).
+It is hence important to realize that Theorem 5 in \cite{Raj:suspension1991} applies only to the \emph{transformed} task set---Theorem 5 in \cite{Raj:suspension1991} does \emph{not} apply to the \emph{original} set of segmented self-suspending tasks.
+This leads to the first question: \emph{Does schedulability of the transformed set of non-self-suspending tasks (subject to release jitter, without period enforcement) also imply schedulability of the original set of segmented self-suspending tasks (under period enforcement)?} In Section \ref{sec:unschedulable}, we answer this question in the negative.
 
-In this paper, which is motivated by the fact that the original proposal \cite{Raj:suspension1991} does not provide an answer to this central question, 
-we make two pertinent observations.
 \begin{enumerate}
@@
 This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, and not in the context of the original segmented self-suspending task set.
+\end{enumerate}
 
-	\item Deriving a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem.
-Recent findings by Nelissen et al.
-\cite{ecrts15nelissen} can be applied, but their method takes exponential time.
+Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the following question: \emph{Given a set of sporadic segmented self-suspending tasks, how do we obtain a corresponding set of (single-segment) deferrable tasks?} That is, the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert self-suspension segments into equivalent bounds on release jitter, but it remains unclear \emph{how} this central step should be accomplished.
+In Section \ref{sec:convert}, we make a pertinent observation.
+
+\begin{enumerate}
+\setcounter{enumi}{1}
+	\item How to derive a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem.
+Recent findings by Nelissen et al.~\cite{ecrts15nelissen} can be applied, but their method takes exponential time.
 \end{enumerate}
-%
-Finally, we make a third observation concerning the use of the period enforcer in conjunction with suspension-based multiprocessor locking protocols for partitioned fixed-priority scheduling (such as the MPCP~\cite{LNR:09,Ra:90} or the FMLP~\cite{BLBA:07,BA:08}): while it is certainly tempting to apply period enforcement with the intention of avoiding the negative effects of deferred execution due to lock contention, it is actually unsafe to do so.
+
+Finally, we make a third observation concerning the use of the period enforcer in conjunction with suspension-based multiprocessor locking protocols for partitioned fixed-priority scheduling (such as the MPCP~\cite{LNR:09,Ra:90} or the FMLP~\cite{BLBA:07,BA:08}).
+While it is certainly tempting to apply period enforcement with the intention of avoiding the negative effects of deferred execution due to lock contention (as previously suggested elsewhere~\cite{Raj:91,Lak:11,LNR:09}), we ask: \emph{Does existing blocking analysis remain safe when combined with the period enforcer algorithm?} In Section \ref{sec:locking}, we show that this is not the case.
 \begin{enumerate}

### 2015-12-15 untitled.tex
Replace figures

@@
   \includegraphics[scale=1]{../figures/not-ok-without-period-enforcer/not-ok.pdf}
-  \caption{Example schedule (\emph{without} period enforcement) of three tasks $\tau_1 = $, $\tau_2$, and $\tau_3$ with periods $T_1 = T_2 = T_3 = 10$.
+  \caption{Example uniprocessor schedule (\emph{without} period enforcement) of three tasks $\tau_1 = $, $\tau_2$, and $\tau_3$ with periods $T_1 = T_2 = T_3 = 10$.
 Tasks $\tau_1$ and $\tau_3$ consist of a single computation segment ($C_1^1 = C_3^1 = 3$); task $\tau_2$ consists of two computation and one suspension segment ($C_2^1 = 1$, $S_2^1 = 4$, $C_2^2 = 2$).
@@
   \includegraphics[scale=1]{../figures/ok-with-period-enforcer/ok.pdf}
-  \caption{Example schedule \emph{with} period enforcement assuming the same scenario as depicted in Figure~\ref{fig:not-ok-without-period-enforcement}.
+  \caption{Example uniprocessor schedule \emph{with} period enforcement assuming the same scenario as depicted in Figure~\ref{fig:not-ok-without-period-enforcement}.
 With period enforcement, task $\tau_3$ does not miss a deadline because task $\tau_2$'s second computation segment is delayed until it no longer imposes undue interference (i.e., it is prevented from resuming ``too soon'').}
@@
 \setcounter{enumi}{1}
-	\item How to derive a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem.
+	\item Deriving a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem.
 Recent findings by Nelissen et al.~\cite{ecrts15nelissen} can be applied, but their method takes exponential time.
@@
   \centering
-  \input{../figures/example-original/example-body.tex}
+  \includegraphics[scale=1]{../figures/example-original/example-original.pdf}
   \caption{An illustrative example of the original self-suspending task set (without period enforcement) assuming periodic job arrivals on a uniprocessor.
@@
   \centering
-  \input{../figures/example/example-body.tex}
+  \includegraphics[scale=1]{../figures/example/example.pdf}
   \caption{An illustrative example demonstrating a deadline miss at time 22

### 2015-12-15 untitled.tex
proofreading fixes

@@
 To date, the real-time literature on self-suspensions has focused on two task models: the \emph{dynamic} and the \emph{segmented} (or \emph{multi-segment}) self-suspension model. 
-The dynamic self-suspension sporadic task model characterizes each
+The dynamic self-suspending sporadic task model characterizes each
 task $\tau_i$ as a four-tuple $(C_i,S_i,T_i,D_i)$: 
@@
 
-In contrast, the segmented sporadic task model extends the above four-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals.
+In contrast, the segmented self-suspending sporadic task model extends the above four-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals.
 These intervals are represented as a tuple
@@
 The advantage of the segmented model is that is allows for more accurate analysis.
-The period enforcer algorithm and its analysis fundamentally applies (only) to the segmented model.
+The period enforcer algorithm and its analysis applies (only) to the segmented model.
 
@@
 since the arrival of $\tau_2$'s second segment (and the release of $\tau_1$) starts a new level-2 busy interval at time $a_{2,1}^2 = 5$.
-The second second segment of $\tau_2$'s first job is hence immediately eligible to execute; however, due to the presence of a pending higher-priority job $\tau_2$ is not actually scheduled until time~8 (just as without period enforcement as depicted in Figure \ref{fig:not-ok-without-period-enforcement}).
+The second segment of $\tau_2$'s first job is hence immediately eligible to execute; however, due to the presence of a pending higher-priority job, $\tau_2$ is not actually scheduled until time~8 (just as without period enforcement as depicted in Figure \ref{fig:not-ok-without-period-enforcement}).
 

### 2016-02-29 untitled.tex
Reviewer A and Reviewer B (first comment)

@@
 
-The advantage of the dynamic model is that it is more flexible since it does not impose any assumptions on control flow.
-The advantage of the segmented model is that is allows for more accurate analysis.
+A special case of the segmented self-suspending sporadic task model is the deferrable sporadic task model, in which each task $\tau_i$ starts with a self-suspension interval followed by one computation segment.
+The deferrable sporadic task model is the central notation in Rajkumar's analysis~\cite{Raj:suspension1991}.
+We will explicitly explain the link in Section~\ref{sec:classic-analysis}.
+
+The advantage of the dynamic model is that it is more flexible since it does not impose any assumptions on the task control flow.
+The advantage of the segmented model is that it allows for more accurate analysis.
 The period enforcer algorithm and its analysis applies (only) to the segmented model.
@@
 
-The central notation in Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{deferrable task}, which matches our notion of multi-segment tasks.  Specifically, Rajkumar states that:
+The central notation in Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{deferrable task}, which matches our notion of segmented tasks.  Specifically, Rajkumar states that:
 \begin{quote}
@@
 
-One may consider to enrich the period enforcer with the following power: when the processor becomes idle, a task immediately becomes eligible to execute regardless of its eligibility time.
+One may consider to enrich the period enforcer with the following property: when the processor becomes idle, a task immediately becomes eligible to execute regardless of its eligibility time.
 However, even with this extension, the above example remains valid by introducing one additional lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from time $3$ to time $9$ and time $13$ to time $20$) and $T_3=D_3=100$.
@@
 
-Furthermore, the example also demonstrates that the conversion to single-segment deferrable tasks does incur a loss of generality since it introduces pessimism: if we convert the two computation segments of task $\tau_2$ into two deferrable tasks $\tau_2^1$ and $\tau_2^2$, where  task $\tau_2^1$ never defers its execution and task $\tau_2^2$ defers its execution by at most \emph{$9$} time units,  then it is immediately obvious that the resulting single-segment deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable under a rate-monotonic priority assignment since we can time the release of a job of $\tau_1$ to coincide with the arrival of a job of $\tau_2^2$ after it has maximally deferred its execution, which clearly results in a deadline miss.
+We can now examine why task $\tau_2$ misses the deadline by the period enforcer algorithm.
+
+Clearly, the suspension time of task $\tau_2^1$ is $0$.
+The suspension time of task $\tau_2^2$ is $C_2^1+C_1+S_2^1=9$.
+In the worst case response time analysis, we can easily see that the worst-case response time of task $\tau_2^2$ is $2+1$ after the second computation segment arrives.
+
+Furthermore, the example also demonstrates that the conversion to single-segment deferrable tasks does incur a loss of generality since it introduces pessimism.
+Recall that the analysis in Theorem 5 in~\cite{Raj:suspension1991} was only compatible with the deferrable task model.
+For the above example, we need to convert the segmented-suspending sporadic task $\tau_2$ into two deferrable tasks, called $\tau_2^1$ and $\tau_2^2$.
+By converting the two computation segments of task $\tau_2$ into two deferrable tasks $\tau_2^1$ and $\tau_2^2$, we can conclude that  task $\tau_2^1$ never defers its execution and task $\tau_2^2$ defers its execution by at most \emph{$9$} time units.
+Then the resulting single-segment deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable under the given fixed-priority scheduling since we can time the release of a job of $\tau_1$ to coincide with the arrival of a job of $\tau_2^2$ after it has maximally deferred its execution.
+We can easily see that the worst-case response time of the computation segment represented by task $\tau_2^2$ is $3$ time units after the second computation segment of task $\tau_2$ arrives.
+Together with the maximum deferred time $9$ time units, we know that the corresponding deferrable task set in fact has deadline misses since $9+3=12 > 11$.
+Therefore, the period enforcer algorithm may also cause a deadline miss.
 

### 2016-02-29 transformation.tex
Reviewer B remark 2

@@
 
-The example in Section \ref{sec:unschedulable} can be easily converted to a corresponding single-segment deferrable task set.
-However, in general,  \emph{precisely} converting a self-suspending task system into a corresponding single-segment  deferrable task set is not an easy problem.
-We demonstrate the inherent difficulty by focusing on a special case.
+The example in Section~\ref{sec:unschedulable} can be easily converted to a corresponding deferrable task set, as explained at the end of Section~\ref{sec:unschedulable}.
+Due to Theorem 5 in~\cite{Raj:suspension1991}, the feasibility of the schedule by the period enforcer algorithm depends on whether the corresponding deferrable task set can be feasibly scheduled.
+Therefore, before applying the period enforcer algorithm to handle segmented self-suspending sporadic tasks, we need to first derive the corresponding deferrable task set as precisely as possible.
 
-Suppose that the system has $k-1$ ordinary sporadic tasks and only one segmented self-suspending task $\tau_k$.
-Converting a computation segment into a deferrable task requires deriving the \emph{worst-case resume time of a computation segment}, denoted as $R_k^j$ for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
+ However, in general, such a conversion is not an easy problem.
+We demonstrate the inherent difficulty by focusing on a special case and by applying the recent result provided by Nelissen et al.
+\cite{ecrts15nelissen}, which analyzed the exact worst-case response time for segmented self-suspending sporadic tasks.
+Suppose that the system has $k-1$ ordinary sporadic tasks and only one segmented self-suspending task $\tau_k$ with $D_k = T_k$.
+Suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.
+Converting a computation segment into a deferrable task requires deriving the worst-case deferrable time, denoted as $R_k^j$, for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
+Formally, if a job of task $\tau_k$ arrives at time $t$, it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than $t+R_k^j$.
 Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$.
-It is not difficult to see that $R_k^1=0$ and $R_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k-1$.
-We therefore need to derive the worst-case response times of the computation segments of task $\tau_k$.
+Therefore,  if we can derive the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ in this special case, we can clearly conclude that $R_k^1=0$ and $R_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k$.
 
 Based on these considerations, it appears that, at least for the simple example, the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems.
-However, it has been recently shown by Nelissen et al.\ \cite{ecrts15nelissen} that calculating the worst-case response time in the above ``simple'' case is already a very challenging problem.
-In particular, Nelissen et al.\ \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.
+Deriving the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not an easy problem.
+The method recently provided by Nelissen et al.
+\cite{ecrts15nelissen} can be used for this specific case to derive the exact $W_k^j$ if we assume that task $\tau_k$ has only the first $j$ computation segments and $j-1$ self-suspension intervals.
+However, Nelissen et al.
+\cite{ecrts15nelissen} also showed that calculating the worst-case response time in the above ``simple'' case is already a very challenging problem, in which calculating $W_k^j$ would need exponential time complexity if $j \geq 2$.
+In particular, Nelissen et al.
+\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.
 

### 2016-02-29 untitled.tex
Comments from Reviewer B.

@@
 
-We can now examine why task $\tau_2$ misses the deadline by the period enforcer algorithm.
-
-Clearly, the suspension time of task $\tau_2^1$ is $0$.
-The suspension time of task $\tau_2^2$ is $C_2^1+C_1+S_2^1=9$.
-In the worst case response time analysis, we can easily see that the worst-case response time of task $\tau_2^2$ is $2+1$ after the second computation segment arrives.
-
 Furthermore, the example also demonstrates that the conversion to single-segment deferrable tasks does incur a loss of generality since it introduces pessimism.

### 2016-03-01 transformation.tex
Comments from Reviewer B.
reorganize Section 4

@@
-\section{Deriving a Corresponding Deferrable Task Set}
+\section{Deriving a Corresponding Deferrable Task Set and Schedulability Test}
 \label{sec:convert}
@@
 Due to Theorem 5 in~\cite{Raj:suspension1991}, the feasibility of the schedule by the period enforcer algorithm depends on whether the corresponding deferrable task set can be feasibly scheduled.
-Therefore, before applying the period enforcer algorithm to handle segmented self-suspending sporadic tasks, we need to first derive the corresponding deferrable task set as precisely as possible.
+Therefore, before applying the period enforcer algorithm to handle segmented self-suspending sporadic tasks, we need to first derive the corresponding deferrable task set and perform the schedulability test for the deferrable task set.
+To reduce the pessimism of the transformation to the corresponding deferrable task set, it would be the best to have a precise transformation and an exact schedulability test.
+Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.
 
- However, in general, such a conversion is not an easy problem.
+With the above discussions, we need to answer two questions before adopting the period enforcer algorithm.
+First, how can we derive the corresponding deferrable task set efficiently?
+Second, how should we perform the schedulability test on the corresponding deferrable task set correctly?
+
+\subsection{Task Set Transformation}
+\label{sec:transformation-exponential}
+
+The transformation from a segmented self-suspending task set to a corresponding deferrable task set was not discussed in \cite{Raj:suspension1991}.
+However, in our opinion, in general, such a transformation is not an easy problem.
 We demonstrate the inherent difficulty by focusing on a special case and by applying the recent result provided by Nelissen et al.
-\cite{ecrts15nelissen}, which analyzed the exact worst-case response time for segmented self-suspending sporadic tasks.
-Suppose that the system has $k-1$ ordinary sporadic tasks and only one segmented self-suspending task $\tau_k$ with $D_k = T_k$.
+\cite{ecrts15nelissen}, which analyzed the exact worst-case response time for segmented self-suspending sporadic tasks with exponential time complexity.
+The worst-case response time analysis in \cite{ecrts15nelissen} is exact under the following conditions: 1) the task set has only one segmented self-suspending task, 2) the self-suspending task is the lowest-priority task, 3) the scheduling policy is preemptive fixed-priority scheduling, and 4) $D_i \leq T_i$ for all task $\tau_i$ in the task system.
+
+Suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$.
+The implicit deadline of a task is equal to its minimum inter-arrival time.
 Suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.
-Converting a computation segment into a deferrable task requires deriving the worst-case deferrable time, denoted as $R_k^j$, for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
-Formally, if a job of task $\tau_k$ arrives at time $t$, it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than $t+R_k^j$.
+Converting a computation segment into a deferrable task requires deriving the worst-case deferrable time (release jitter), denoted as $\rho_k^j$, for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
+Formally, if a job of task $\tau_k$ arrives at time $t$, it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than $t+\rho_k^j$.
 Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$.
-Therefore,  if we can derive the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ in this special case, we can clearly conclude that $R_k^1=0$ and $R_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k$.
+Therefore,  if we can derive the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ in this special case, we can clearly conclude that $\rho_k^1=0$ and $\rho_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k$.
 
@@
 
-In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $R_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
-Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.
+In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
+However, finding a safe upper bound of $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ can be done in pseudo-polynomial time \cite{PH:rtss98}, if over-approximations can be tolerated.
+
+\subsection{Schedulability Test for A Corresponding Deferrable Task Set}
+\label{sec:schedulability-test-deferrable}
+
+Here, we explain the potential schedulability test that can be used for a corresponding deferrable task set.
+It may be seemingly correct at beginning that we do not have to consider the interference from the other deferrable tasks created from one segmented self-suspending task.
+For example, in the special task set considered in Section~\ref{sec:transformation-exponential}, one may argue that it is not necessary to consider the interference from the other computation segments of task $\tau_k$ while analyzing the feasibility of the last computation segment.
+However, this statement is incorrect.
+
+Consider a task system consisting of $2$ tasks.
+Let $\tau_1$ denote a sporadic task without self-suspensions and parameters $C_1 = 2$ and $T_1=D_1=10$, and let $\tau_2$ denote a self-suspending task consisting of three segments with parameters  $C_2^1 = 1$,  $S_2^1 = 6$, $C_2^2=1$, $S_2^2=8$, $C_2^3=1$, and $ T_2=D_2=21$.
+That is, this task set is pretty similar to the task set used in Section~\ref{sec:unschedulable}.
+One additional computation
+segment is created to illustrate the potential concerns.
+This task set is schedulable by fixed-priority preemptive scheduling without any enforcement since the worst-case response time of the second computation segment is $10$ and the third computation segment will suffer from the interference by at most one job from task $\tau_1$.
+Moreover, by using the analysis in Section~\ref{sec:transformation-exponential}, we can conclude that
+\begin{equation*}
+  \rho_2^1 = 0, \rho_2^2 = 9, \rho_2^3 = 18.
+\end{equation*}
+If we do not consider the interference from the first two computation segments of task $\tau_2$, we can conclude that the last computation segment of task $\tau_2$ has worst-case deferrable time (release jitter) $18$ and the worst-case response time of a job of task $\tau_2$ is hence $18+C_1+C_2^3=21$.
+
+However, the above statement is not correct. Let us consider this task set under control of the period enforcer algorithm, as defined in Section \ref{sec:pe}.
+Figure \ref{fig:example-deferrable-incompatible} shows the resulting schedule for a periodic release pattern. This changes the schedule of the second job of task $\tau_2$. 
+The second job of task $\tau_2$ (which is released at time $21$) is affected by period enforcement.
+The first segment of the second job arrives at time $a^1_{2,2} = 21$, is interfered for one time unit, and suspends at time $23$.
+The  second segment of the second job hence resumes only at time $a^2_{2,2} = 29$.
+Thus, we have
+\begin{align*}
+	ET_{2,2}^1 & = \max\left(0 + 21,\ \mathit{busy}(\tau_2, 21)\right) = 21  \text{ and }
+\\
+	ET_{2,2}^2 & = \max\left(9 + 21,\ \mathit{busy}(\tau_2, 29\right) ) = 30.
+\end{align*}
+According to the rules of the period enforcer algorithm, the processor therefore remains idle at time $29$ because the segment is not eligible to execute until time $ET_{2,2}^2 = 30$.
+However, at time $30$, the fourth job of $\tau_1$ is released.
+As a result, the second job of $\tau_2$ suffers from additional interference and misses its deadline at time $42$.
+
+If we further allow task $\tau_1$ to be released sporadically, the fifth job of task $\tau_1$ can be released at time $41$ instead of time $40$ in Figure~\ref{fig:example-deferrable-incompatible}.
+For such a case, the response time of the second job of task $\tau_2$ is $23$.
+This means that the worst-case response time of the deferrable task created to represent the third computation segment of task $\tau_2$ must be at least $23-\rho_2^3=5$, which equals to $C_1 + C_2^1+C_2^2+C_2^3$.
+
+Therefore, for this specific example, we know that the schedulability test for the  corresponding deferrable task set has also to be pessimistic enough to consider the interferences from all the computation segments.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.4]{../figures/deferrable-incompatible.pdf}
+  \caption{}
+  \label{fig:example-deferrable-incompatible}
+\end{figure}
+
+\subsection{Discussions}
+\label{sec:discussions-deferrable}
+
+The transformation from a segmented self-suspending task set to a corresponding deferrable task set and the corresponding schedulability test were not discussed in \cite{Raj:suspension1991}.
+With the above discussions in Section~\ref{sec:transformation-exponential} and Section~\ref{sec:schedulability-test-deferrable}, we find that the transformation is in fact non-trivial and the schedulability test of the corresponding deferrable task set has also to be pessimistic.
+As shown in Section~\ref{sec:schedulability-test-deferrable}, even if we derive the precise deferrable time (release jitter) information, we may still have to account for the computation segments that have no \emph{direct} interference with a computation segment that is under analysis (e.g., the third computation segment of task $\tau_2$ in Section~\ref{sec:schedulability-test-deferrable}).
+
+Fundamentally, the design of the period enforcer algorithm relies on the schedulability of a deferrable task set.
+However, as we demonstrated here, there is a gap between the segmented self-suspending task set and the corresponding deferrable task set in the transformation and also in the schedulability test analysis.
+The authors in this paper do not find any simple way to provide a schedulability test without any information loss.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "LITES/LITES-Paper.tex"
+%%% End:

### 2016-03-01 untitled.tex
reorganize Section 4

@@
 We will explicitly explain the link in Section~\ref{sec:classic-analysis}.
+A deferrable sporadic task is equivalent to
+an ordinary sporadic task with \emph{release jitter}.
 
@@
 The advantage of the segmented model is that it allows for more accurate analysis.
-The period enforcer algorithm and its analysis applies (only) to the segmented model.
+The period enforcer algorithm and its analysis applies (only) to the segmented model, to be explained in Section~\ref{sec:classic-analysis}.
 
@@
 With deferred execution, a task $\tau_i$ can execute its $C_i$ units of execution in discrete amounts $C_i^1, C_i^2$, $\ldots$ with suspension in between $C_i^j$ and $C_i^{j+1}$.
-\cite[Section 3]{Raj:suspension1991}\footnote{The notation has been altered here for the sake of consistency.}
+\cite[Section 3]{Raj:suspension1991}\footnote{The notation has been altered here for the sake of consistency.
+}
 \end{quote}
+
+Note that it is not possible to convert a dynamic self-suspending task into the notion of deferred execution defined in \cite[Section 3]{Raj:suspension1991}, since the definition of a dynamic self-suspending task creates dynamic execution amounts for different jobs of a task.
+A simple example can explain why the period enforcer algorithm is not compatible with the dynamic self-suspending task model.
+Suppose that the system has only one task with execution time $C_1=1$, $S_1=1$, and $D_1=T_1=2$.
+The first job of task $\tau_1$ arrives at time $0$, suspends itself for one time unit, and then executes for one time unit.
+The second job of task $\tau_1$ arrives at time $2$, first executes for $0.5$ time unit, then suspends for $1$ time unit, and executes for $0.5$ time unit.
+With the period enforcer algorithm, the second job of task $\tau_1$ starts its execution at time $3$ and will clearly miss the deadline at time $4$.
+
 %
@@
 In contrast to the first job, the second job of task $\tau_2$ (which is released at time $11$) is affected by period enforcement.
-The first segment of the second job arrives at time $a^1_{2,2} = 11$, is interfered with for one time unit, and suspends at time $13$.
+The first segment of the second job arrives at time $a^1_{2,2} = 11$, is interfered for one time unit, and suspends at time $13$.
 The  second segment of the second job hence resumes only at time $a^2_{2,2} = 19$.
@@
 However, at time $20$, the third job of $\tau_1$ is released.
-As a result, the second job of $\tau_2$ incurs additional interference and misses its deadline at time $22$.
+As a result, the second job of $\tau_2$ suffers from additional interference and misses its deadline at time $22$.
 

### 2016-03-01 LITES/response.tex
add response letter

\documentclass[12pt]{article}
\usepackage{fullpage}

\usepackage{times}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
% \usepackage{cite}
\usepackage[sort]{natbib}
\usepackage{setspace}
\usepackage{multido}
\usepackage[justification=Centering]{subfig}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{shadows,patterns,shapes,arrows,decorations.pathmorphing,backgrounds,positioning,fit,plotmarks}

 \def\myendproof{{\ \vbox{\hrule\hbox{%
   \vrule height1.3ex\hskip0.8ex\vrule}\hrule }}\par}
 \newenvironment{proof}{\noindent{\bf Proof. }}{\myendproof}

\newcommand{\response}[1]{\textcolor{red}{\emph{Response: }} \textcolor{blue}{#1}\vspace{0.25in}}
\newcommand{\action}[1]{\textcolor{blue}{\emph{\bf Actions Taken: }} \emph{#1}\myendproof \vspace{0.25in}}

\begin{document}

{\bf Reviewer A}:

	The paper discusses the application of the period enforcer algorithm to
segmented self-suspending tasks in a unicore/partitioned-multicore settings.
In particular, the authors contrived multiple scenarios and task sets to
exemplify situations where (1) schedulability is achieved without using the
period enforcer but deadlines are missed if the enforcer is used, and (2)
the enforcer is not compatible with multicore resource sharing mechanisms in
the sense that it invalidates the analyses currently employed. As an extra
contribution, the authors also show that converting sporadic segmented
self-suspending tasks into single-segment deferrable tasks (i.e. tasks with
release jitters) incurs extra pessimism in the analysis (which means that
the original task set may be schedulable whereas its corresponding
deferrable tasks set is not), and doing the conversion in polynomial time
is an open problem.

The problem is relevant and interesting and I believe the paper fits very
well in the scope of Lites. The structure of the paper is good and the
presentation is excellent, both in terms of language, and general structure.
The paper is easy and pleasant to read and presents very pedagogic content.
I warmly recommend its publication.

{\bf Q:A.1}:    
I only have a few very minor comments:\\
- p2, line 60: assumptions on control flow $\rightarrow$ assumptions on
the task control flow\\
- p2, line 60: the segmented model is that is allows for $\rightarrow$ the
segmented model is that **it** allows for\\
- p4, line112: I would say segmented self-suspending tasks rather than
multi-segment tasks to remain consistent with the naming introduced in
the system model.\\
- p7, line 212:  to enrich the period enforcer with the following
power, consider replacing power with ``property'' or
``scheduling rule''.\\
- p7, line 221: ``it is immediately obvious that'' $\rightarrow$ I would simply
remove that part of the sentence since nothing is obvious with
self-suspending tasks and particularly for the problem addressed in this
paper.

\vskip 0.1in\noindent
\response{Thanks for the suggestions and comments. We have rephrased those parts accordingly.}

{\bf Reviewer B}:

	This work makes the following observations
(i) period enforcement is not strictly superior to scheduling without
enforcement
(ii) period enforcement is incompatible with existing analysis of
suspension-based locking protocols.

The observations described above are valid. However, there are some unstated
assumptions about the period enforcer in how the technique is being applied,
which should be clarified in this paper.

{\bf Q:B.1}:    
First, the period enforcer analysis only applies to the case where a task
defers its entire execution time but not parts of it. Tasks that suspend
after executing for a while should be considered as two or more tasks with
their own worst-case execution time. It is known that if tasks are broken up
into multiple segments with individual relative deadlines adding up to the
original deadline, the schedulability is no longer guaranteed. For example,
a task set (C,T,D), with tasks (4, 5, 5) and (2, 10, 10) is schedulable,
whereas breaking up the (2, 10, 10) into (1.5, 10, 7.5) and (0.5, 10, 2.5)
is not schedulable.

Section 3 should address the above clarification and the task $\tau_2$ should
be statically broken up before applying the period enforcer. It is true that
transforming a task set by breaking up tasks into sub tasks with sub
deadlines can lead to unschedulability on an otherwise schedulable task set.

\response{
Reviewer B was partially correct. Since the period enforcer analysis only applies to the case where a task defers its entire execution time but not parts of it, we should create the corresponding deferrable task set. However, it is not necessary that we need to create those deferrable task sets before using the period enforcer algorithm. Since the computation segments of a segmented self-suspending task are independent from each other, we can simply use the period enforcer algorithm without knowing the corresponding deferrable task set. This is actually how we demonstrate the examples in Section 2 and Section 3. However, there is a risk that the resulting schedule may not be feasible. Therefore, one would need to verify the feasibility of the corresponding deferrable task set before applying the period enforcer algorithm. Such a step was assumed in [21], and we explain the difficulty in this paper. The period enforcement in the example in Section 3 is only used to enforce the periodic behavior of the 2nd computation segment of task $\tau_2$. We do not specifically discuss the schedulability test over there. Our intention is to show that the period enforcer algorithm is not always superior to the original fixed-priority scheduling. That is, the enforcement makes this segment miss the deadline, but, in the fixed-priority scheduling there is no deadline miss. We leave the discussions about the task transformation and the schedulability test in Section 4. To be more precise about the above statement, we have rewritten Section 4 to answer two fundamental questions: First, how can we derive the corresponding deferrable task set efficiently? Second, how should we perform the schedulability test on the corresponding deferrable task set correctly?
}

\action{Section 4 was completely rewritten to be more clear.}

{\bf Q:B.2}:    
Section 4 should be reworded. 'precisely converting a self-suspending task
system into a corresponding single-segment deferrable task set is not an
easy problem'  $\rightarrow$  'converting a self-suspending tasks system into a
single-segment deferrable task set introduced pessimism due to period
enforcement'. It is unclear how the Nelissen paper helps since it does not
assume period enforcement. If each self-suspending task is broken into
deferrable sub tasks (considering all the higher priority tasks as
non-suspending), why is the Nelissen analysis needed? Also, It should be
clarified as to why the problem of precisely computing $W_k^{j}$ is useful in
the first place since period enforcement would introduce pessimism anyway
(as shown in Section 3).

\response{We do not fully capture the first sentence here from the reviewer. There is no additional pessimism due to the period enforcer algorithm. The pessimism is due to the assumption made by the period enforcer algorithm to allow only a set of deferrable tasks (that only suspend once at beginning). In the revision, we completely rewrote Section 4 to explain the gap between the segmented self-suspending task model and the corresponding deferrable sporadic task set. We need the analysis from Nelissen et al. to precisely break a self-suspending task into multiple deferrable subtasks so that we can calculate the worst-case release jitter of a computation segment. The new structure of Section 4 first explains why it is difficult to precisely calculate the release jitter of the computation segments and then explains that it is rather important to also perform the schedulability test correctly. We show that the schedulability test can be wrong if the test is not done well. }

{\bf Q:B.3}:    
Section 5 should also include the scenario where period enforcement is only
applicable to the normal execution segments. It seems practical to not
enforce critical sections since they are usually short in most systems i.e.
the back-to-back execution overhead from most critical sections in
real-world scenarios should be minimal. The case 3 that should be discussed
is period enforcement applied to execution segment after the critical
section, without enforcing the critical section and just including its
back-to-back execution penalty in the worst-case analysis.

\response{ leave to Bj\"orn}

{\bf Reviewer C}:

{\bf Q:C.1}:    
This reviewer was not able to download a complete version of [20] including
the figures (the .ps available online is highly corrupted). Please provide a
precise link to the full version of this paper or a way to access to it.

\response{We are not able to provide a complete version of the paper since the paper is not from us. We also used the same version (without figures) to study the period enforcer algorithm. In the revision, we have decided to detail more steps to improve the consistency of our paper. }

{\bf Reviewer D}:

	The period enforcer algorithm (proposed by Rajkumar [20]) was designed to
reduce the scheduling penalty caused by self-suspending tasks. This paper
presents three important observations regarding the applicability of period
enforcer algorithm and its schedulability analysis to multi-segment
self-suspending tasks. First, it is shown that period enforcement may cause
deadline miss in multi-segment self-suspending tasks that are schedulable
without period enforcement. The original analysis of period enforcer
algorithm [20] shows that if a set of single-segment (deferrable) tasks is
schedulable, then the task set is also schedulable under period enforcement.
Therefore, to apply this result from [20] to a set of multi-segment
self-suspending tasks, a transformation from multi-segment to single-segment
task set is needed. Second, it is shown that such transformation is quite
difficult (i.e., currently only exponential time solution is known). Third,
it is shown that period enforcer algorithm requires new analysis of existing
suspension-based locking protocol to compute blocking factor. In other
words, period enforcer algorithm is incompatible with all existing analyses
of suspension-based locking protocols. All these three (negative)
observations show the limitations of the applicability of period enforcer
algorithm to multi-segment self-suspending task set. These limitations are
demonstrated using examples that are easy to follow (Section 3, 4 and 5).

{\bf Q:D.1}:  
This paper is heavily dependent on the assumptions and analysis of the
original period enforcer algorithm in [20]. Although the authors highlighted
some of the assumptions and important result (e.g., Theorem 5) from [20], it
may be difficult for readers to understand this paper without reading in
details the analysis from [20]. The reviewer had difficulty in understanding
the discussion in section 2.3 and 2.4. After the reviewer read [20], the
discussion in section 2.3 and section 2.4 becomes clearer. Important details
from [20] need to be added to this paper to make it more readable and
self-contained.

\response{leave this to Bj\"orn}

{\bf Q:D.2}:  
There are some terms in section 2.3 and 2.4 that the authors have used
assuming that readers are already familiar with those terms. For example,
the terms “deferrable task”, “task set transformation”,
“single-segment deferrable tasks”, “non-self-suspending sporadic
task”, etc. may be formally defined before using them. For example, the
set of transformed non-self-suspending tasks subject to release jitter
(section 2.4) essentially means a set of deferrable tasks. Since Theorem 5
uses the term deferrable task, the relationship between transformed task set
and deferrable task set may be explicitly stated for better understanding.

\response{leave this to Bj\"orn as well}

{\bf Q:D.3}:  
Theorem 5 (page 5) mentions about “worst-case” condition. What does it
mean by worst-case condition for deferrable tasks? Paper [20] explains the
worst-case. The authors could explain the worst-case or restate the theorem
in a different way so that reader does not need to seek for the worst-case
in [20].

\response{leave this to Bj\"orn as well.
message from JJ:
Can you think how to improve that? I think a table would be nice. We may need to put more words from Raj’s paper.. 
But, let’s assume that Raj decides to take the paper down from his website. Then, no one knows what we are talking about precisely. So, maybe it is a good idea to make Section 2 more self-contained?
}

{\bf Q:D.4}:  
Is there any way to quantify (theoretically or using simulation) the
schedulability loss/gain using period enforcer algorithm for multi-segment
self-suspending tasks? The reviewer is curious to learn the impact of the
negative (theoretical) results of period enforcer algorithm on randomly
generated multi-segment self-suspending task sets.

\response{We are not able to provide the (theoretical or simulated) loss/gain due to the following reasons. 1) We are not able to convert a segmented self-suspending task set into the corresponding deferrable task set efficiently and without any information loss. In Section 4.1, we explain the task set transformation for a special case. If there are multiple self-suspending tasks, there is no existing solution to exactly construct the deferrable task set. 2) There is no clear evidence on how such schedulability tests for the corresponding deferrable task set should be designed precisely. We have demonstrated in Section 4.2 that all the deferrable tasks (even constructed from the same self-suspending task) should be considered for analyzing the interference. But, this implies that the worst-case behaviour of the last computation segment can be very pessimistic. Based on the example provided in Section 4.2, one can imagine that the calculation of the deferrable time already considers the earlier computation segments and the calculation of the worst-case interference has to account for the earlier computation segments as well (even if there is no overlap). We are not sure whether this is due to the design or due to the pessimistic analysis. Please refer to Section 4.2 for this. 3) There are also other methods that have been developed recently to handle the self-suspensions, and we are not sure what to compare with. }

\action{Section 4 has been completely rewritten to explain the difficulty.}

{\bf Q:D.5}:   
Based on the notes of this paper, it is not clear whether one should consider period enforcer algorithm or not for self-suspending tasks.

\response{From our note, we can only support the period enforcer algorithm as stated in the conclusion: ``\emph{Nevertheless, Theorem 5 in [21] could be useful for handling self-suspending tasks (that do not use suspension-based locks) if there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm. However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.}'' However, there are several drawbacks that were never discussed earlier, as pointed out at the end of Section 1. The period enforcer algorithm can still be potentially used with cares. }

{\bf Q:D.6}:   
Why the dynamic self-suspension model is presented and compared with
segmented model in section 2.1? It is not clear why period enforcer
algorithm applies only to the segmented model. Is there any reference or
argument to support this claim in line 61-62?

\response{There was no reference to support this claim. We explain the reason in Section 2.3 in the revision.}

\action{The following paragraph is added to Section 2.3.  Note that it is not possible to convert a dynamic self-suspending task into the notion of deferred execution defined in [21], since the definition of a dynamic self-suspending task creates dynamic execution amounts for different jobs of a task. A simple example can explain why the period enforcer algorithm is not compatible with the dynamic self-suspending task model. Suppose that the system has only one task with execution time $C_1=1$, $S_1=1$, and $D_1=T_1=2$. The first job of task $\tau_1$ arrives at time $0$, suspends itself for one time unit, and then executes for one time unit. The second job of task $\tau_1$ arrives at time $2$, first executes for $0.5$ time unit, then suspends for $1$ time unit, and executes for $0.5$ time unit. With the period enforcer algorithm, the second job of task $\tau_1$ starts its execution at time $3$ and will clearly miss the deadline at time $4$.}
\end{document}

### 2016-05-19 locking.tex
take out unneeded macros (compilation trouble)
minor tweaks in Section 5

@@
 Such self-suspensions due to lock contention, just like any other self-suspension, result in deferred execution and thus can detrimentally affect a task's interference on lower-priority tasks.
-It may thus seem natural to apply the period enforcer  to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91} and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.} However, as we demonstrate with two examples, it is actually not safe to use the period enforcer in the presence of suspension-based locks.
+It may thus seem natural to apply the period enforcer  to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91} and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.
+As we show in this paper, this is unsafe.} However, as we demonstrate with two examples, it is actually not safe to apply period enforcement to lock-induced self-suspensions.
 
@@
 
-\ifpaper
 \begin{figure}[t]
@@
 \end{figure}
-\fi
 
@@
 Furthermore, if both tasks continue the illustrated execution pattern, the period enforcer continues to increase their response times.
-As a result, the pattern may be repeated to construct schedules in which an arbitrarily large implicit or constrained deadline is violated.
+As a result, the pattern may be repeated to construct schedules in which any arbitrarily large implicit or constrained deadline is violated.
 
@@
 
-While it is intuitively appealing to combine period enforcement with suspension-based locking protocols, we observe that this causes non-trivial difficulties.
+While it is intuitively appealing to combine period enforcement with suspension-based locking protocols~\cite{Raj:91,Lak:11,LNR:09}, we observe that this causes non-trivial difficulties.
 In particular, our examples show that the addition of period enforcement invalidates all existing blocking analyses.

### 2016-05-19 transformation.tex
redraw figure 5
polish Section 4 and include 2 references

@@
 Second, how should we perform the schedulability test on the corresponding deferrable task set correctly?
+We will discuss the difficulty to derive the corresponding deferrable task set in Section~\ref{sec:transformation-exponential} and the potential pessimism for the schedulability test based on the deferrable task set in Section~\ref{sec:schedulability-test-deferrable}.
 
@@
 \cite{ecrts15nelissen}, which analyzed the exact worst-case response time for segmented self-suspending sporadic tasks with exponential time complexity.
-The worst-case response time analysis in \cite{ecrts15nelissen} is exact under the following conditions: 1) the task set has only one segmented self-suspending task, 2) the self-suspending task is the lowest-priority task, 3) the scheduling policy is preemptive fixed-priority scheduling, and 4) $D_i \leq T_i$ for all task $\tau_i$ in the task system.
+The worst-case response time analysis in \cite{ecrts15nelissen} is exact under the following conditions: 1) the task set has only one segmented self-suspending task, 2) the self-suspending task is the lowest-priority task, 3) the scheduling policy is preemptive fixed-priority scheduling, and 4) $D_i \leq T_i$ for all task $\tau_i$ in the task system.\footnote{Here, we refer to the characteristics of the worst-case release pattern provided in Lemma 2 in \cite{ecrts15nelissen}.
+The exact worst-case response time is to explore all the release patterns that satisfy the conditions in Lemma 2 in \cite{ecrts15nelissen} instead of the proposed MILP in \cite{ecrts15nelissen} (which only provides an upper bound).
+}
 
@@
 In particular, Nelissen et al.
-\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.
+\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.\footnote{The schedulability test problem for such a case has been recently proved to be coNP-hard in the strong sense by Jian-Jia Chen in a report.}
 
 In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
-However, finding a safe upper bound of $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ can be done in pseudo-polynomial time \cite{PH:rtss98}, if over-approximations can be tolerated.
+However, finding a safe upper bound of $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ can be done in pseudo-polynomial time \cite{PH:rtss98,Huang:multiseg}, if over-approximations can be tolerated.
 
@@
 
-Here, we explain the potential schedulability test that can be used for a corresponding deferrable task set.
-It may be seemingly correct at beginning that we do not have to consider the interference from the other deferrable tasks created from one segmented self-suspending task.
-For example, in the special task set considered in Section~\ref{sec:transformation-exponential}, one may argue that it is not necessary to consider the interference from the other computation segments of task $\tau_k$ while analyzing the feasibility of the last computation segment.
-However, this statement is incorrect.
+Suppose that the corresponding deferrable task set can be derived precisely by ignoring the complexity issues raised in Section~\ref{sec:transformation-exponential}.
+The next question is whether there is any information loss, with respect to the schedulability test, by inspecting only the corresponding deferrable task set.
+We will demonstrate an example to explain the information loss.
 
 Consider a task system consisting of $2$ tasks.
-Let $\tau_1$ denote a sporadic task without self-suspensions and parameters $C_1 = 2$ and $T_1=D_1=10$, and let $\tau_2$ denote a self-suspending task consisting of three segments with parameters  $C_2^1 = 1$,  $S_2^1 = 6$, $C_2^2=1$, $S_2^2=8$, $C_2^3=1$, and $ T_2=D_2=21$.
+Let $\tau_1$ denote a sporadic task without self-suspensions, in which $C_1 = 2$ and $T_1=D_1=10$.
+Let $\tau_2$ denote a self-suspending task consisting of three computation segments with parameters  $C_2^1 = 1$,  $S_2^1 = 6$, $C_2^2=1$, $S_2^2=8$, $C_2^3=1$, and $ T_2=D_2=21$.
 That is, this task set is pretty similar to the task set used in Section~\ref{sec:unschedulable}.
@@
 \end{equation*}
-If we do not consider the interference from the first two computation segments of task $\tau_2$, we can conclude that the last computation segment of task $\tau_2$ has worst-case deferrable time (release jitter) $18$ and the worst-case response time of a job of task $\tau_2$ is hence $18+C_1+C_2^3=21$.
+Therefore, in the corresponding deferrable task set, there are $2$ sporadic tasks $\tau_1$ and $\tau_2^1$ (with execution time $1$ and period $21$), and two deferrable tasks $\tau_2^2$ (with deferrable length $9$, period $21$, and execution time $1$) and $\tau_2^3$ (with deferrable length $18$, period $21$, and execution time $1$).
 
-However, the above statement is not correct. Let us consider this task set under control of the period enforcer algorithm, as defined in Section \ref{sec:pe}.
+Suppose that we are interested to analyze the worst-case response time of the deferrable task $\tau_2^3$ (in the corresponding deferrable task set) under fixed-priority scheduling.
+The question is \emph{whether we should ignore or should include the interference due to $\tau_2^1$ and $\tau_2^2$}.
+Since the computation segments that are used to create $\tau_2^1, \tau_2^2$, and $\tau_2^3$ do not have any overlap in their available time to be executed, it may be seemingly correct at beginning that we do not have to consider the interference from $\tau_2^1$ and $\tau_2^2$ when analyzing the response time of the deferrable task $\tau_2^3$.
+However, this statement is in fact incorrect.
+
+If we do not consider the interference from $\tau_2^1$ and $\tau_2^2$, we can conclude that the worst-case response time of $\tau_2^3$ is  $18+C_1+C_2^3=21$.
+Let us consider this task set under the control of the period enforcer algorithm, as defined in Section \ref{sec:pe}.
 Figure \ref{fig:example-deferrable-incompatible} shows the resulting schedule for a periodic release pattern. This changes the schedule of the second job of task $\tau_2$. 
@@
 For such a case, the response time of the second job of task $\tau_2$ is $23$.
-This means that the worst-case response time of the deferrable task created to represent the third computation segment of task $\tau_2$ must be at least $23-\rho_2^3=5$, which equals to $C_1 + C_2^1+C_2^2+C_2^3$.
+This means that the worst-case response time of the deferrable task$\tau_2^3$ must be at least $23$, which equals to $18+C_1 + C_2^1+C_2^2+C_2^3$.
 
-Therefore, for this specific example, we know that the schedulability test for the  corresponding deferrable task set has also to be pessimistic enough to consider the interferences from all the computation segments.
+Therefore, for this specific example, we know that the schedulability test for the  corresponding deferrable task set has also to be pessimistic enough to consider the interferences from all the computation segments created from the same self-suspending sporadic task.
 
@@
   \centering  
-  \includegraphics[scale=0.4]{../figures/deferrable-incompatible.pdf}
-  \caption{}
+  \includegraphics[scale=1]{../figures/deferrable-incompatible/deferrable-incompatible.pdf}
+  \caption{An example for explaining why it is not safe to ignore the interference from the other computation segments in the same task.}
   \label{fig:example-deferrable-incompatible}

### 2016-05-19 untitled.tex
take out unneeded macros (compilation trouble)
polish Section 4 and include 2 references

@@
 When real-time tasks suspend themselves (due to blocking I/O, lock contention, etc.), they defer a part of their execution to be processed at a later time.
-A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}.
+A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen}.
 This penalty, which is maximized when a task defers the completion of one job just until the release of the next job, can manifest as response-time increases and thus may lead to deadline misses.
@@
 %\clearpage
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "LITES/LITES-Paper.tex"
+%%% End:

### 2016-05-31 untitled.tex
new discussion: review relevant task models and clearly define task set transformation

@@
 \section{Introduction} 
-When real-time tasks suspend themselves (due to blocking I/O, lock contention, etc.), they defer a part of their execution to be processed at a later time.
+When real-time tasks suspend themselves (due to blocking I/O, lock contention, \textit{etc.}), they defer a part of their execution to be processed at a later time.
 A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen}.
@@
 The period enforcer algorithm~\cite{Raj:suspension1991} applies to self-suspending tasks on uniprocessors under fixed-priority scheduling, and hence by extension also to multiprocessors under partitioned fixed-priority scheduling (where tasks are statically assigned to processors and each processor is scheduled as a uniprocessor).
-In this section, we review the underlying task model (Section \ref{sec:taskmodel}), introduce the period enforcer algorithm (Section \ref{sec:pe}), summarize its analysis (Section \ref{sec:classic-analysis}), and finally restate our observations (Section \ref{sec:questions}).
+In this section, we review the underlying task model (Section \ref{sec:taskmodel}), introduce the period enforcer algorithm (Section \ref{sec:pe}), summarize its analysis (Section \ref{sec:classic-analysis}), and finally restate our observations in more precise terms (Section \ref{sec:questions}).
 
-\subsection{Self-Suspending Tasks}
+\subsection{Task Models}
 \label{sec:taskmodel}
 
-To date, the real-time literature on self-suspensions has focused on two task models: the \emph{dynamic} and the \emph{segmented} (or \emph{multi-segment}) self-suspension model. 
-The dynamic self-suspending sporadic task model characterizes each
-task $\tau_i$ as a four-tuple $(C_i,S_i,T_i,D_i)$: 
-$C_i$ denotes an upper bound on the total execution time of any job of $\tau_i$,
-$S_i$ denotes an upper bound on the total self-suspension time of any job of $\tau_i$,
-$T_i$ denotes the minimum inter-arrival time (or period) of $\tau_i$, and $D_i$ is the relative deadline.
+Since the analysis of the period enforcer requires reasoning about different task models and their relationships, we carefully introduce and precisely define the relevant models in this section.
+
+\subsubsection{Periodic Tasks}
+
+The most basic and best understood task model is the \emph{periodic task model} due to Liu and Layland~\cite{LL:73}.
+In this model, each task $\tau_i$ is characterized as a tuple $(C_i,T_i)$, where $C_i$ denotes an upper bound on the total execution time of any job of $\tau_i$ and $T_i$ denotes the (exact) \emph{inter-arrival time} (or \emph{period}) of $\tau_i$.
+Each such periodic task $\tau_i$ releases a job at time~0, and periodically every $T_i$ units thereafter.
+Each job must finish by the time the next arrives.
+Importantly, Liu and Layland assumed both that the $k$\xth job of $\tau_i$ arrives \emph{exactly} at time $(k-1)\times T_i$, and that an incomplete job is \emph{always} available for execution (i.e., jobs never block on I/O or locks).
+
+A straightforward generalization of the periodic task model is to introduce an explicit \emph{relative deadline} parameter $D_i$.
+In this case, each task is represented by a three-tuple $(C_i, T_i, D_i)$, with the interpretation that every job of $\tau_i$ must finish within $D_i$ time units after its release.
+Task $\tau_i$ is said to have an \emph{implicit deadline}  if $D_i = T_i$, a \emph{constrained deadline} if $D_i \leq T_i$, and an \emph{arbitrary deadline} otherwise.
+We primarily consider implicit deadlines in this note.
+
+\subsubsection{Sporadic Tasks}
+
+Mok~\cite{Mo:83} introduced the \emph{sporadic task model}, in which each task $\tau_i$ is still specified by a tuple $(C_i, T_i, D_i)$.
+However, the sporadic task model relaxes the inter-arrival constraint $T_i$ to specify a \emph{minimum} (rather than an exact) separation between jobs.
+In this interpretation, the first job is not necessarily released at time~0, and the exact release times of future jobs cannot be predicted, which is an appropriate modeling assumption for event-triggered tasks.
+
+On uniprocessors, this relaxation does not introduce additional pessimism: since any two jobs of a sporadic task $\tau_i$ are  known to be released \emph{at least} $T_i$ time units apart, the sporadic task model~\cite{Mo:83} still allows for schedulability analysis that is as accurate as Liu and Layland's analysis of periodic tasks~\cite{LL:73}.
+
+Mok retained the assumption that incomplete jobs are always ready for execution (i.e., no suspensions), and that jobs, once released, are \emph{immediately} available for execution.
+
+\subsubsection{Release Jitter}
+\label{sec:jitter}
+
+The latter assumption --- immediate availability for execution --- is inappropriate in many practical systems (especially in networked systems) if events (e.g., messages) that trigger job releases can incur non-negligible delays (e.g., network congestion).
+Such delays in task activation can accounted for by introducing a notion of \emph{release jitter}.
+To this end, each task is represented by a four-tuple $(C_i, J_i, T_i, D_i)$, where the parameter $J_i$ is a bound on the maximum time that a job remains unavailable for execution after it should have started to run.
+Release jitter can be incorporated in both the periodic and the sporadic task models.
+
+In the presence of release jitter, the terms ``job arrival'' and ``job release,'' which are often used interchangeably, take on distinct meanings: a job's \emph{arrival time} denotes the point in time when it actually becomes available for execution, whereas a job's \emph{release time} is the instant that is relevant for the (minimum) inter-arrival time constraint.
+Any job of task $\tau_i$ thus arrives at most $J_i$ time units after it is released.
+
+Notably, non-zero release jitter \emph{does} cause additional pessimism: in the worst case, two consecutive jobs of a task $\tau_i$ can be separated by as little as $T_i - J_i$ time units (if the earlier job incurs maximum release jitter and the successor job incurs none).
+As a result, a task may ``carry in'' some additional work into a given analysis interval.
+Taking this effect into account, Audsley et al.~\cite{ABRTW:93} developed a response-time analysis for sporadic and periodic constrained-deadline tasks subject to release jitter under preemptive fixed-priority scheduling.
+
+However, even in the presence of release jitter, a key assumption remains that jobs do not self-suspend (e.g., wait for I/O).\footnote{Audsley et al.~\cite{ABRTW:93} do present a response-time analysis that takes into account a limited form of suspensions due to semaphores (``blocking'').
+However, their analysis does not apply to general self-suspensions (i.e., the kind of self-suspensions targeted by period enforcer algorithm)  and is not relevant in the context of this paper.} That is, it is assumed that, once a job has arrived, it continuously remains available for dispatching until it completes.
+This restriction is removed next.
+
+\subsubsection{Self-Suspending Tasks}
+\label{sec:dynamic}
+
+When a job \emph{self-suspends}, it becomes unavailable for execution until some external event occurs (e.g., a disk I/O operation completes, a network packet arrives, a co-processor signals completion, \textit{etc.}).
+This has the effect of \emph{deferring} (a part of) the job's processing requirement until the time that it \emph{resumes} from its suspension, which causes massive analytical difficulties~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen,Ri:04}.
+
+To date, the real-time literature on self-suspensions has focused on two task models: the \emph{dynamic} self-suspension model, which we discuss first,  and the \emph{segmented} (or \emph{multi-segment}) suspension model, which we discuss next in Section~\ref{sec:segmented}.
+Self-suspensions can arise in both periodic and sporadic tasks (i.e, both interpretations of the $T_i$ parameter are possible); in this note we focus on sporadic tasks.
+
+The dynamic self-suspending task model characterizes each
+task $\tau_i$ as a four-tuple $(C_i,S_i,T_i,D_i)$: the parameters $C_i$, $T_i$, and $D_i$ have their usual meaning (i.e., as in the periodic and sporadic task models), and $S_i$ denotes an upper bound on the total self-suspension time of any job of $\tau_i$.
 The dynamic self-suspension model does not impose a bound on the maximum number of self-suspensions, nor does it make any assumptions as to where during a job's execution self-suspensions occur.
+That is, how often a job defers its execution, when it does so, and how much it defers varies unpredictably from job to job.
 
-In contrast, the segmented self-suspending sporadic task model extends the above four-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals.
+Allowing tasks to self-suspend can impose substantial scheduling penalties (an example is provided shortly in Section~\ref{sec:pe}) and greatly complicates  schedulability analysis (e.g., see~\cite{ecrts15nelissen,Ri:04,Chen2016}).
+In particular, release jitter and self-suspensions are not interchangeable concepts and it is not safe~\cite{Chen2016,ecrts15nelissen} to simply substitute $J_i$ with $S_i$ in Audsley et al.'s analysis~\cite{ABRTW:93}.
+However, under the dynamic suspension model, it is possible for jobs of self-suspending tasks to defer their entire execution requirement.
+
+The period enforcer algorithm aims to mitigate the negative effects of self-suspensions.
+However, for reasons that will be explained in Section~\ref{sec:pe}, the period enforcer algorithm cannot be meaningfully combined with the dynamic suspension model.
+Instead, it requires a segmented suspension model, which we discuss next.
+
+\subsubsection{Segmented Self-Suspending Tasks (aka Deferrable Tasks)}
+\label{sec:segmented}
+
+The (multi-)segmented self-suspending sporadic task model extends the  four-tuple $(C_i,S_i,T_i,D_i)$ by characterizing each self-suspending task as a fixed, finite linear sequence of computation and suspension intervals.
 These intervals are represented as a tuple
-$(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, which is composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals.
-For simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment.
-Any suspension before the first computation segment is equivalent to \emph{release jitter}~\cite{ABRTW:93}.
+$(S_{i}^0,C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, which is composed of $m_i$ computation segments separated by $m_i$ suspension intervals.
 
-A special case of the segmented self-suspending sporadic task model is the deferrable sporadic task model, in which each task $\tau_i$ starts with a self-suspension interval followed by one computation segment.
-The deferrable sporadic task model is the central notation in Rajkumar's analysis~\cite{Raj:suspension1991}.
-We will explicitly explain the link in Section~\ref{sec:classic-analysis}.
-A deferrable sporadic task is equivalent to
-an ordinary sporadic task with \emph{release jitter}.
-
-The advantage of the dynamic model is that it is more flexible since it does not impose any assumptions on the task control flow.
-The advantage of the segmented model is that it allows for more accurate analysis.
-The period enforcer algorithm and its analysis applies (only) to the segmented model, to be explained in Section~\ref{sec:classic-analysis}.
+The first self-suspension segment $S_i^0$, prior to the first execution segment, is equivalent to release jitter ($J_i$, Section~\ref{sec:jitter}).
+However, in many works on the segmented self-suspending task model, the segment $S_i^0$ assumed to be absent (i.e., $S_i^0 = 0$), such that there are only $m_i - 1$ suspension intervals (and jobs arrive jitter-free).
+Unless noted otherwise we adopt this convention.
 
 We say that a segment \emph{arrives} when it becomes available for execution.
-The first computation segment arrives immediately when the job is released; the second computation segment (if any) arrives when the job resumes from its first self-suspension, \textit{etc.} For simplicity, we assume that tasks are indexed in order of decreasing priority (i.e., $\tau_1$ is the highest-priority task).
-A \emph{level-$i$ busy interval} is a maximal interval during which  the processor executes only segments of tasks with priority $i$ or higher.
+The first computation segment arrives immediately when the job is released (unless $S_i^0 \neq 0$); the second computation segment (if any) arrives when the job resumes from its first self-suspension, \textit{etc.}
+
+The advantage of the dynamic model (Section~\ref{sec:dynamic}) is that it is more flexible since it does not impose any assumptions on the task control flow.
+The advantage of the segmented model is that it allows for more accurate analysis.
+The period enforcer algorithm and its analysis~\cite{Raj:suspension1991} applies (only) to the segmented model, as explained in Section~\ref{sec:classic-analysis}.
+
+A note on terminology: for the sake of consistency with the recent literature on self-suspensions in real-time systems, we favor the term ``segmented self-suspending tasks'' to refer to tasks under the just-introduced model.
+However,  Rajkumar's original description of the period enforcer~\cite{Raj:suspension1991} refers to such tasks as  \emph{deferrable tasks}, as it predates the widespread adoption of the former term.
+We use the former terminology in this note, unless when quoting Rajkumar~\cite{Raj:suspension1991}, in which case we use Rajkumar's original terminology.
+
+\subsubsection{Single-Segment Self-Suspending (aka Deferrable) Tasks}
+\label{sec:single-segmented}
+
+Segmented self-suspending tasks with exactly one self-suspension interval followed by exactly one computation segment ($m_i = 1$, $S_i^0 \neq 0$) are an important special case, which we refer to as \emph{single-segment self-suspending tasks}.
+This special case is central to Rajkumar's original analysis of the period enforcer~\cite{Raj:suspension1991},  as we will explain in Section~\ref{sec:classic-analysis}.
+Regarding terminology, Rajkumar~\cite{Raj:suspension1991} does not use a special term for single-segment self-suspending tasks, simply referring to them as deferrable tasks.
+To avoid ambiguity, we instead explicitly mention the ``single-segment'' qualifier.
+
+Note also that single-segment self-suspending sporadic tasks, which are  ``suspended'' only prior to commencing execution, are analytically fully equivalent to sporadic task subject to release jitter (i.e., the model described in Section~\ref{sec:jitter}).
+We nonetheless use the term ``single-segment self-suspending sporadic task'' to remain close to Rajkumar's original description~\cite{Raj:suspension1991}, and to highlight the connection to the (multi-)segmented self-suspending task model (Section~\ref{sec:segmented}).
+
+This concludes our review of relevant task models. Before reviewing the period enforcer and its original analysis, we briefly introduce to some essential concepts.
+
+\subsubsection{Assumptions, Busy Periods, and Task Set Transformations}
+\label{sec:misc-defs}
+
+We focus exclusively on preemptive fixed-priority scheduling in this note, as the period enforcer is explicitly designed for this setting.
+For simplicity, we assume that tasks are indexed in order of decreasing priority (i.e., $\tau_1$ is the highest-priority task).
+
+A key concept in the period enforcer's runtime rules (discussed next) is the notion of a \emph{level-$i$ busy interval}, which is a maximal interval during which  the processor executes only segments of tasks with priority $i$ or higher.
+
+Finally, Rajkumar's original analysis~\cite{Raj:suspension1991} of the period enforcer is rooted in the concept of a \emph{task set transformation}.
+In general, such a task set transformation is simply a function $f$ that maps a given task set $\tau$ to a transformed task set $\tau' = f(\tau)$ \emph{such that $\tau'$ is schedulable \textbf{only if} the original task set $\tau$ is schedulable}, too.
+The basic idea is that such a transformation allows a form of reduction: $\tau$ can now \emph{indirectly} be shown to be schedulable by establishing that $\tau'$ is schedulable.
+
+Importantly, the tasks in $\tau$ and $\tau'$ do \emph{not} have to be of the same task model, nor does the number of tasks have to remain constant (i.e.,  $|\tau| \neq |\tau'|$ is possible).
+Specifically, the task set transformation underlying the analysis of the period enforcer maps each \emph{multi-}segmented self-suspending task $\tau_i \in \tau$  to $m_i$ \emph{single-}segmented self-suspending tasks in $\tau'$ (i.e., $|\tau'| = \sum_{\tau} m_i$ ).
+
+With these definitions in place, we can now introduce the period enforcer.
+We first introduce the actual algorithms (i.e., the rules enforced at runtime) in Section~\ref{sec:pe} and then review the corresponding original analysis in Section~\ref{sec:classic-analysis}.
 

### 2016-05-31 untitled.tex
snapshot — work in progress section 2

@@
 
-\subsubsection{Segmented Self-Suspending Tasks (aka Deferrable Tasks)}
+\subsubsection{Segmented Self-Suspending Tasks}
 \label{sec:segmented}
@@
 
-Recall that the scheduling penalty associated with self-suspensions is maximized when a task defers the completion of one job just until the release of the next job.
+XXX
+
+\subsubsection{The Problem: Back-to-Back Execution}
+
+The scheduling penalty associated with self-suspensions is maximized when a task defers the completion of one job just until the release of the next job.
 For example, this effect is illustrated in Figure \ref{fig:not-ok-without-period-enforcement}, which shows a case in which the self-suspension of the higher-priority task $\tau_2$  results in a deadline miss of the lower-priority task $\tau_3$.
+
 The root cause is increased interference due to the ``back-to-back'' execution effect~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}, where two jobs of $\tau_2$ execute in close succession (i.e, separated by less than a period) because the second job self-suspended for a (much) shorter duration than the first job.
-That is, $\tau_3$ misses its deadline because $\tau_2$ resumed ``too soon.''
+In the example shown in Figure \ref{fig:not-ok-without-period-enforcement},  $\tau_3$ misses its deadline because of this effect as $\tau_2$ resumed ``too soon.''
 
@@
 \end{figure}
+
+\subsubsection{The Period Enforcement Rule}
 
@@
 
-Figure \ref{fig:ok-with-period-enforcement} illustrates how this definition of eligibility times restores the schedulability of the task set depicted in Figure~\ref{fig:not-ok-without-period-enforcement}.
+\subsubsection{Example: Hiding the Effects of Deferred Execution}
+
+Figure \ref{fig:ok-with-period-enforcement} illustrates how the definition of eligibility time in Equation~\ref{eq:ET-def} restores the schedulability of the task set depicted in Figure~\ref{fig:not-ok-without-period-enforcement}.
 Consider the eligibility times of the second segment of task $\tau_2$.
@@
 The examples in Figures \ref{fig:not-ok-without-period-enforcement} and \ref{fig:ok-with-period-enforcement} suggest an intuition for the benefits provided by period enforcement: computation segments of a self-suspending task $\tau_i$ are forced to execute at least $T_i$ time units apart (hence the name), which ensures that it causes no more interference than a regular (non-self-suspending) sporadic task.
+
+\subsection{Incompatibility with the Dynamic Self-Suspension Model}
+
+Note that it is not possible to convert a dynamic self-suspending task into the notion of deferred execution defined in \cite[Section 3]{Raj:suspension1991}, since the definition of a dynamic self-suspending task creates dynamic execution amounts for different jobs of a task.
+A simple example can explain why the period enforcer algorithm is not compatible with the dynamic self-suspending task model.
+Suppose that the system has only one task with execution time $C_1=1$, $S_1=1$, and $D_1=T_1=2$.
+The first job of task $\tau_1$ arrives at time $0$, suspends itself for one time unit, and then executes for one time unit.
+The second job of task $\tau_1$ arrives at time $2$, first executes for $0.5$ time unit, then suspends for $1$ time unit, and executes for $0.5$ time unit.
+With the period enforcer algorithm, the second job of task $\tau_1$ starts its execution at time $3$ and will clearly miss the deadline at time $4$.
+
  Next, we revisit the original analysis of this technique.
@@
 
-Note that it is not possible to convert a dynamic self-suspending task into the notion of deferred execution defined in \cite[Section 3]{Raj:suspension1991}, since the definition of a dynamic self-suspending task creates dynamic execution amounts for different jobs of a task.
-A simple example can explain why the period enforcer algorithm is not compatible with the dynamic self-suspending task model.
-Suppose that the system has only one task with execution time $C_1=1$, $S_1=1$, and $D_1=T_1=2$.
-The first job of task $\tau_1$ arrives at time $0$, suspends itself for one time unit, and then executes for one time unit.
-The second job of task $\tau_1$ arrives at time $2$, first executes for $0.5$ time unit, then suspends for $1$ time unit, and executes for $0.5$ time unit.
-With the period enforcer algorithm, the second job of task $\tau_1$ starts its execution at time $3$ and will clearly miss the deadline at time $4$.
-
-%
 Central to Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{task set transformation} that splits each deferrable task with multiple segments into a corresponding number of single-segment deferrable tasks.

### 2016-06-06 untitled.tex
revise section 2.2

@@
 With these definitions in place, we can now introduce the period enforcer.
-We first introduce the actual algorithms (i.e., the rules enforced at runtime) in Section~\ref{sec:pe} and then review the corresponding original analysis in Section~\ref{sec:classic-analysis}.
 
@@
 
-XXX
+The period enforcer consists of two parts: a runtime rule that governs when a self-suspending task may be scheduled, and an (offline) analysis that may be used to assess the temporal correctness of a set of segmented self-suspending tasks (Section~\ref{sec:segmented}) subject to period enforcement.
+Initially, we focus on the runtime rule (i.e., the actual period enforcer algorithm) and then review the corresponding original analysis thereafter in Section~\ref{sec:classic-analysis}.
+We begin with a simple example that highlights the effect that the period enforcer is designed to control.
 
@@
 The scheduling penalty associated with self-suspensions is maximized when a task defers the completion of one job just until the release of the next job.
-For example, this effect is illustrated in Figure \ref{fig:not-ok-without-period-enforcement}, which shows a case in which the self-suspension of the higher-priority task $\tau_2$  results in a deadline miss of the lower-priority task $\tau_3$.
+This effect is illustrated in Figure \ref{fig:not-ok-without-period-enforcement}, which shows a case in which the self-suspension of the higher-priority task $\tau_2$ from time~1 until time~5  results in a deadline miss of the lower-priority task $\tau_3$ at time~15.
 
 The root cause is increased interference due to the ``back-to-back'' execution effect~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}, where two jobs of $\tau_2$ execute in close succession (i.e, separated by less than a period) because the second job self-suspended for a (much) shorter duration than the first job.
-In the example shown in Figure \ref{fig:not-ok-without-period-enforcement},  $\tau_3$ misses its deadline because of this effect as $\tau_2$ resumed ``too soon.''
+In the example shown in Figure \ref{fig:not-ok-without-period-enforcement},  $\tau_3$ misses its deadline at time~15 because of this effect as $\tau_2$'s second job resumes ``too soon'' (i.e, already at time~12 after having been suspended for only one time unit, rather than four time units like the first job of $\tau_2$).
 
@@
 
-The key idea underlying the period enforcer algorithm is to artificially delay the execution of computation segments if a job resumes ``too soon.'' To this end,  the period enforcer algorithm determines for each computation segment an \emph{eligibility time}.
+The key idea underlying the period enforcer algorithm is to artificially delay the execution of computation segments if a job resumes ``too soon.'' To this end,  the period enforcer determines for each computation segment an \emph{eligibility time}.
 If a segment resumes  before its eligibility time, the execution of the segment is delayed until the eligibility time is reached.
@@
 Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$ (i.e., the processor executes only $\tau_i$ or higher-priority tasks throughout the interval $[\mathit{busy}(\tau_i, t'), t']$).
-According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
+The period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
 \begin{align}\label{eq:ET-def}
@@
 \end{align}
-where $ET_{i,0}^k = -T_i$.
+where $ET_{i,0}^k = -T_i$~\cite[Section 3.1]{Raj:suspension1991}.
+This simple and elegant rule has the desirable effect of avoiding all back-to-back execution, which can be easily observed with an example.
 
-\subsubsection{Example: Hiding the Effects of Deferred Execution}
+\subsubsection{Example: Avoiding Back-to-Back Execution}
 
@@
 
-By definition, we have $ET_{2,0}^2 = -T_2 = -10$. At time~5, when the second computation segment of the first job resumes ($a_{2,1}^2 = 5$), we thus have
+By definition,  $ET_{2,0}^2 = -T_2 = -10$. At time~5, when the second computation segment of the first job resumes ($a_{2,1}^2 = 5$), we thus have
 \begin{align*}
@@
   \caption{Example uniprocessor schedule \emph{with} period enforcement assuming the same scenario as depicted in Figure~\ref{fig:not-ok-without-period-enforcement}.
-With period enforcement, task $\tau_3$ does not miss a deadline because task $\tau_2$'s second computation segment is delayed until it no longer imposes undue interference (i.e., it is prevented from resuming ``too soon'').}
+With period enforcement, task $\tau_3$ does not miss a deadline because task $\tau_2$'s second computation segment is delayed until time~15 when it no longer imposes undue interference (i.e., it is prevented from resuming ``too soon'' at time~12).}
   \label{fig:ok-with-period-enforcement}
@@
 
-\subsection{Incompatibility with the Dynamic Self-Suspension Model}
+\subsubsection{Incompatibility with the Dynamic Self-Suspension Model}
 
-Note that it is not possible to convert a dynamic self-suspending task into the notion of deferred execution defined in \cite[Section 3]{Raj:suspension1991}, since the definition of a dynamic self-suspending task creates dynamic execution amounts for different jobs of a task.
+Before reviewing the classic analysis based on this intuition, we briefly comment on the difficulty of combining period enforcement with the dynamic self-suspension model (Section~\ref{sec:dynamic}).
+
+In short, the period enforcer fundamentally requires the segmented self-suspension model~(Section~\ref{sec:segmented}) to be effective because a job of a dynamic self-suspending task exhibits unpredictable execution times between its unpredictably many self-suspensions.
+
 A simple example can explain why the period enforcer algorithm is not compatible with the dynamic self-suspending task model.
-Suppose that the system has only one task with execution time $C_1=1$, $S_1=1$, and $D_1=T_1=2$.
-The first job of task $\tau_1$ arrives at time $0$, suspends itself for one time unit, and then executes for one time unit.
-The second job of task $\tau_1$ arrives at time $2$, first executes for $0.5$ time unit, then suspends for $1$ time unit, and executes for $0.5$ time unit.
-With the period enforcer algorithm, the second job of task $\tau_1$ starts its execution at time $3$ and will clearly miss the deadline at time $4$.
+Consider a trivial system that has only one task with a total execution time $C_1=1$, a total self-suspension length $S_1=1$, and a period and relative deadline of $D_1=T_1=2$.
+Suppose the first job of task $\tau_1$ arrives at time~$0$, suspends itself for one time unit, and then executes for one time unit.
+Further suppose the second job of task $\tau_1$ arrives at time $2$, first executes for $0.5$ time units, then suspends for $1$ time unit, and finally executes for $0.5$ time units.
+With the period enforcer algorithm in place, the second job of task $\tau_1$ starts its execution at time $3$, at which point it will clearly miss its deadline at time $4$.
 
- Next, we revisit the original analysis of this technique.
+In this example, the problem is that the eligibility time of the first computation ``segment'' of the second job is determined by the self-suspension pattern of the first job, even though the first job deferred all of its execution, whereas the second job deferred only a part of its execution.
+Under the more restrictive segmented self-suspension model~(Section~\ref{sec:segmented}), the pattern of self-suspension and computation times is statically fixed by the model; such a mismatch is hence not possible.
+
+Next, we revisit the original analysis of the period enforcer algorithm.
 

### 2016-06-06 untitled.tex
edit section 2.3

@@
 \subsubsection{Incompatibility with the Dynamic Self-Suspension Model}
+\label{sec:incompat}
 
@@
 
-The central notation in Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{deferrable task}, which matches our notion of segmented tasks.  Specifically, Rajkumar states that:
+The central notation in Rajkumar's analysis~\cite{Raj:suspension1991} is a deferrable task, which matches our notion of segmented tasks, as already discussed in Section~\ref{sec:segmented}.
+Specifically, Rajkumar states that:
 \begin{quote}
@@
 
-Central to Rajkumar's analysis~\cite{Raj:suspension1991} is a \emph{task set transformation} that splits each deferrable task with multiple segments into a corresponding number of single-segment deferrable tasks.
+Central to Rajkumar's analysis~\cite{Raj:suspension1991} is a task set transformation (recall Section~\ref{sec:misc-defs}) that splits each deferrable task with multiple segments (Section~\ref{sec:segmented}) into a corresponding number of single-segment deferrable tasks (Section~\ref{sec:single-segmented}).
 In the words of Rajkumar~\cite[Section 3]{Raj:suspension1991}:
@@
 %
-In other words, the transformation can be understood as splitting each self-suspending task into a matching number of non-self-suspending sporadic tasks subject to release jitter, which can be easily analyzed with classic fixed-priority response-time analysis~\cite{ABRTW:93}.
+In other words, the transformation can be understood as splitting each self-suspending task into a matching number of single-segment deferrable tasks (Section~\ref{sec:single-segmented}), which are equivalent to non-self-suspending sporadic tasks subject to release jitter (Section~\ref{sec:jitter}), which can be easily analyzed with classic fixed-priority response-time analysis~\cite{ABRTW:93}.
+To constitute an effective schedulability analysis, the transformation must ensure that, if the transformed set of single-segment deferrable tasks can be shown to be schedulable (e.g., with response-time analysis~\cite{ABRTW:93}), then the original set of multi-segment deferrable tasks is also schedulable under period enforcement.
 
-It is well known that uncontrolled deferred execution (i.e, release jitter) can impose  increased interference on lower-priority tasks because of the potential for ``back-to-back'' execution~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}, as illustrated in Figure \ref{fig:not-ok-without-period-enforcement}.
-
+To summarize, as illustrated in Figure \ref{fig:not-ok-without-period-enforcement},
+uncontrolled deferred execution can impose  increased interference on lower-priority tasks because of the potential for ``back-to-back'' execution~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}.
 The purpose of the period enforcer algorithm is to reduce such penalties for lower-priority tasks without detrimentally affecting the schedulability of self-suspending, higher-priority tasks.
@@
 \begin{quote}
-{\bf Theorem 5}: A deferrable task that is schedulable under its worst-case conditions is also schedulable under the period enforcer algorithm.  \cite{Raj:suspension1991}
+{\bf Theorem 5}: A [single-segment] deferrable task that is schedulable under its worst-case conditions is also schedulable under the period enforcer algorithm \cite{Raj:suspension1991}. 
 \end{quote}
+The ``worst-case conditions'' mentioned in the theorem simply correspond to the case when \textbf{(i)}~a job of a single-segment deferrable task defers its execution for the maximally allowed time $S_i^0$ (i.e., when it incurs maximal release jitter) and \textbf{(ii)} it incurs maximum higher-priority interference (i.e., when its start of execution coincides with a critical instant~\cite{LL:73}).
 

### 2016-06-06 untitled.tex
edit section 2.4

@@
 However,  Rajkumar's original description of the period enforcer~\cite{Raj:suspension1991} refers to such tasks as  \emph{deferrable tasks}, as it predates the widespread adoption of the former term.
-We use the former terminology in this note, unless when quoting Rajkumar~\cite{Raj:suspension1991}, in which case we use Rajkumar's original terminology.
+We use both terms interchangeably in this paper.
 
@@
 
-Theorem 5 (in \cite{Raj:suspension1991}) is a strong result that seemingly enables a powerful analysis approach: if the corresponding transformed set of non-self-suspending tasks (subject to release jitter, without period enforcement) can be shown to be schedulable under fixed-priority scheduling using \emph{any} applicable analysis (e.g.,~\cite{ABRTW:93}), then the period enforcer algorithm also yields a correct schedule.
+Theorem 5 (in \cite{Raj:suspension1991}) is a strong result, as it implies that the period enforcer does not induce any deadline misses.
+This seemingly enables a powerful analysis approach: if the corresponding transformed set of single-segment deferrable tasks can be shown to be schedulable  \emph{without} period enforcement under fixed-priority scheduling using \emph{any} applicable analysis (e.g.,~\cite{ABRTW:93}), then the period enforcer algorithm also yields a correct schedule.
 
 However, recall that, in the original analysis~\cite{Raj:suspension1991}, deferrable tasks are assumed to defer their  execution either completely or not at all (but not parts of it).
-It is hence important to realize that Theorem 5 in \cite{Raj:suspension1991} applies only to the \emph{transformed} task set---Theorem 5 in \cite{Raj:suspension1991} does \emph{not} apply to the \emph{original} set of segmented self-suspending tasks.
-This leads to the first question: \emph{Does schedulability of the transformed set of non-self-suspending tasks (subject to release jitter, without period enforcement) also imply schedulability of the original set of segmented self-suspending tasks (under period enforcement)?}  In Section \ref{sec:unschedulable}, we answer this question in the negative.
+It is hence important to realize that Theorem 5 in \cite{Raj:suspension1991} applies only to the transformed set of \emph{single-segment} deferrable tasks, and that it does \emph{not} apply to the \emph{original} set of multi-segmented self-suspending tasks.
+
+This leads to the first question: \emph{Does schedulability of the  original set of segmented self-suspending tasks (without period enforcement) imply schedulability under period enforcement?} That is, can Theorem 5 (in \cite{Raj:suspension1991}) be generalized to multi-segmented self-suspending tasks?
+In Section \ref{sec:unschedulable}, we answer this question in the negative.
 
 \begin{enumerate}
-	\item There exist sporadic segmented self-suspending task sets that are schedulable under fixed-priority scheduling without any enforcement, but the corresponding schedule by using the period enforcer algorithm is not feasible.
-This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, and not in the context of the original segmented self-suspending task set.
+	\item There exist sporadic segmented self-suspending task sets that are schedulable under fixed-priority scheduling without any enforcement, but that are infeasible under period enforcement.
+This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, and not in the context of the original multi-segmented self-suspending task set.
 \end{enumerate}
 
-Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the following question: \emph{Given a set of sporadic segmented self-suspending tasks, how do we obtain a corresponding set of (single-segment) deferrable tasks?} That is, the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert self-suspension segments into equivalent bounds on release jitter, but it remains unclear \emph{how} this central step should be accomplished.
+Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to solve the task-set transformation problem: \emph{given a set of sporadic segmented self-suspending tasks, how do we obtain a corresponding set of (single-segment) deferrable tasks?} That is, as discussed in Section~\ref{sec:classic-analysis} the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert multi-segmented self-suspending tasks into corresponding sets of single-segment deferrable tasks, but it remains unclear in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished.
 In Section \ref{sec:convert}, we make a pertinent observation.
@@
 \setcounter{enumi}{1}
-	\item Deriving a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem.
+	\item It is an open problem whether it is possible to derive a single-segment deferrable task set corresponding to a given set of  multi-segmented self-suspending tasks in polynomial time.
 Recent findings by Nelissen et al.~\cite{ecrts15nelissen} can be applied, but their method takes exponential time.
@@
 
-Finally, we make a third observation concerning the use of the period enforcer in conjunction with suspension-based multiprocessor locking protocols for partitioned fixed-priority scheduling (such as the MPCP~\cite{LNR:09,Ra:90} or the FMLP~\cite{BLBA:07,BA:08}).
-While it is certainly tempting to apply period enforcement with the intention of avoiding the negative effects of deferred execution due to lock contention (as previously suggested elsewhere~\cite{Raj:91,Lak:11,LNR:09}), we ask: \emph{Does existing blocking analysis remain safe when combined with the period enforcer algorithm?} In Section \ref{sec:locking}, we show that this is not the case.
+Finally, we consider the use of the period enforcer in conjunction with suspension-based multiprocessor locking protocols for partitioned fixed-priority scheduling (such as the MPCP~\cite{LNR:09,Ra:90} or the FMLP~\cite{BLBA:07,BA:08}).
+While it is certainly tempting to apply period enforcement with the intention of avoiding the negative effects of deferred execution due to lock contention (as previously suggested elsewhere~\cite{Raj:91,Lak:11,LNR:09}), we ask: \emph{does existing blocking analysis remain safe when combined with the period enforcer algorithm?} In Section \ref{sec:locking}, we show that this is not the case.
+
 \begin{enumerate}
 \setcounter{enumi}{2}
-	\item The period enforcer algorithm invalidates all existing blocking analyses as there exist non-trivial feedback cycles between the period enforcer rules and blocking durations.
+	\item The period enforcer algorithm invalidates all existing blocking analyses for shared-memory real-time semaphore protocols as there exist non-trivial feedback cycles between the period enforcer rules and blocking durations.
 \end{enumerate}

### 2016-06-06 untitled.tex
edit section 3

@@
 
-Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to solve the task-set transformation problem: \emph{given a set of sporadic segmented self-suspending tasks, how do we obtain a corresponding set of (single-segment) deferrable tasks?} That is, as discussed in Section~\ref{sec:classic-analysis} the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert multi-segmented self-suspending tasks into corresponding sets of single-segment deferrable tasks, but it remains unclear in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished.
+Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the task-set transformation question: \emph{given a set of sporadic segmented self-suspending tasks $\tau$, how do we obtain a corresponding set of single-segment deferrable tasks $\tau'$ such that $\tau'$ is schedulable  only if $\tau$ is schedulable?} That is, as discussed in Section~\ref{sec:classic-analysis} the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert multi-segmented self-suspending tasks into corresponding sets of single-segment deferrable tasks, but it remains unclear in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished.
 In Section \ref{sec:convert}, we make a pertinent observation.
@@
 
-In this section, we demonstrate with an example that there exist sporadic  segmented self-suspending task sets that both \textbf{(i)} are schedulable \emph{without} period enforcement and \textbf{(ii)} are not schedulable with period enforcement.
+In this section, we demonstrate with an example that there exist sets of sporadic  segmented self-suspending tasks that both \textbf{(i)} are schedulable \emph{without} period enforcement and \textbf{(ii)} are not schedulable with period enforcement.
 
@@
 In contrast to the first job, the second job of task $\tau_2$ (which is released at time $11$) is affected by period enforcement.
-The first segment of the second job arrives at time $a^1_{2,2} = 11$, is interfered for one time unit, and suspends at time $13$.
+The first segment of the second job arrives at time $a^1_{2,2} = 11$, incurs interference for one time unit during $[11, 12)$, and suspends at time $13$.
 The  second segment of the second job hence resumes only at time $a^2_{2,2} = 19$.
@@
 
-One may consider to enrich the period enforcer with the following property: when the processor becomes idle, a task immediately becomes eligible to execute regardless of its eligibility time.
-However, even with this extension, the above example remains valid by introducing one additional lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from time $3$ to time $9$ and time $13$ to time $20$) and $T_3=D_3=100$.
+One may consider to enrich the period enforcer with the following scheduling rule: when the processor becomes idle, a task immediately becomes eligible to execute regardless of its eligibility time.
+However, even with this extension, the above example remains valid by introducing one additional lower-priority task $\tau_3$ with execution time $C_3=13$ (to be executed from time $3$ to time $9$ and time $13$ to time $20$) and $T_3=D_3=100$.
 With task $\tau_3$, the processor is always busy from time $0$ to time $23$ and consequently $\tau_2$ still misses its deadline at time $22$.
@@
 Furthermore, the example also demonstrates that the conversion to single-segment deferrable tasks does incur a loss of generality since it introduces pessimism.
-Recall that the analysis in Theorem 5 in~\cite{Raj:suspension1991} was only compatible with the deferrable task model.
-For the above example, we need to convert the segmented-suspending sporadic task $\tau_2$ into two deferrable tasks, called $\tau_2^1$ and $\tau_2^2$.
-By converting the two computation segments of task $\tau_2$ into two deferrable tasks $\tau_2^1$ and $\tau_2^2$, we can conclude that  task $\tau_2^1$ never defers its execution and task $\tau_2^2$ defers its execution by at most \emph{$9$} time units.
-Then the resulting single-segment deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable under the given fixed-priority scheduling since we can time the release of a job of $\tau_1$ to coincide with the arrival of a job of $\tau_2^2$ after it has maximally deferred its execution.
-We can easily see that the worst-case response time of the computation segment represented by task $\tau_2^2$ is $3$ time units after the second computation segment of task $\tau_2$ arrives.
-Together with the maximum deferred time $9$ time units, we know that the corresponding deferrable task set in fact has deadline misses since $9+3=12 > 11$.
-Therefore, the period enforcer algorithm may also cause a deadline miss.
+In the context of the above example, if we convert the multi-segmented suspending task $\tau_2$ into two single-segment deferrable tasks, called $\tau_2^1$ and $\tau_2^2$, where task $\tau_2^1$ never defers its execution and task $\tau_2^2$ defers its execution by at most \emph{$9$} time units, the resulting single-segment deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable under the given priority assignment: if a job of $\tau_1$ coincides with the arrival of a job of $\tau_2^2$ after it has maximally deferred its execution, the job of $\tau_2^2$ has a response time of $9 + 2 + 1$ time units, which exceeds its relative deadline of 11 time units.
+This shows that any restriction to single-segment deferrable tasks (i.e., assuming that ``a task $\tau_i$ can defer its entire execution time but not parts of it''~\cite{Raj:suspension1991}, recall Section~\ref{sec:classic-analysis}) inherently comes with a loss of generality.
 

### 2016-06-06 transformation.tex
edit section 3
edit section 4.1

@@
-\section{Deriving a Corresponding Deferrable Task Set and Schedulability Test}
+\section{Deriving and Analyzing a Corresponding Deferrable Task Set}
 \label{sec:convert}
 
-The example in Section~\ref{sec:unschedulable} can be easily converted to a corresponding deferrable task set, as explained at the end of Section~\ref{sec:unschedulable}.
-Due to Theorem 5 in~\cite{Raj:suspension1991}, the feasibility of the schedule by the period enforcer algorithm depends on whether the corresponding deferrable task set can be feasibly scheduled.
-Therefore, before applying the period enforcer algorithm to handle segmented self-suspending sporadic tasks, we need to first derive the corresponding deferrable task set and perform the schedulability test for the deferrable task set.
-To reduce the pessimism of the transformation to the corresponding deferrable task set, it would be the best to have a precise transformation and an exact schedulability test.
-Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.
-
-With the above discussions, we need to answer two questions before adopting the period enforcer algorithm.
-First, how can we derive the corresponding deferrable task set efficiently?
-Second, how should we perform the schedulability test on the corresponding deferrable task set correctly?
-We will discuss the difficulty to derive the corresponding deferrable task set in Section~\ref{sec:transformation-exponential} and the potential pessimism for the schedulability test based on the deferrable task set in Section~\ref{sec:schedulability-test-deferrable}.
+To apply an analysis of the period enforcer based on Theorem 5 in~\cite{Raj:suspension1991}, we need to convert a given set of multi-segment self-suspending tasks into a corresponding set of single-segment deferrable tasks.
+This raises two questions: first, how can we efficiently derive the corresponding set of single-segment deferrable tasks?
+And second, how should we perform the schedulability test on the corresponding single-segment deferrable tasks correctly?
+We will discuss the difficulty of deriving a suitable corresponding task set in Section~\ref{sec:transformation-exponential} and the potential for pessimism in any schedulability test based on the set of single-segment deferrable tasks  in Section~\ref{sec:schedulability-test-deferrable}.
 
@@
 
-The transformation from a segmented self-suspending task set to a corresponding deferrable task set was not discussed in \cite{Raj:suspension1991}.
-However, in our opinion, in general, such a transformation is not an easy problem.
-We demonstrate the inherent difficulty by focusing on a special case and by applying the recent result provided by Nelissen et al.
-\cite{ecrts15nelissen}, which analyzed the exact worst-case response time for segmented self-suspending sporadic tasks with exponential time complexity.
-The worst-case response time analysis in \cite{ecrts15nelissen} is exact under the following conditions: 1) the task set has only one segmented self-suspending task, 2) the self-suspending task is the lowest-priority task, 3) the scheduling policy is preemptive fixed-priority scheduling, and 4) $D_i \leq T_i$ for all task $\tau_i$ in the task system.\footnote{Here, we refer to the characteristics of the worst-case release pattern provided in Lemma 2 in \cite{ecrts15nelissen}.
-The exact worst-case response time is to explore all the release patterns that satisfy the conditions in Lemma 2 in \cite{ecrts15nelissen} instead of the proposed MILP in \cite{ecrts15nelissen} (which only provides an upper bound).
-}
+The needed transformation from a multi-segmented self-suspending task to a corresponding set of single-segment deferrable tasks was not discussed in \cite{Raj:suspension1991}.
+However, in our opinion, performing such a transformation without introducing additional pessimism is not easy in the general case, nor is it obvious how to do so.
 
-Suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$.
-The implicit deadline of a task is equal to its minimum inter-arrival time.
-Suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.
-Converting a computation segment into a deferrable task requires deriving the worst-case deferrable time (release jitter), denoted as $\rho_k^j$, for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$.
-Formally, if a job of task $\tau_k$ arrives at time $t$, it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than $t+\rho_k^j$.
-Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$.
-Therefore,  if we can derive the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ in this special case, we can clearly conclude that $\rho_k^1=0$ and $\rho_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k$.
+We demonstrate the inherent difficulty of the problem by focusing on a special case to which we can apply a recent result of Nelissen et al.
+\cite{ecrts15nelissen}, which allows analyzing the exact worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity.
+Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} is exact under the following conditions:\footnote{We refer to the characteristics of the worst-case release pattern provided in Lemma 2 in \cite{ecrts15nelissen}.
+The exact worst-case response time can be obtained by exploring all  release patterns that satisfy these conditions.}
+\begin{enumerate}
+	\item the task set contains only one self-suspending task, 
+	\item the self-suspending task is the lowest-priority task, 
+	\item the scheduling policy is preemptive fixed-priority scheduling, and 
+	\item all tasks have constrained deadlines.	
+\end{enumerate}
 
-Based on these considerations, it appears that, at least for the simple example, the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems.
-Deriving the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not an easy problem.
-The method recently provided by Nelissen et al.
-\cite{ecrts15nelissen} can be used for this specific case to derive the exact $W_k^j$ if we assume that task $\tau_k$ has only the first $j$ computation segments and $j-1$ self-suspension intervals.
-However, Nelissen et al.
-\cite{ecrts15nelissen} also showed that calculating the worst-case response time in the above ``simple'' case is already a very challenging problem, in which calculating $W_k^j$ would need exponential time complexity if $j \geq 2$.
-In particular, Nelissen et al.
-\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.\footnote{The schedulability test problem for such a case has been recently proved to be coNP-hard in the strong sense by Jian-Jia Chen in a report.}
+Suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines ($T_i = D_i$ for all $\tau_i$).
+Further suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.
+To  convert a computation segment into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible arrival time}, relative to the release of the job.
+Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than at time $t+\rho_k^j$.
 
-In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
-However, finding a safe upper bound of $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ can be done in pseudo-polynomial time \cite{PH:rtss98,Huang:multiseg}, if over-approximations can be tolerated.
+How can we compute $\rho_k^j$?
+Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$, and recall that $S_k^{j}$ denotes the maximum self-suspension length before the $j^{\mathrm{th}}$ computation segment of $\tau_k$.
+Then $\rho_k^j$ can be expressed in terms of $W_k^{j-1}$:
+$$
+	\rho_k^j = W_k^{j-1}+S_k^{j-1},
+$$
+where $W_k^0$.
+Therefore, if we can derive the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$, we can easily compute $\rho_k^j$  for $j=1,2,\ldots,m_k$.
+And conversely, if we can somehow compute $\rho_k^j$  for $j=2,\ldots,m_k$, we  can trivially infer $W_k^j$ for $j=1,2,\ldots,m_k-1$.
+Based on these considerations, it appears that the transformation problem is  --- at least in the considered special case --- equivalent to worst-case response time analysis of multi-segmented self-suspending task systems.
+
+However, deriving an exact bound $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not easy at all: 
+even for the above ``simple'' case, Nelissen et al.'s solution~\cite{ecrts15nelissen} for calculating the exact worst-case response time requires exponential time complexity if $j \geq 2$.
+Furthermore, Nelissen et al.
+\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem indeed.\footnote{In fact, in ongoing work, this problem has recently been shown to be coNP-hard in the strong sense~\cite{Chen2016b}.}
+
+In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the considered special case of a single self-suspending task).
+If over-approximations can be tolerated, finding a safe upper bound on $W_k^j$ for $j=1,2,\ldots,m_k-1$ can be done in pseudo-polynomial time \cite{PH:rtss98,Huang:multiseg}, albeit at a loss of accuracy and hence generality.
 

### 2016-06-06 transformation.tex
snapshot editing (not yet finished with section 4)

@@
 
-In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the considered special case of a single self-suspending task).
+In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case of a single self-suspending task).
 If over-approximations can be tolerated, finding a safe upper bound on $W_k^j$ for $j=1,2,\ldots,m_k-1$ can be done in pseudo-polynomial time \cite{PH:rtss98,Huang:multiseg}, albeit at a loss of accuracy and hence generality.
+
+Next, we consider another problem that arises once a suitable set of single-segment deferrable tasks has been found.
 
@@
 
-Suppose that the corresponding deferrable task set can be derived precisely by ignoring the complexity issues raised in Section~\ref{sec:transformation-exponential}.
-The next question is whether there is any information loss, with respect to the schedulability test, by inspecting only the corresponding deferrable task set.
-We will demonstrate an example to explain the information loss.
+Ignoring the complexity issues raised in Section~\ref{sec:transformation-exponential}, suppose that we have an oracle for precisely deriving the corresponding single-segment deferrable task set $\tset'$ for any given set of multi-segmented self-suspending tasks $\tset$.
+The next question is then how to analyze $\tset'$.
 
-Consider a task system consisting of $2$ tasks.
-Let $\tau_1$ denote a sporadic task without self-suspensions, in which $C_1 = 2$ and $T_1=D_1=10$.
-Let $\tau_2$ denote a self-suspending task consisting of three computation segments with parameters  $C_2^1 = 1$,  $S_2^1 = 6$, $C_2^2=1$, $S_2^2=8$, $C_2^3=1$, and $ T_2=D_2=21$.
-That is, this task set is pretty similar to the task set used in Section~\ref{sec:unschedulable}.
-One additional computation
-segment is created to illustrate the potential concerns.
-This task set is schedulable by fixed-priority preemptive scheduling without any enforcement since the worst-case response time of the second computation segment is $10$ and the third computation segment will suffer from the interference by at most one job from task $\tau_1$.
-Moreover, by using the analysis in Section~\ref{sec:transformation-exponential}, we can conclude that
-\begin{equation*}
-  \rho_2^1 = 0, \rho_2^2 = 9, \rho_2^3 = 18.
-\end{equation*}
-Therefore, in the corresponding deferrable task set, there are $2$ sporadic tasks $\tau_1$ and $\tau_2^1$ (with execution time $1$ and period $21$), and two deferrable tasks $\tau_2^2$ (with deferrable length $9$, period $21$, and execution time $1$) and $\tau_2^3$ (with deferrable length $18$, period $21$, and execution time $1$).
+Recall from Section~\ref{sec:classic-analysis} that a set single-segment deferrable tasks (i.e., a set of sporadic tasks subject to release jitter) can be analyzed with Audsley et al.'s response-time analysis~\cite{ABRTW:93}.
+However, when analyzing the response time of a single-segment deferrable task that corresponds to the  $j^{\mathrm{th}}$ computation segment of a multi-segmented self-suspending task $\tau_k$, it seems needlessly pessimistic to take into account interference from other single-segment deferrable tasks \emph{that also correspond to segments of $\tau_k$} --- after all, they reflect the same \emph{sequential} task, and  the execution of earlier segments of the same job is already reflected in $\rho_k^j$ .
+When analyzing a single-segment deferrable task in $\tset'$, one may thus be tempted to disregard any tasks stemming from the same multi-segmented self-suspending task (in $\tset$).
+We show with an example that this is generally not safe.
 
-Suppose that we are interested to analyze the worst-case response time of the deferrable task $\tau_2^3$ (in the corresponding deferrable task set) under fixed-priority scheduling.
-The question is \emph{whether we should ignore or should include the interference due to $\tau_2^1$ and $\tau_2^2$}.
+Consider a task system $\tset$ consisting of two tasks.
+Let $\tau_1$ denote a sporadic task without self-suspensions, with parameters $C_1 = 2$ and $T_1=D_1=10$.
+Further, let $\tau_2$ denote a self-suspending task consisting of three computation segments with parameters $S_2^0 = 0$, $C_2^1 = 1$,  $S_2^1 = 6$, $C_2^2=1$, $S_2^2=8$, $C_2^3=1$, and $ T_2=D_2=21$.
+This task set is similar to the task set used in Section~\ref{sec:unschedulable}; one additional computation segment has been added to illustrate the potential concerns.
+
+By design, this task set is schedulable under fixed-priority preemptive scheduling without any enforcement since the worst-case response time of $\tau_2$'s second computation segment is $10$ and the third computation segment will suffer interference due to at most one job of task $\tau_1$.
+
+Moreover, by using Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} and the relationship discussed in Section~\ref{sec:transformation-exponential}, we can conclude that
+$\rho_2^1 = 0$, $\rho_2^2 = 9$, and $\rho_2^3 = 18$.
+Therefore, in the corresponding task set $\tset'$, there are two sporadic tasks $\tau_1$ (with  $C_1 = 2$ and $T_1=D_1=10$) and $\tau_2^1$ (with  $C_2^1 = 1$ and $T^1_2=D^1_2=21$), and two single-segment deferrable tasks $\tau_2^2$ (with  $C_2^2 = 1$, $T^2_2=D^2_2=21$, and $S_2^{0,2} = \rho_2^2 = 9$) and $\tau_2^3$ (with  $C_2^3 = 1$, $T^3_2=D^3_2=21$, and $S_2^{0,3} = \rho_2^3 = 18$).
+
+Suppose that we seek to analyze the worst-case response time of the deferrable task $\tau_2^3 \in \tset'$.
+The question is: \emph{can we safely ignore the interference due to $\tau_2^1$ and $\tau_2^2$}, as they reflect computation of the same sequential task $\tau_2 \in \tset$?
+
+XXX stopped editing here XXX
+
 Since the computation segments that are used to create $\tau_2^1, \tau_2^2$, and $\tau_2^3$ do not have any overlap in their available time to be executed, it may be seemingly correct at beginning that we do not have to consider the interference from $\tau_2^1$ and $\tau_2^2$ when analyzing the response time of the deferrable task $\tau_2^3$.

### 2016-06-06 untitled.tex
snapshot editing (not yet finished with section 4)

@@
 
+\newcommand{\tset}[0]{\mathcal{T}}
+
 Finally, Rajkumar's original analysis~\cite{Raj:suspension1991} of the period enforcer is rooted in the concept of a \emph{task set transformation}.
-In general, such a task set transformation is simply a function $f$ that maps a given task set $\tau$ to a transformed task set $\tau' = f(\tau)$ \emph{such that $\tau'$ is schedulable \textbf{only if} the original task set $\tau$ is schedulable}, too.
-The basic idea is that such a transformation allows a form of reduction: $\tau$ can now \emph{indirectly} be shown to be schedulable by establishing that $\tau'$ is schedulable.
+In general, such a task set transformation is simply a function $f$ that maps a given task set $\tset$ to a transformed task set $\tset' = f(\tset)$ \emph{such that $\tset'$ is schedulable \textbf{only if} the original task set $\tset$ is schedulable}, too.
+The basic idea is that such a transformation allows a form of reduction: $\tset$ can now \emph{indirectly} be shown to be schedulable by establishing that $\tset'$ is schedulable.
 
-Importantly, the tasks in $\tau$ and $\tau'$ do \emph{not} have to be of the same task model, nor does the number of tasks have to remain constant (i.e.,  $|\tau| \neq |\tau'|$ is possible).
-Specifically, the task set transformation underlying the analysis of the period enforcer maps each \emph{multi-}segmented self-suspending task $\tau_i \in \tau$  to $m_i$ \emph{single-}segmented self-suspending tasks in $\tau'$ (i.e., $|\tau'| = \sum_{\tau} m_i$ ).
+Importantly, the tasks in $\tset$ and $\tset'$ do \emph{not} have to be of the same task model, nor does the number of tasks have to remain constant (i.e.,  $|\tset| \neq |\tset'|$ is possible).
+Specifically, the task set transformation underlying the analysis of the period enforcer maps each \emph{multi-}segmented self-suspending task $\tau_i \in \tset$  to $m_i$ \emph{single-}segmented self-suspending tasks in $\tset'$ (i.e., $|\tset'| = \sum_{\tset} m_i$).
 
@@
 
-Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the task-set transformation question: \emph{given a set of sporadic segmented self-suspending tasks $\tau$, how do we obtain a corresponding set of single-segment deferrable tasks $\tau'$ such that $\tau'$ is schedulable  only if $\tau$ is schedulable?} That is, as discussed in Section~\ref{sec:classic-analysis} the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert multi-segmented self-suspending tasks into corresponding sets of single-segment deferrable tasks, but it remains unclear in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished.
+Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the task-set transformation question: \emph{given a set of sporadic segmented self-suspending tasks $\tset$, how do we obtain a corresponding set of single-segment deferrable tasks $\tset'$ such that $\tset'$ is schedulable  only if $\tset$ is schedulable?} That is, as discussed in Section~\ref{sec:classic-analysis} the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert multi-segmented self-suspending tasks into corresponding sets of single-segment deferrable tasks, but it remains unclear in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished.
 In Section \ref{sec:convert}, we make a pertinent observation.

### 2016-06-07 locking.tex
edit section 4

@@
 
-\emph{Binary semaphores}, i.e., suspension-based locks used to realize mutually exclusive access to shared resources, are a common source of self-suspensions in multiprocessor real-time systems.
+Binary semaphores, i.e., suspension-based locks used to realize mutually exclusive access to shared resources, are a common source of self-suspensions in multiprocessor real-time systems.
 When a task tries to use a resource that has already been locked, it self-suspends until the resource becomes available.

### 2016-06-07 transformation.tex
edit section 4

--- 2016-06-06 transformation.tex
@@
 
-To apply an analysis of the period enforcer based on Theorem 5 in~\cite{Raj:suspension1991}, we need to convert a given set of multi-segment self-suspending tasks into a corresponding set of single-segment deferrable tasks.
-This raises two questions: first, how can we efficiently derive the corresponding set of single-segment deferrable tasks?
-And second, how should we perform the schedulability test on the corresponding single-segment deferrable tasks correctly?
-We will discuss the difficulty of deriving a suitable corresponding task set in Section~\ref{sec:transformation-exponential} and the potential for pessimism in any schedulability test based on the set of single-segment deferrable tasks  in Section~\ref{sec:schedulability-test-deferrable}.
+To apply an analysis of the period enforcer based on Theorem 5 in~\cite{Raj:suspension1991}, we first need to convert a given set of multi-segment self-suspending tasks into a corresponding set of single-segment deferrable tasks.
+This raises the question: how can we efficiently derive the corresponding set of single-segment deferrable tasks?
 
-\subsection{Task Set Transformation}
-\label{sec:transformation-exponential}
+The original period enforcer proposal~\cite{Raj:suspension1991} is silent on this issue and does not spell out a procedure for converting  multi-segmented self-suspending task to a corresponding set of single-segment deferrable tasks.
+However, in our opinion, performing such a transformation without introducing additional pessimism is not at all easy in the general case.
 
-The needed transformation from a multi-segmented self-suspending task to a corresponding set of single-segment deferrable tasks was not discussed in \cite{Raj:suspension1991}.
-However, in our opinion, performing such a transformation without introducing additional pessimism is not easy in the general case, nor is it obvious how to do so.
-
-We demonstrate the inherent difficulty of the problem by focusing on a special case to which we can apply a recent result of Nelissen et al.
+In the following, we illustrate the inherent difficulty of the problem by focusing on a special case to which we can apply a recent result of Nelissen et al.
 \cite{ecrts15nelissen}, which allows analyzing the exact worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity.
@@
 	\item the scheduling policy is preemptive fixed-priority scheduling, and 
-	\item all tasks have constrained deadlines.	
+	\item all tasks have constrained deadlines (i.e., $D_i \leq T_i$ for all $\tau_i$).
 \end{enumerate}
 
-Suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines ($T_i = D_i$ for all $\tau_i$).
+For an arbitrary number of tasks $k \geq 2$, 
+suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines (i.e., $D_i = T_i$ for all $\tau_i$).
 Further suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.
-To  convert a computation segment into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible arrival time}, relative to the release of the job.
+
+To  convert a computation segment of $\tau_k$ into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible arrival time}, relative to the release of a job.
 Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than at time $t+\rho_k^j$.
@@
 $$
-where $W_k^0$.
-Therefore, if we can derive the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$, we can easily compute $\rho_k^j$  for $j=1,2,\ldots,m_k$.
-And conversely, if we can somehow compute $\rho_k^j$  for $j=2,\ldots,m_k$, we  can trivially infer $W_k^j$ for $j=1,2,\ldots,m_k-1$.
-Based on these considerations, it appears that the transformation problem is  --- at least in the considered special case --- equivalent to worst-case response time analysis of multi-segmented self-suspending task systems.
+where $W_k^0 = 0$.
+Therefore, if we can derive the exact segment worst-case response time $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$, we can easily compute $\rho_k^j$  for $j=1,2,\ldots,m_k$.
+And conversely, if we can somehow obtain $\rho_k^j$  for $j=2,\ldots,m_k$, we  can trivially infer $W_k^j$ for $j=1,2,\ldots,m_k-1$.
+Based on these considerations, it appears that the transformation problem is  --- at least in the considered special case --- equivalent to the  worst-case response time analysis of a multi-segmented self-suspending task.
 
-However, deriving an exact bound $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not easy at all: 
+However, deriving an exact bound $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not easy: 
 even for the above ``simple'' case, Nelissen et al.'s solution~\cite{ecrts15nelissen} for calculating the exact worst-case response time requires exponential time complexity if $j \geq 2$.
 Furthermore, Nelissen et al.
-\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem indeed.\footnote{In fact, in ongoing work, this problem has recently been shown to be coNP-hard in the strong sense~\cite{Chen2016b}.}
+\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.\footnote{In fact, in ongoing work, this problem has recently been shown to be coNP-hard in the strong sense~\cite{Chen2016b}.}
 
-In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the considered special case of a single self-suspending task).
-If over-approximations can be tolerated, finding a safe upper bound on $W_k^j$ for $j=1,2,\ldots,m_k-1$ can be done in pseudo-polynomial time \cite{PH:rtss98,Huang:multiseg}, albeit at a loss of accuracy and hence generality.
+Nelissen et al.
+\cite{ecrts15nelissen} did not study the period enforcer; rather, they considered unrestricted self-suspensions.
+However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that these observations apply to the period enforcement case, too.
 
-\subsection{Schedulability Test for A Corresponding Deferrable Task Set}
-\label{sec:schedulability-test-deferrable}
+To summarize, to analyze the period enforcer based on Theorem~5 in~\cite{Raj:suspension1991}, a procedure for transforming multi-segmented self-suspending tasks into sets of single-segment deferrable tasks is needed, but no such procedure was given in the original proposal~\cite{Raj:suspension1991}.
+Based on the presented considerations, we conclude that this missing step is non-trivial and observe that the closest known solution by Nelissen et al.
+\cite{ecrts15nelissen} requires exponential time even in the greatly simplified special case of a single self-suspending task.
+It thus remains unclear how Theorem~5 in~\cite{Raj:suspension1991} can be used in the schedulability analysis of sets of multi-segmented self-suspending tasks.
 
-Suppose that the corresponding deferrable task set can be derived precisely by ignoring the complexity issues raised in Section~\ref{sec:transformation-exponential}.
-The next question is whether there is any information loss, with respect to the schedulability test, by inspecting only the corresponding deferrable task set.
-We will demonstrate an example to explain the information loss.
+We further note that we did seek alternatives, but we did not find a simple or efficient schedulability test for the period enforcer without introducing additional pessimism.
 
-Consider a task system consisting of $2$ tasks.
-Let $\tau_1$ denote a sporadic task without self-suspensions, in which $C_1 = 2$ and $T_1=D_1=10$.
-Let $\tau_2$ denote a self-suspending task consisting of three computation segments with parameters  $C_2^1 = 1$,  $S_2^1 = 6$, $C_2^2=1$, $S_2^2=8$, $C_2^3=1$, and $ T_2=D_2=21$.
-That is, this task set is pretty similar to the task set used in Section~\ref{sec:unschedulable}.
-One additional computation
-segment is created to illustrate the potential concerns.
-This task set is schedulable by fixed-priority preemptive scheduling without any enforcement since the worst-case response time of the second computation segment is $10$ and the third computation segment will suffer from the interference by at most one job from task $\tau_1$.
-Moreover, by using the analysis in Section~\ref{sec:transformation-exponential}, we can conclude that
-\begin{equation*}
-  \rho_2^1 = 0, \rho_2^2 = 9, \rho_2^3 = 18.
-\end{equation*}
-Therefore, in the corresponding deferrable task set, there are $2$ sporadic tasks $\tau_1$ and $\tau_2^1$ (with execution time $1$ and period $21$), and two deferrable tasks $\tau_2^2$ (with deferrable length $9$, period $21$, and execution time $1$) and $\tau_2^3$ (with deferrable length $18$, period $21$, and execution time $1$).
-
-Suppose that we are interested to analyze the worst-case response time of the deferrable task $\tau_2^3$ (in the corresponding deferrable task set) under fixed-priority scheduling.
-The question is \emph{whether we should ignore or should include the interference due to $\tau_2^1$ and $\tau_2^2$}.
-Since the computation segments that are used to create $\tau_2^1, \tau_2^2$, and $\tau_2^3$ do not have any overlap in their available time to be executed, it may be seemingly correct at beginning that we do not have to consider the interference from $\tau_2^1$ and $\tau_2^2$ when analyzing the response time of the deferrable task $\tau_2^3$.
-However, this statement is in fact incorrect.
-
-If we do not consider the interference from $\tau_2^1$ and $\tau_2^2$, we can conclude that the worst-case response time of $\tau_2^3$ is  $18+C_1+C_2^3=21$.
-Let us consider this task set under the control of the period enforcer algorithm, as defined in Section \ref{sec:pe}.
-Figure \ref{fig:example-deferrable-incompatible} shows the resulting schedule for a periodic release pattern. This changes the schedule of the second job of task $\tau_2$. 
-The second job of task $\tau_2$ (which is released at time $21$) is affected by period enforcement.
-The first segment of the second job arrives at time $a^1_{2,2} = 21$, is interfered for one time unit, and suspends at time $23$.
-The  second segment of the second job hence resumes only at time $a^2_{2,2} = 29$.
-Thus, we have
-\begin{align*}
-	ET_{2,2}^1 & = \max\left(0 + 21,\ \mathit{busy}(\tau_2, 21)\right) = 21  \text{ and }
-\\
-	ET_{2,2}^2 & = \max\left(9 + 21,\ \mathit{busy}(\tau_2, 29\right) ) = 30.
-\end{align*}
-According to the rules of the period enforcer algorithm, the processor therefore remains idle at time $29$ because the segment is not eligible to execute until time $ET_{2,2}^2 = 30$.
-However, at time $30$, the fourth job of $\tau_1$ is released.
-As a result, the second job of $\tau_2$ suffers from additional interference and misses its deadline at time $42$.
-
-If we further allow task $\tau_1$ to be released sporadically, the fifth job of task $\tau_1$ can be released at time $41$ instead of time $40$ in Figure~\ref{fig:example-deferrable-incompatible}.
-For such a case, the response time of the second job of task $\tau_2$ is $23$.
-This means that the worst-case response time of the deferrable task$\tau_2^3$ must be at least $23$, which equals to $18+C_1 + C_2^1+C_2^2+C_2^3$.
-
-Therefore, for this specific example, we know that the schedulability test for the  corresponding deferrable task set has also to be pessimistic enough to consider the interferences from all the computation segments created from the same self-suspending sporadic task.
-
-\begin{figure}[t]
-  \centering  
-  \includegraphics[scale=1]{../figures/deferrable-incompatible/deferrable-incompatible.pdf}
-  \caption{An example for explaining why it is not safe to ignore the interference from the other computation segments in the same task.}
-  \label{fig:example-deferrable-incompatible}
-\end{figure}
-
-\subsection{Discussions}
-\label{sec:discussions-deferrable}
-
-The transformation from a segmented self-suspending task set to a corresponding deferrable task set and the corresponding schedulability test were not discussed in \cite{Raj:suspension1991}.
-With the above discussions in Section~\ref{sec:transformation-exponential} and Section~\ref{sec:schedulability-test-deferrable}, we find that the transformation is in fact non-trivial and the schedulability test of the corresponding deferrable task set has also to be pessimistic.
-As shown in Section~\ref{sec:schedulability-test-deferrable}, even if we derive the precise deferrable time (release jitter) information, we may still have to account for the computation segments that have no \emph{direct} interference with a computation segment that is under analysis (e.g., the third computation segment of task $\tau_2$ in Section~\ref{sec:schedulability-test-deferrable}).
-
-Fundamentally, the design of the period enforcer algorithm relies on the schedulability of a deferrable task set.
-However, as we demonstrated here, there is a gap between the segmented self-suspending task set and the corresponding deferrable task set in the transformation and also in the schedulability test analysis.
-The authors in this paper do not find any simple way to provide a schedulability test without any information loss.
+Next, we take a look at the period enforcer in the context of lock contention.
 

### 2016-06-07 conclusion.tex
more fixes to address Jim's comments
- mention FP and partitioning early on
- spell out small numbers

Also add acknowledgements.
shorten acknowledgement
ack
add ack sfb876
snapshot editing (not yet finished with section 4)
add remarks on DPCP/DFLP and some minor edits

@@
 
-Furthermore, the original proposal \cite{Raj:suspension1991} left open the question of how to convert a segmented self-suspending task set to a corresponding set of deferrable tasks.
-Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a task set transformation is non-trivial in the general case.
+The original proposal \cite{Raj:suspension1991} further left open the question of how to convert a segmented self-suspending task set to a corresponding set of deferrable tasks.
+Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a transformation is non-trivial in the general case.
 
@@
 However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.
+
+\section*{Acknowledgements}
+
+We thank James H.\ Anderson and Raj Rajkumar for their comments on early drafts of this paper.
+This work has been supported by DFG, as part of the Collaborative
+Research Center SFB876. % (\url{http://sfb876.tu-dortmund.de/}).

### 2016-06-07 locking.tex
add remarks on DPCP/DFLP and some minor edits

@@
 \subsection{Case 1: Locking Takes Effect at Earliest Segment Eligibility Time}
+\label{sec:sem-case-1}
 In the following example, we assume the first interpretation, i.e., that the processing of lock requests is delayed until the point when a resuming segment would no longer be subject to any delay due to period enforcement.
@@
 \subsection{Case 2: Locking Takes Effect Immediately}
+\label{sec:sem-case-2}
  From now on, we assume the second interpretation: all lock requests are processed immediately when they are made, even if this causes the shared resource to be locked by a task that is not yet eligible to execute according to  the rules of the period enforcer algorithm.
@@
 
+Notably, the examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a shared-memory locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors.
+One may wonder whether  effects similar to those described in in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the DPCP~\cite{RSL:88} or the DFLP~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}.
+In this case, the self-suspension occurs on the task's \emph{application processor}, which is different from the (remote) synchronization processor on which the critical section is executed.
+This separation allows employing period enforcement  only on application processors (while avoiding it on synchronization processors) without incurring the feedback cycle between blocking times and self-suspension times highlighted in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2}.
+However, period enforcement still invalidates all existing blocking analyses for distributed real-time semaphore protocols~\cite{RSL:88,Raj:91,Br:13} because it artificially increases blocking times if tasks contain multiple accesses to shared resources.
+
 Finally, it is worth noting that our examples can be trivially extended with lower-priority tasks to ensure that no processor idles before the described deadline misses occur.

### 2016-06-07 untitled.tex
add remarks on DPCP/DFLP and some minor edits

@@
 
-On uniprocessors, this relaxation does not introduce additional pessimism: since any two jobs of a sporadic task $\tau_i$ are  known to be released \emph{at least} $T_i$ time units apart, the sporadic task model~\cite{Mo:83} still allows for schedulability analysis that is as accurate as Liu and Layland's analysis of periodic tasks~\cite{LL:73}.
+On uniprocessors, the relaxation from periodic to sporadic job arrivals does not introduce additional pessimism:\footnote{Assuming that all periodic tasks release a job synchronously at time zero.} since any two jobs of a sporadic task $\tau_i$ are  known to be released \emph{at least} $T_i$ time units apart, the sporadic task model~\cite{Mo:83} still allows for schedulability analysis that is as accurate as Liu and Layland's analysis of periodic tasks~\cite{LL:73}.
 
@@
 \setcounter{enumi}{2}
-	\item The period enforcer algorithm invalidates all existing blocking analyses for shared-memory real-time semaphore protocols as there exist non-trivial feedback cycles between the period enforcer rules and blocking durations.
+	\item The period enforcer algorithm invalidates all existing blocking analyses for real-time semaphore protocols as there exist non-trivial feedback cycles between the period enforcer rules and blocking durations.
 \end{enumerate}

### 2016-06-08 locking.tex
proofreading and polishing pass

@@
 Such self-suspensions due to lock contention, just like any other self-suspension, result in deferred execution and thus can detrimentally affect a task's interference on lower-priority tasks.
-It may thus seem natural to apply the period enforcer  to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91} and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.
-As we show in this paper, this is unsafe.} However, as we demonstrate with two examples, it is actually not safe to apply period enforcement to lock-induced self-suspensions.
+It may thus seem natural to apply the period enforcer  to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91}, stated as a motivation and possible use case in the original period enforcer proposal~\cite{Raj:suspension1991}, and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.
+As we show here, this is actually unsafe.} However, as we demonstrate with two examples, it is actually unsafe to apply period enforcement to lock-induced self-suspensions.
 

### 2016-06-08 transformation.tex
proofreading and polishing pass

@@
-\section{Deriving and Analyzing a Corresponding Deferrable Task Set}
+\section{Deriving a Corresponding Deferrable Task Set}
 \label{sec:convert}
@@
 where $W_k^0 = 0$.
-Therefore, if we can derive the exact segment worst-case response time $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$, we can easily compute $\rho_k^j$  for $j=1,2,\ldots,m_k$.
+Therefore, if we can derive the exact segment worst-case response time $W_k^j$ for $j=1,2,\ldots,m_k-1$, we can easily compute $\rho_k^j$  for $j=1,2,\ldots,m_k$.
 And conversely, if we can somehow obtain $\rho_k^j$  for $j=2,\ldots,m_k$, we  can trivially infer $W_k^j$ for $j=1,2,\ldots,m_k-1$.
@@
 
-Nelissen et al.
-\cite{ecrts15nelissen} did not study the period enforcer; rather, they considered unrestricted self-suspensions.
-However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that these observations apply to the period enforcement case, too.
+Nelissen et al.~\cite{ecrts15nelissen} did not study the period enforcer; rather, they considered unrestricted self-suspensions.
+However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that these observations transfer to the period enforcement case.
 
-To summarize, to analyze the period enforcer based on Theorem~5 in~\cite{Raj:suspension1991}, a procedure for transforming multi-segmented self-suspending tasks into sets of single-segment deferrable tasks is needed, but no such procedure was given in the original proposal~\cite{Raj:suspension1991}.
-Based on the presented considerations, we conclude that this missing step is non-trivial and observe that the closest known solution by Nelissen et al.
+To summarize, to analyze the period enforcer based on Theorem~5 in~\cite{Raj:suspension1991}, a procedure for transforming multi-segmented self-suspending tasks into sets of single-segment deferrable tasks is needed, but no such procedure is given in the original proposal~\cite{Raj:suspension1991}.
+%
+Based on the presented considerations, we conclude that filling in this missing step is non-trivial and observe that the closest known solution by Nelissen et al.
 \cite{ecrts15nelissen} requires exponential time even in the greatly simplified special case of a single self-suspending task.
-It thus remains unclear how Theorem~5 in~\cite{Raj:suspension1991} can be used in the schedulability analysis of sets of multi-segmented self-suspending tasks.
+It thus remains unclear how Theorem~5 in~\cite{Raj:suspension1991} can be used for schedulability analysis of sets of multi-segmented self-suspending tasks.
+While we did search for alternative analysis approaches that do not rely on Theorem 5, we did not find a simple or efficient schedulability test for the period enforcer without introducing substantial additional pessimism.
+The problem remains open.
 
-We further note that we did seek alternatives, but we did not find a simple or efficient schedulability test for the period enforcer without introducing additional pessimism.
-
-Next, we take a look at the period enforcer in the context of lock contention.
+Next, we take a look at the period enforcer in the context of synchronization protocols.
 

### 2016-06-08 untitled.tex
proofreading and polishing pass

@@
 First, in the words of Rajkumar, it ``forces tasks to behave like ideal periodic tasks from the scheduling point of view with no associated scheduling penalties''~\cite{Raj:suspension1991}, which is obviously highly desirable in many practical applications in which self-suspensions are inevitable (e.g., when offloading computations to co-processors such as GPUs or DSPs).
-Second, the later-proposed, but more widely-known \emph{released guard} algorithm~\cite{SL:96} uses a technique quite similar to period enforcement to control scheduling penalties due to release jitter in distributed systems.
+Second, the later-proposed, but more widely-known \emph{release guard} algorithm~\cite{SL:96} uses a technique quite similar to period enforcement to control scheduling penalties due to release jitter in distributed systems.
 The period enforcer algorithm has also attracted renewed attention in recent years and has been discussed in several current works  (e.g.,~\cite{DBLP:conf/rtss/ChenL14,LNR:09,LR:10,Lak:11,LC:14,KANR:13,HY:11,CA:09,CA:10,CA:10b}), at times controversially~\cite{BA:08a}.
@@
 
-We briefly introduce the needed background in Section~\ref{sec:prelim}, restate our contributions more precisely in Section~\ref{sec:questions}, and then establish the three above  observations in detail in Sections \ref{sec:unschedulable}--\ref{sec:locking} before concluding in Section \ref{sec:conclusion}.
+We introduce the needed background in Section~\ref{sec:prelim}, restate our contributions more precisely in Section~\ref{sec:questions}, and then establish the three above  observations in detail in Sections \ref{sec:unschedulable}--\ref{sec:locking} before concluding in Section \ref{sec:conclusion}.
 
@@
 Each job must finish by the time the next arrives.
-Importantly, Liu and Layland assumed both that the $k$\xth job of $\tau_i$ arrives \emph{exactly} at time $(k-1)\times T_i$, and that an incomplete job is \emph{always} available for execution (i.e., jobs never block on I/O or locks).
+Importantly, Liu and Layland assume both that the $k$\xth job of $\tau_i$ arrives \emph{exactly} at time $(k-1)\times T_i$, and that an incomplete job is \emph{always} available for execution (i.e., jobs never block on I/O or locks).
 
@@
 
-Mok~\cite{Mo:83} introduced the \emph{sporadic task model}, in which each task $\tau_i$ is still specified by a tuple $(C_i, T_i, D_i)$.
+Mok~\cite{Mo:83} introduced the \emph{sporadic task model}, a widely used generalization of the periodic task model in which each task $\tau_i$ is still specified by a tuple $(C_i, T_i, D_i)$.
 However, the sporadic task model relaxes the inter-arrival constraint $T_i$ to specify a \emph{minimum} (rather than an exact) separation between jobs.
@@
 
-On uniprocessors, the relaxation from periodic to sporadic job arrivals does not introduce additional pessimism:\footnote{Assuming that all periodic tasks release a job synchronously at time zero.} since any two jobs of a sporadic task $\tau_i$ are  known to be released \emph{at least} $T_i$ time units apart, the sporadic task model~\cite{Mo:83} still allows for schedulability analysis that is as accurate as Liu and Layland's analysis of periodic tasks~\cite{LL:73}.
+On uniprocessors, the relaxation from periodic to sporadic job arrivals does not introduce additional pessimism:\footnote{Assuming that all periodic tasks synchronously release a job at time zero.} since any two jobs of a sporadic task $\tau_i$ are  known to be released \emph{at least} $T_i$ time units apart, the sporadic task model~\cite{Mo:83} still allows for schedulability analysis that is as accurate as Liu and Layland's analysis of periodic tasks~\cite{LL:73}.
 
@@
 The latter assumption --- immediate availability for execution --- is inappropriate in many practical systems (especially in networked systems) if events (e.g., messages) that trigger job releases can incur non-negligible delays (e.g., network congestion).
-Such delays in task activation can accounted for by introducing a notion of \emph{release jitter}.
+Such delays in task activation can be accounted for by introducing a notion of \emph{release jitter}.
 To this end, each task is represented by a four-tuple $(C_i, J_i, T_i, D_i)$, where the parameter $J_i$ is a bound on the maximum time that a job remains unavailable for execution after it should have started to run.
@@
 In the presence of release jitter, the terms ``job arrival'' and ``job release,'' which are often used interchangeably, take on distinct meanings: a job's \emph{arrival time} denotes the point in time when it actually becomes available for execution, whereas a job's \emph{release time} is the instant that is relevant for the (minimum) inter-arrival time constraint.
-Any job of task $\tau_i$ thus arrives at most $J_i$ time units after it is released.
+Any job of task $\tau_i$  \emph{arrives} at most $J_i$ time units after it is \emph{released}.
 
 Notably, non-zero release jitter \emph{does} cause additional pessimism: in the worst case, two consecutive jobs of a task $\tau_i$ can be separated by as little as $T_i - J_i$ time units (if the earlier job incurs maximum release jitter and the successor job incurs none).
-As a result, a task may ``carry in'' some additional work into a given analysis interval.
+As a result, a task may ``carry in'' some additional work into a given interval.
 Taking this effect into account, Audsley et al.~\cite{ABRTW:93} developed a response-time analysis for sporadic and periodic constrained-deadline tasks subject to release jitter under preemptive fixed-priority scheduling.
@@
 However, even in the presence of release jitter, a key assumption remains that jobs do not self-suspend (e.g., wait for I/O).\footnote{Audsley et al.~\cite{ABRTW:93} do present a response-time analysis that takes into account a limited form of suspensions due to semaphores (``blocking'').
-However, their analysis does not apply to general self-suspensions (i.e., the kind of self-suspensions targeted by period enforcer algorithm)  and is not relevant in the context of this paper.} That is, it is assumed that, once a job has arrived, it continuously remains available for dispatching until it completes.
+However, their analysis does not apply to general self-suspensions (i.e., the kind of self-suspensions targeted by period enforcer algorithm)  and is not relevant in the context of this paper.} That is, Audsley et al.~\cite{ABRTW:93} assume that, once a job has arrived, it continuously remains available for dispatching until it completes.
 This restriction is removed next.
@@
 When a job \emph{self-suspends}, it becomes unavailable for execution until some external event occurs (e.g., a disk I/O operation completes, a network packet arrives, a co-processor signals completion, \textit{etc.}).
-This has the effect of \emph{deferring} (a part of) the job's processing requirement until the time that it \emph{resumes} from its suspension, which causes massive analytical difficulties~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen,Ri:04}.
+This has the effect of \emph{deferring} (a part of) the job's processing requirement until the time that it \emph{resumes} from its suspension, which causes massive analytical difficulties~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen,Ri:04,Raj:suspension1991,Chen2016}.
 
-To date, the real-time literature on self-suspensions has focused on two task models: the \emph{dynamic} self-suspension model, which we discuss first,  and the \emph{segmented} (or \emph{multi-segment}) suspension model, which we discuss next in Section~\ref{sec:segmented}.
-Self-suspensions can arise in both periodic and sporadic tasks (i.e, both interpretations of the $T_i$ parameter are possible); in this note we focus on sporadic tasks.
+To date, the real-time literature on self-suspensions has focused on two task models: the \emph{dynamic} self-suspension model, which we discuss first,  and the  (\emph{multi}-)\emph{segmented} suspension model, which we discuss next in Section~\ref{sec:segmented}.
+Self-suspensions can arise in both periodic and sporadic tasks (i.e, both interpretations of the $T_i$ parameter are possible).
+The observations that we make in this note apply equally to both periodic and sporadic tasks; for convenience, we focus primarily on periodic tasks.
 
@@
 The dynamic self-suspension model does not impose a bound on the maximum number of self-suspensions, nor does it make any assumptions as to where during a job's execution self-suspensions occur.
-That is, how often a job defers its execution, when it does so, and how much it defers varies unpredictably from job to job.
+That is, how often a job defers its execution, when it does so, and how much of its execution it defers may vary unpredictably from job to job.
 
@@
 In particular, release jitter and self-suspensions are not interchangeable concepts and it is not safe~\cite{Chen2016,ecrts15nelissen} to simply substitute $J_i$ with $S_i$ in Audsley et al.'s analysis~\cite{ABRTW:93}.
-However, under the dynamic suspension model, it is possible for jobs of self-suspending tasks to defer their entire execution requirement.
+(Nonetheless, under the dynamic suspension model, it is possible for jobs of self-suspending tasks to defer their entire execution requirement, so self-suspensions can be seen as a generalization of release jitter.)
 
 The period enforcer algorithm aims to mitigate the negative effects of self-suspensions.
-However, for reasons that will be explained in Section~\ref{sec:pe}, the period enforcer algorithm cannot be meaningfully combined with the dynamic suspension model.
-Instead, it requires a segmented suspension model, which we discuss next.
+However, for reasons that will be explained in Section~\ref{sec:incompat}, the period enforcer algorithm cannot be meaningfully combined with the dynamic suspension model.
+Instead, it requires the segmented suspension model, which we discuss next.
 
@@
 
-The first self-suspension segment $S_i^0$, prior to the first execution segment, is equivalent to release jitter ($J_i$, Section~\ref{sec:jitter}).
-However, in many works on the segmented self-suspending task model, the segment $S_i^0$ assumed to be absent (i.e., $S_i^0 = 0$), such that there are only $m_i - 1$ suspension intervals (and jobs arrive jitter-free).
+The first self-suspension segment $S_i^0$, prior to the first execution segment, is equivalent to release jitter (i.e., the parameter $J_i$ in Section~\ref{sec:jitter}).
+However, in in much of the literature on the segmented self-suspending task model, the segment $S_i^0$ assumed to be absent (i.e., $S_i^0 = 0$), such that there are only $m_i - 1$ suspension intervals (and jobs arrive jitter-free).
 Unless noted otherwise we adopt this convention.
@@
 
-The advantage of the dynamic model (Section~\ref{sec:dynamic}) is that it is more flexible since it does not impose any assumptions on the task control flow.
+The advantage of the dynamic model (Section~\ref{sec:dynamic}) is that it is more flexible since it does not impose any assumptions on a task's control flow.
 The advantage of the segmented model is that it allows for more accurate analysis.
-The period enforcer algorithm and its analysis~\cite{Raj:suspension1991} applies (only) to the segmented model, as explained in Section~\ref{sec:classic-analysis}.
+The period enforcer algorithm and its analysis~\cite{Raj:suspension1991} applies (only) to the segmented model, as explained in Sections~\ref{sec:incompat} and~\ref{sec:classic-analysis}.
 
@@
 \label{sec:single-segmented}
-
-Segmented self-suspending tasks with exactly one self-suspension interval followed by exactly one computation segment ($m_i = 1$, $S_i^0 \neq 0$) are an important special case, which we refer to as \emph{single-segment self-suspending tasks}.
+An important special case is segmented self-suspending tasks with exactly one self-suspension interval followed by exactly one computation segment ($m_i = 1$, $S_i^0 \neq 0$), which we refer to as \emph{single-segment self-suspending tasks}.
 This special case is central to Rajkumar's original analysis of the period enforcer~\cite{Raj:suspension1991},  as we will explain in Section~\ref{sec:classic-analysis}.
@@
 
-Note also that single-segment self-suspending sporadic tasks, which are  ``suspended'' only prior to commencing execution, are analytically fully equivalent to sporadic task subject to release jitter (i.e., the model described in Section~\ref{sec:jitter}).
-We nonetheless use the term ``single-segment self-suspending sporadic task'' to remain close to Rajkumar's original description~\cite{Raj:suspension1991}, and to highlight the connection to the (multi-)segmented self-suspending task model (Section~\ref{sec:segmented}).
+Note also that single-segment self-suspending sporadic tasks, which are  ``suspended'' only prior to commencing execution, are analytically fully equivalent to sporadic tasks subject to release jitter (i.e., the model described in Section~\ref{sec:jitter}).
+We nonetheless use the term ``single-segment self-suspending task,'' or interchangeably ``single-segment deferrable task,'' to remain close to Rajkumar's original description~\cite{Raj:suspension1991}, and to highlight the connection to the (multi-)segmented self-suspending task model (Section~\ref{sec:segmented}).
 
-This concludes our review of relevant task models. Before reviewing the period enforcer and its original analysis, we briefly introduce to some essential concepts.
+This concludes our review of relevant task models. Before reviewing the period enforcer and its original analysis, we briefly introduce some essential concepts.
 
@@
 In general, such a task set transformation is simply a function $f$ that maps a given task set $\tset$ to a transformed task set $\tset' = f(\tset)$ \emph{such that $\tset'$ is schedulable \textbf{only if} the original task set $\tset$ is schedulable}, too.
-The basic idea is that such a transformation allows a form of reduction: $\tset$ can now \emph{indirectly} be shown to be schedulable by establishing that $\tset'$ is schedulable.
+The basic idea is that such a transformation allows schedulability analysis by reduction: given a suitable transformation $f$, $\tset$ can be  \emph{indirectly} shown to be schedulable by computing $\tset' = f(\tset)$ and establishing that $\tset'$ is schedulable.
 
@@
 
-The period enforcer consists of two parts: a runtime rule that governs when a self-suspending task may be scheduled, and an (offline) analysis that may be used to assess the temporal correctness of a set of segmented self-suspending tasks (Section~\ref{sec:segmented}) subject to period enforcement.
+The period enforcer consists of two parts: a runtime rule that governs when each segment of a self-suspending task may be scheduled, and an (offline) analysis that may be used to assess the temporal correctness of a set of self-suspending tasks (Section~\ref{sec:segmented}) subject to period enforcement.
 Initially, we focus on the runtime rule (i.e., the actual period enforcer algorithm) and then review the corresponding original analysis thereafter in Section~\ref{sec:classic-analysis}.
@@
 
-The root cause is increased interference due to the ``back-to-back'' execution effect~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}, where two jobs of $\tau_2$ execute in close succession (i.e, separated by less than a period) because the second job self-suspended for a (much) shorter duration than the first job.
-In the example shown in Figure \ref{fig:not-ok-without-period-enforcement},  $\tau_3$ misses its deadline at time~15 because of this effect as $\tau_2$'s second job resumes ``too soon'' (i.e, already at time~12 after having been suspended for only one time unit, rather than four time units like the first job of $\tau_2$).
+The root cause is increased interference due to the ``back-to-back'' execution effect~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}.
+In the example shown in Figure \ref{fig:not-ok-without-period-enforcement}, two jobs of $\tau_2$ execute in close succession (i.e, separated by less than a period) because the second job, released at time~10, self-suspended for a (much) shorter duration than the first job.
+Consequently,  $\tau_3$ suffers from increased interference when $\tau_2$'s second job resumes ``too soon'' at time~12 after having been suspended for only one time unit, rather than four time units like the first job of $\tau_2$.
 
@@
 
-In short, the period enforcer fundamentally requires the segmented self-suspension model~(Section~\ref{sec:segmented}) to be effective because a job of a dynamic self-suspending task exhibits unpredictable execution times between its unpredictably many self-suspensions.
+In short, to be effective, the period enforcer fundamentally requires the segmented self-suspension model~(Section~\ref{sec:segmented})  because it cannot cope with the unpredictable execution times between (the unpredictably many) self-suspensions that jobs may exhibit under the dynamic self-suspension model.
 
@@
 In this example, the problem is that the eligibility time of the first computation ``segment'' of the second job is determined by the self-suspension pattern of the first job, even though the first job deferred all of its execution, whereas the second job deferred only a part of its execution.
-Under the more restrictive segmented self-suspension model~(Section~\ref{sec:segmented}), the pattern of self-suspension and computation times is statically fixed by the model; such a mismatch is hence not possible.
+Under the more restrictive segmented self-suspension model~(Section~\ref{sec:segmented}), the pattern of self-suspension and computation times is statically fixed; such a mismatch is hence not possible.
 
@@
 \begin{quote}
-With deferred execution, a task $\tau_i$ can execute its $C_i$ units of execution in discrete amounts $C_i^1, C_i^2$, $\ldots$ with suspension in between $C_i^j$ and $C_i^{j+1}$.
-\cite[Section 3]{Raj:suspension1991}\footnote{The notation has been altered here for the sake of consistency.
+``With deferred execution, a task $\tau_i$ can execute its $C_i$ units of execution in discrete amounts $C_i^1, C_i^2$, $\ldots$ with suspension in between $C_i^j$ and $C_i^{j+1}$.'' \cite[Section 3]{Raj:suspension1991}\footnote{The notation has been altered here for the sake of consistency.
 }
@@
 \begin{quote}
-	 Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
+	 ``Without any loss of generality, we shall assume that a task $\tau_i$ can defer its entire execution time but not parts of it.
 That is, a task $\tau_i$ executes for $C_i$ units with no suspensions once it begins execution.
@@
 The only difference is that if a task $\tau_i$ is split into two tasks $\tau_i'$ followed by $\tau_i''$, then $\tau_i''$ has the same deadlines as $\tau_i{{'}}$.
+''
 \end{quote}
@@
 
-Theorem 5 (in \cite{Raj:suspension1991}) is a strong result, as it implies that the period enforcer does not induce any deadline misses.
+Theorem 5 (in \cite{Raj:suspension1991}) is a strong result: it implies that the period enforcer does not induce any deadline misses.
 This seemingly enables a powerful analysis approach: if the corresponding transformed set of single-segment deferrable tasks can be shown to be schedulable  \emph{without} period enforcement under fixed-priority scheduling using \emph{any} applicable analysis (e.g.,~\cite{ABRTW:93}), then the period enforcer algorithm also yields a correct schedule.
@@
 \begin{enumerate}
-	\item There exist sporadic segmented self-suspending task sets that are schedulable under fixed-priority scheduling without any enforcement, but that are infeasible under period enforcement.
+	\item There exist sets of segmented self-suspending tasks that are schedulable under fixed-priority scheduling without any enforcement, but that are infeasible under period enforcement.
 This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, and not in the context of the original multi-segmented self-suspending task set.
@@
 
-Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the task-set transformation question: \emph{given a set of sporadic segmented self-suspending tasks $\tset$, how do we obtain a corresponding set of single-segment deferrable tasks $\tset'$ such that $\tset'$ is schedulable  only if $\tset$ is schedulable?} That is, as discussed in Section~\ref{sec:classic-analysis} the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert multi-segmented self-suspending tasks into corresponding sets of single-segment deferrable tasks, but it remains unclear in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished.
+Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the task-set transformation question: \emph{given a set of segmented self-suspending tasks $\tset$, how do we obtain a corresponding set of single-segment deferrable tasks $\tset'$ such that $\tset'$ is schedulable (without period enforcement)  only if $\tset$ is schedulable (with period enforcement)?} That is, as discussed in Section~\ref{sec:classic-analysis}, the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert multi-segmented self-suspending tasks into corresponding sets of single-segment deferrable tasks, but it is left undefined in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished.
 In Section \ref{sec:convert}, we make a pertinent observation.
@@
 \setcounter{enumi}{1}
-	\item It is an open problem whether it is possible to derive a single-segment deferrable task set corresponding to a given set of  multi-segmented self-suspending tasks in polynomial time.
-Recent findings by Nelissen et al.~\cite{ecrts15nelissen} can be applied, but their method takes exponential time.
+	\item How to  to derive a single-segment deferrable task set corresponding to a given set of  multi-segmented self-suspending tasks is an open problem.
+Recent findings by Nelissen et al.~\cite{ecrts15nelissen} can be applied in a special case, but their method takes exponential time (even in the  special case).
 \end{enumerate}
@@
 In the context of the above example, if we convert the multi-segmented suspending task $\tau_2$ into two single-segment deferrable tasks, called $\tau_2^1$ and $\tau_2^2$, where task $\tau_2^1$ never defers its execution and task $\tau_2^2$ defers its execution by at most \emph{$9$} time units, the resulting single-segment deferrable task set $\{\tau_1, \tau_2^1, \tau_2^2\}$ is in fact not schedulable under the given priority assignment: if a job of $\tau_1$ coincides with the arrival of a job of $\tau_2^2$ after it has maximally deferred its execution, the job of $\tau_2^2$ has a response time of $9 + 2 + 1$ time units, which exceeds its relative deadline of 11 time units.
-This shows that any restriction to single-segment deferrable tasks (i.e., assuming that ``a task $\tau_i$ can defer its entire execution time but not parts of it''~\cite{Raj:suspension1991}, recall Section~\ref{sec:classic-analysis}) inherently comes with a loss of generality.
+This shows that any restriction to single-segment deferrable tasks --- that is, assuming that ``[w]ithout any loss of generality [\ldots] a task $\tau_i$ can defer its entire execution time but not parts of it''~\cite{Raj:suspension1991} (recall Section~\ref{sec:classic-analysis}) --- does in fact come with a loss of generality.
 

### 2016-06-09 conclusion.tex
some edits to address reviewer concerns

@@
 
-Nevertheless, Theorem 5 in \cite{Raj:suspension1991} could be useful for handling self-suspending tasks (that do not use suspension-based locks) if there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm.
-However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.
+Nevertheless, the period enforcer algorithm \emph{per se}, and Theorem 5 in \cite{Raj:suspension1991},  could still prove to be useful for handling self-suspending tasks (that do not use suspension-based locks) if  \emph{efficient} schedulability tests or methods for constructing sets of single-segment deferrable tasks can be found.
+However, such tests or transformations have not yet been found  and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.
 

### 2016-06-09 locking.tex
some edits to address reviewer concerns

@@
 
-Notably, the examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a shared-memory locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors.
+Notably, the examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a \emph{shared-memory} locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors.
 One may wonder whether  effects similar to those described in in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the DPCP~\cite{RSL:88} or the DFLP~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}.
@@
 
+As a third possible interpretation in the shared-memory case, one could also exclude critical sections from period enforcement such that only the rest of the computation segment \emph{after} a critical section is subject to period enforcement (i.e., making critical sections immediately eligible to execute).\footnote{This interpretation does not fit the assumptions stated in~\cite{Raj:suspension1991,Raj:91}.} As in the case of distributed semaphore protocols, this interpretation breaks the feedback cycle highlighted in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2}, but still invalidates all existing blocking analyses as it artificially inflates the synchronization delay, just as if a single remote critical section could effectively block a job multiple times.
+
 Finally, it is worth noting that our examples can be trivially extended with lower-priority tasks to ensure that no processor idles before the described deadline misses occur.

### 2016-06-09 LITES/response.tex
edit response to reviewers

@@
 \documentclass[12pt]{article}
-\usepackage{fullpage}
+\usepackage[margin=2cm,a4paper]{geometry}
 
@@
 
- \def\myendproof{{\ \vbox{\hrule\hbox{%
-   \vrule height1.3ex\hskip0.8ex\vrule}\hrule }}\par}
- \newenvironment{proof}{\noindent{\bf Proof. }}{\myendproof}
+\newenvironment{response}[0]{\textcolor{red}{\emph{Response: }}\color{blue}}{\medskip}
 
-\newcommand{\response}[1]{\textcolor{red}{\emph{Response: }} \textcolor{blue}{#1}\vspace{0.25in}}
-\newcommand{\action}[1]{\textcolor{blue}{\emph{\bf Actions Taken: }} \emph{#1}\myendproof \vspace{0.25in}}
+\setlength{\parindent}{0cm}
+\setlength{\parskip}{6pt}
+
+\newcommand{\action}[1]{\textcolor{red}{\emph{\it Actions taken: }}{\color{blue}#1}\medskip}
 
@@
 
-{\bf Reviewer A}:
+\pagestyle{empty}
 
-	The paper discusses the application of the period enforcer algorithm to
-segmented self-suspending tasks in a unicore/partitioned-multicore settings.
-In particular, the authors contrived multiple scenarios and task sets to
-exemplify situations where (1) schedulability is achieved without using the
-period enforcer but deadlines are missed if the enforcer is used, and (2)
-the enforcer is not compatible with multicore resource sharing mechanisms in
-the sense that it invalidates the analyses currently employed. As an extra
-contribution, the authors also show that converting sporadic segmented
-self-suspending tasks into single-segment deferrable tasks (i.e. tasks with
-release jitters) incurs extra pessimism in the analysis (which means that
-the original task set may be schedulable whereas its corresponding
-deferrable tasks set is not), and doing the conversion in polynomial time
-is an open problem.
+\section*{Reviewer A}
 
-The problem is relevant and interesting and I believe the paper fits very
-well in the scope of Lites. The structure of the paper is good and the
-presentation is excellent, both in terms of language, and general structure.
-The paper is easy and pleasant to read and presents very pedagogic content.
-I warmly recommend its publication.
-
-{\bf Q:A.1}:    
-I only have a few very minor comments:\\
-- p2, line 60: assumptions on control flow $\rightarrow$ assumptions on
-the task control flow\\
-- p2, line 60: the segmented model is that is allows for $\rightarrow$ the
-segmented model is that **it** allows for\\
-- p4, line112: I would say segmented self-suspending tasks rather than
+\begin{itemize}
+	\item p2, line 60: assumptions on control flow $\rightarrow$ assumptions on
+the task control flow
+	\item p2, line 60: the segmented model is that is allows for $\rightarrow$ the
+segmented model is that **it** allows for
+	\item p4, line112: I would say segmented self-suspending tasks rather than
 multi-segment tasks to remain consistent with the naming introduced in
-the system model.\\
-- p7, line 212:  to enrich the period enforcer with the following
+the system model.
+	\item p7, line 212:  to enrich the period enforcer with the following
 power, consider replacing power with ``property'' or
-``scheduling rule''.\\
-- p7, line 221: ``it is immediately obvious that'' $\rightarrow$ I would simply
+``scheduling rule''.
+	\item p7, line 221: ``it is immediately obvious that'' $\rightarrow$ I would simply
 remove that part of the sentence since nothing is obvious with
@@
 
-\vskip 0.1in\noindent
-\response{Thanks for the suggestions and comments. We have rephrased those parts accordingly.}
+\begin{response}
+Thanks you for the suggestions. We have rephrased those parts accordingly.
 
-{\bf Reviewer B}:
+\end{response}
 
-	This work makes the following observations
-(i) period enforcement is not strictly superior to scheduling without
-enforcement
-(ii) period enforcement is incompatible with existing analysis of
-suspension-based locking protocols.
+\end{itemize}
 
-The observations described above are valid. However, there are some unstated
-assumptions about the period enforcer in how the technique is being applied,
-which should be clarified in this paper.
-
-{\bf Q:B.1}:    
-First, the period enforcer analysis only applies to the case where a task
+\section*{Reviewer B}
+\begin{itemize}
+	\item  First, the period enforcer analysis only applies to the case where a task
 defers its entire execution time but not parts of it. Tasks that suspend
@@
 is not schedulable.
-
-Section 3 should address the above clarification and the task $\tau_2$ should
+	\item Section 3 should address the above clarification and the task $\tau_2$ should
 be statically broken up before applying the period enforcer. It is true that
@@
 
-\response{
-Reviewer B was partially correct.
-Since the period enforcer analysis only applies to the case where a task defers its entire execution time but not parts of it, we should create the corresponding deferrable task set.
-However, it is not necessary that we need to create those deferrable task sets before using the period enforcer algorithm.
-Since the computation segments of a segmented self-suspending task are independent from each other, we can simply use the period enforcer algorithm without knowing the corresponding deferrable task set.
-This is actually how we demonstrate the examples in Section 2 and Section 3.
-However, there is a risk that the resulting schedule may not be feasible.
+\begin{response}
+We must distinguish two cases: the period enforcer \emph{runtime rules}, and the \emph{classic analysis} of the period enforcer.
+
+Section 3 is concerned only with the runtime rules.
+Specifically, it shows that the runtime rules may cause otherwise schedulable sets of multi-segment self-suspending tasks to become unschedulable.
+Importantly, the runtime rules do not require a conversion to single-segment deferrable tasks.
+
+The conversion to multi-segment deferrable tasks is relevant \emph{only} in the context of schedulability analysis. However, section 3 does not discuss schedulability analysis.
+
+To elaborate a bit more, since the period enforcer \emph{analysis} only applies to the case where a task defers its entire execution time but not parts of it, we should create the corresponding deferrable task set to carry out schedulability analysis.
+However, it is \emph{not} necessary to create those deferrable task sets for \emph{using} the period enforcer algorithm at runtime: since the computation segments of a segmented self-suspending task are independent from each other, we can simply use the period enforcer algorithm without knowing the corresponding deferrable task set.
+This is actually how we were able to create the examples in Section 2 and Section 3.
+
+%However, there is a risk that the resulting schedule may not be feasible.
 Therefore, one would need to verify the feasibility of the corresponding deferrable task set before applying the period enforcer algorithm.
@@
 Second, how should we perform the schedulability test on the corresponding deferrable task set correctly?
+\end{response}
+
+\action{Section 2 was rewritten and expanded to clearly distinguish between the runtime rule (Section 2.2) and the classic analysis (Section 2.3).
+The runtime rule --- Equation (1) --- is stated in terms of multi-segmented deferrable tasks.
 }
 
-\action{Section 4 was completely rewritten to be more clear.}
+\action{Section 4 was completely rewritten to more clearly explain that the problem of finding a suitable corresponding task set is open, and likely difficult.}
 
-{\bf Q:B.2}:    
+	\item 
 Section 4 should be reworded. 'precisely converting a self-suspending task
@@
 single-segment deferrable task set introduced pessimism due to period
-enforcement'. It is unclear how the Nelissen paper helps since it does not
+enforcement'.
+
+\begin{response}
+We do not fully understand this recommendation. In any case, we have completely rewritten this section, which we hope addresses this point as well.
+\end{response}
+
+\item 
+It is unclear how the Nelissen paper helps since it does not
 assume period enforcement. If each self-suspending task is broken into
@@
 
-\response{We do not fully capture the first sentence here from the reviewer.
-There is no additional pessimism due to the period enforcer algorithm.
+\begin{response}
+We need the analysis from Nelissen et al.\ to precisely break a self-suspending task into multiple deferrable subtasks so that we can calculate the worst-case release jitter of a computation segment.
+Section 4 explains why it is difficult to precisely calculate the release jitter of the computation segments and clearly points out the connection to Nelissen et al.'s analysis.
+
+%There is no additional pessimism due to the period enforcer algorithm.
 The pessimism is due to the assumption made by the period enforcer algorithm to allow only a set of deferrable tasks (that only suspend once at beginning).
-In the revision, we completely rewrote Section 4 to explain the gap between the segmented self-suspending task model and the corresponding deferrable sporadic task set.
-We need the analysis from Nelissen et al.
-to precisely break a self-suspending task into multiple deferrable subtasks so that we can calculate the worst-case release jitter of a computation segment.
-The new structure of Section 4 first explains why it is difficult to precisely calculate the release jitter of the computation segments and then explains that it is rather important to also perform the schedulability test correctly.
-We show that the schedulability test can be wrong if the test is not done well.
+\end{response}
+
+\action{In the revision, we completely rewrote Section 4 to explain the gap between the segmented self-suspending task model and the corresponding deferrable sporadic task set, and to precisely state the equivalence of the conversion to the worst-case response-time analysis of individual computation segments (i.e., Nelissen et al.'s analysis).
 }
 
-{\bf Q:B.3}:    
+	\item 
 Section 5 should also include the scenario where period enforcement is only
@@
 
-\response{ leave to Bj\"orn}
+\begin{response}
+Thank you for the suggestion; this is indeed an interesting option to explore. However, we believe this to be out of the scope of this paper.
 
-{\bf Reviewer C}:
+Prior work has explicitly suggested to use the period enforcer to control self-suspensions caused by the MPCP and the DPCP such that there is \emph{no} scheduling penalty, which is only possible if the period enforcer is applied to critical sections.
+Specifically, this is assumed in Rajkumar's 1991 book [27], and also mentioned in more recent works by his group.
+In fact, Rajkumar writes in the original period enforcer proposal (emphasis added):
 
-{\bf Q:C.1}:    
+\begin{quote}
+	``The scheduling penalty of deferred execution [when using synchronization protocols] can be enormous to the extent that it can make such protocols almost useless in several cases.
+\emph{The development of the period enforcer algorithm was primarily motivated by the need to reduce or eliminate this penalty} with a small, if not negligible, overhead.
+If tasks using these synchronization protocols use the period enforcer, their worst-case blocking durations would be considerably smaller and the schedulable utilization correspondingly higher.'' [26, Section 5.1]
+\end{quote}
+
+Our intention here is to document that this simply does not work. We do not claim to evaluate all possible ways in which the period enforcer might be used otherwise.
+
+\end{response}
+
+\action{Section 5.4 has been extended to briefly comment on this third possible interpretation. We now also comment on distributed semaphore protocols (e.g., DPCP, DFLP).}
+
+\end{itemize}
+
+\section*{Reviewer C}
+
+\begin{itemize}
+	\item   
 This reviewer was not able to download a complete version of [20] including
@@
 
-\response{We are not able to provide a complete version of the paper since the paper is not from us.
-We also used the same version (without figures) to study the period enforcer algorithm.
-In the revision, we have decided to detail more steps to improve the consistency of our paper.
-}
+\begin{response}We are not able to provide a complete version of the original paper since we are not its authors.
+In the preparation of the paper under review, we also relied on the same online version (without figures) to study the period enforcer algorithm.
+While figures are indeed missing, they can be inferred from the accompanying detailed  descriptions; the rest of the paper is not corrupted.
 
-{\bf Reviewer D}:
+\end{response}
 
-	The period enforcer algorithm (proposed by Rajkumar [20]) was designed to
-reduce the scheduling penalty caused by self-suspending tasks. This paper
-presents three important observations regarding the applicability of period
-enforcer algorithm and its schedulability analysis to multi-segment
-self-suspending tasks. First, it is shown that period enforcement may cause
-deadline miss in multi-segment self-suspending tasks that are schedulable
-without period enforcement. The original analysis of period enforcer
-algorithm [20] shows that if a set of single-segment (deferrable) tasks is
-schedulable, then the task set is also schedulable under period enforcement.
-Therefore, to apply this result from [20] to a set of multi-segment
-self-suspending tasks, a transformation from multi-segment to single-segment
-task set is needed. Second, it is shown that such transformation is quite
-difficult (i.e., currently only exponential time solution is known). Third,
-it is shown that period enforcer algorithm requires new analysis of existing
-suspension-based locking protocol to compute blocking factor. In other
-words, period enforcer algorithm is incompatible with all existing analyses
-of suspension-based locking protocols. All these three (negative)
-observations show the limitations of the applicability of period enforcer
-algorithm to multi-segment self-suspending task set. These limitations are
-demonstrated using examples that are easy to follow (Section 3, 4 and 5).
+\action{In the revision, we have expanded and improved the description of the background material in Section 2 to make our paper much more self-contained.}
 
-{\bf Q:D.1}:  
+\end{itemize}
+
+\section*{Reviewer D}
+
+\begin{itemize}
+	\item   
 This paper is heavily dependent on the assumptions and analysis of the
@@
 
-\response{leave this to Bj\"orn}
+\action{We have expanded the discussion of necessary background in Section 2. Sections 2.3 and 2.4 have been edited and clarified to make the discussion self-contained.}
 
-{\bf Q:D.2}:  
+	\item   
 There are some terms in section 2.3 and 2.4 that the authors have used
@@
 
-\response{leave this to Bj\"orn as well}
+\action{We added Section 2.2, which reviews in detail all relevant task models  and clearly defines the terminology used in the paper.}
 
-{\bf Q:D.3}:  
+	\item   
 Theorem 5 (page 5) mentions about “worst-case” condition. What does it
@@
 
-\response{leave this to Bj\"orn as well.
-message from JJ:
-Can you think how to improve that? I think a table would be nice. We may need to put more words from Raj’s paper.. 
-But, let’s assume that Raj decides to take the paper down from his website. Then, no one knows what we are talking about precisely. So, maybe it is a good idea to make Section 2 more self-contained?
-}
+\action{We added an explanation of the term ``worst-case conditions'' immediately after Theorem 5.}
 
-{\bf Q:D.4}:  
+	\item  
 Is there any way to quantify (theoretically or using simulation) the
@@
 
-\response{We are not able to provide the (theoretical or simulated) loss/gain due to the following reasons.
-1) We are not able to convert a segmented self-suspending task set into the corresponding deferrable task set efficiently and without any information loss.
-In Section 4.1, we explain the task set transformation for a special case.
-If there are multiple self-suspending tasks, there is no existing solution to exactly construct the deferrable task set.
-2) There is no clear evidence on how such schedulability tests for the corresponding deferrable task set should be designed precisely.
+\begin{response}We are not able to provide a meaningful empirical characterization of the loss/gain because there is neither a precise baseline to compare against, nor a practical analysis for the period enforcer.
+\begin{enumerate}
+	\item We would be curious to learn this, too, but unfortunately we are not able to convert a segmented self-suspending task set into the corresponding deferrable task set efficiently and without  information loss.
+In Section 4, we explain the task set transformation for a limited special case.
+If there are multiple self-suspending tasks, there is no existing solution for exactly constructing the deferrable task set: the original proposal [26] failed to define such a conversion, and we did not find one that is both efficient and not too pessimistic.
+
+%\item There is no clear evidence on how such schedulability tests for the corresponding deferrable task set should be designed precisely.
 We have demonstrated in Section 4.2 that all the deferrable tasks (even constructed from the same self-suspending task) should be considered for analyzing the interference.
@@
 Based on the example provided in Section 4.2, one can imagine that the calculation of the deferrable time already considers the earlier computation segments and the calculation of the worst-case interference has to account for the earlier computation segments as well (even if there is no overlap).
-We are not sure whether this is due to the design or due to the pessimistic analysis.
-Please refer to Section 4.2 for this.
-3) There are also other methods that have been developed recently to handle the self-suspensions, and we are not sure what to compare with.
-}
+%We are not sure whether this is due to the design or due to the pessimistic analysis. Please refer to Section 4.2 for this.
 
-\action{Section 4 has been completely rewritten to explain the difficulty.}
+\item There are also no other known methods or analyses to handle self-suspensions in the general case without incurring a potentially large degree of pessimism; we are hence not sure what to use as a fair baseline.
+\end{enumerate}
 
-{\bf Q:D.5}:   
+\end{response}
+
+\action{Section 4 has been rewritten to  more clearly explain the difficulty of analyzing self-suspensions (both with or without the period enforcer).}
+
+\end{itemize}
+
+\begin{itemize}
+	\item 
 Based on the notes of this paper, it is not clear whether one should consider period enforcer algorithm or not for self-suspending tasks.
 
-\response{From our note, we can only support the period enforcer algorithm as stated in the conclusion: ``\emph{Nevertheless, Theorem 5 in [21] could be useful for handling self-suspending tasks (that do not use suspension-based locks) if there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm.
-However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.}'' However, there are several drawbacks that were never discussed earlier, as pointed out at the end of Section 1.
-The period enforcer algorithm can still be potentially used with cares.
-}
+\begin{response}
+	There is currently no known schedulability analysis for the period enforcer (if any task has more than one computation segment).
+Until future work finds such an analysis, the period enforcer should not be considered as a solution for handling self-suspensions.
+\end{response}
 
-{\bf Q:D.6}:   
+\action{We have rephrased the last paragraph of the conclusion to make this point more clearly.}
+
+	\item    
 Why the dynamic self-suspension model is presented and compared with
@@
 
-\response{There was no reference to support this claim. We explain the reason in Section 2.3 in the revision.}
+\action{We added Section 2.2.4 to explain this limitation.}
+\end{itemize}
 
-\action{The following paragraph is added to Section 2.3.
-Note that it is not possible to convert a dynamic self-suspending task into the notion of deferred execution defined in [21], since the definition of a dynamic self-suspending task creates dynamic execution amounts for different jobs of a task.
-A simple example can explain why the period enforcer algorithm is not compatible with the dynamic self-suspending task model.
-Suppose that the system has only one task with execution time $C_1=1$, $S_1=1$, and $D_1=T_1=2$.
-The first job of task $\tau_1$ arrives at time $0$, suspends itself for one time unit, and then executes for one time unit.
-The second job of task $\tau_1$ arrives at time $2$, first executes for $0.5$ time unit, then suspends for $1$ time unit, and executes for $0.5$ time unit.
-With the period enforcer algorithm, the second job of task $\tau_1$ starts its execution at time $3$ and will clearly miss the deadline at time $4$.}
 \end{document}

### 2016-06-09 transformation.tex
fix proofreading comments

@@
 %
-Based on the presented considerations, we conclude that filling in this missing step is non-trivial and observe that the closest known solution by Nelissen et al.
-\cite{ecrts15nelissen} requires exponential time even in the greatly simplified special case of a single self-suspending task.
+Based on the presented considerations, we conclude that filling in this missing step is non-trivial and observe that the closest known solution by Nelissen et al.~\cite{ecrts15nelissen} requires exponential time even in the greatly simplified special case of a single self-suspending task.
 It thus remains unclear how Theorem~5 in~\cite{Raj:suspension1991} can be used for schedulability analysis of sets of multi-segmented self-suspending tasks.

### 2016-06-09 untitled.tex
fix proofreading comments

@@
 However, even in the presence of release jitter, a key assumption remains that jobs do not self-suspend (e.g., wait for I/O).\footnote{Audsley et al.~\cite{ABRTW:93} do present a response-time analysis that takes into account a limited form of suspensions due to semaphores (``blocking'').
-However, their analysis does not apply to general self-suspensions (i.e., the kind of self-suspensions targeted by period enforcer algorithm)  and is not relevant in the context of this paper.} That is, Audsley et al.~\cite{ABRTW:93} assume that, once a job has arrived, it continuously remains available for dispatching until it completes.
+However, their analysis does not apply to general self-suspensions (i.e., the kind of self-suspensions targeted by the period enforcer algorithm)  and is not relevant in the context of this paper.} That is, Audsley et al.~\cite{ABRTW:93} assume that, once a job has arrived, it continuously remains available for dispatching until it completes.
 This restriction is removed next.
@@
 The first self-suspension segment $S_i^0$, prior to the first execution segment, is equivalent to release jitter (i.e., the parameter $J_i$ in Section~\ref{sec:jitter}).
-However, in in much of the literature on the segmented self-suspending task model, the segment $S_i^0$ assumed to be absent (i.e., $S_i^0 = 0$), such that there are only $m_i - 1$ suspension intervals (and jobs arrive jitter-free).
+However, in much of the literature on the segmented self-suspending task model, the segment $S_i^0$ is assumed to be absent (i.e., $S_i^0 = 0$), such that there are only $m_i - 1$ suspension intervals (and jobs arrive jitter-free).
 Unless noted otherwise we adopt this convention.
@@
   \includegraphics[scale=1]{../figures/not-ok-without-period-enforcer/not-ok.pdf}
-  \caption{Example uniprocessor schedule (\emph{without} period enforcement) of three tasks $\tau_1 = $, $\tau_2$, and $\tau_3$ with periods $T_1 = T_2 = T_3 = 10$.
+  \caption{Example uniprocessor schedule (\emph{without} period enforcement) of three tasks $\tau_1$, $\tau_2$, and $\tau_3$ with periods $T_1 = T_2 = T_3 = 10$.
 Tasks $\tau_1$ and $\tau_3$ consist of a single computation segment ($C_1^1 = C_3^1 = 3$); task $\tau_2$ consists of two computation and one suspension segment ($C_2^1 = 1$, $S_2^1 = 4$, $C_2^2 = 2$).
@@
 Any task that does suspend after it executes for a while can be considered to be two or more tasks each with its own worst-case execution time.
-The only difference is that if a task $\tau_i$ is split into two tasks $\tau_i'$ followed by $\tau_i''$, then $\tau_i''$ has the same deadlines as $\tau_i{{'}}$.
-''
+The only difference is that if a task $\tau_i$ is split into two tasks $\tau_i'$ followed by $\tau_i''$, then $\tau_i''$ has the same deadlines as $\tau_i{{'}}$.''
 \end{quote}
@@
 \setcounter{enumi}{1}
-	\item How to  to derive a single-segment deferrable task set corresponding to a given set of  multi-segmented self-suspending tasks is an open problem.
+	\item How to derive a single-segment deferrable task set corresponding to a given set of  multi-segmented self-suspending tasks is an open problem.
 Recent findings by Nelissen et al.~\cite{ecrts15nelissen} can be applied in a special case, but their method takes exponential time (even in the  special case).

### 2016-06-09 untitled.tex
more proofreading fixes

@@
 To date, the real-time literature on self-suspensions has focused on two task models: the \emph{dynamic} self-suspension model, which we discuss first,  and the  (\emph{multi}-)\emph{segmented} suspension model, which we discuss next in Section~\ref{sec:segmented}.
-Self-suspensions can arise in both periodic and sporadic tasks (i.e, both interpretations of the $T_i$ parameter are possible).
+Self-suspensions can arise in both periodic and sporadic tasks (i.e., both interpretations of the $T_i$ parameter are possible).
 The observations that we make in this note apply equally to both periodic and sporadic tasks; for convenience, we focus primarily on periodic tasks.
@@
 The root cause is increased interference due to the ``back-to-back'' execution effect~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}.
-In the example shown in Figure \ref{fig:not-ok-without-period-enforcement}, two jobs of $\tau_2$ execute in close succession (i.e, separated by less than a period) because the second job, released at time~10, self-suspended for a (much) shorter duration than the first job.
+In the example shown in Figure \ref{fig:not-ok-without-period-enforcement}, two jobs of $\tau_2$ execute in close succession (i.e., separated by less than a period) because the second job, released at time~10, self-suspended for a (much) shorter duration than the first job.
 Consequently,  $\tau_3$ suffers from increased interference when $\tau_2$'s second job resumes ``too soon'' at time~12 after having been suspended for only one time unit, rather than four time units like the first job of $\tau_2$.

### 2016-06-09 conclusion.tex
update conclusion

@@
 One key assumption in the original proposal \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it.
-This creates some mismatches between the original self-suspending task set and the corresponding deferrable task set, which we have demonstrated with an example that shows that Theorem 5 in \cite{Raj:suspension1991} does not reflect the schedulability of the original self-suspending task system.
+This creates some mismatches between the original segmented self-suspending task set and the corresponding single-segment deferrable task set, which we have demonstrated with an example that shows that Theorem 5 in \cite{Raj:suspension1991} does not reflect the schedulability of the original segmented self-suspending task system.
 
-The original proposal \cite{Raj:suspension1991} further left open the question of how to convert a segmented self-suspending task set to a corresponding set of deferrable tasks.
+The original proposal \cite{Raj:suspension1991} further left open the question of how to convert a segmented self-suspending task set to a corresponding set of single-segment deferrable tasks.
+This problem remains open.
 Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a transformation is non-trivial in the general case.
@@
 Nevertheless, the period enforcer algorithm \emph{per se}, and Theorem 5 in \cite{Raj:suspension1991},  could still prove to be useful for handling self-suspending tasks (that do not use suspension-based locks) if  \emph{efficient} schedulability tests or methods for constructing sets of single-segment deferrable tasks can be found.
-However, such tests or transformations have not yet been found  and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.
+However, such tests or transformations have not yet been obtained  and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.
 

### 2016-06-09 abstract.tex
Update second claim

--- 2015-10-17 abstract.tex
@@
 Originally proposed in 1991, the algorithm has attracted renewed interest in recent years.
-This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~to match the assumptions underlying the analysis of the period enforcer, a schedulability analysis of self-suspending tasks subject to period enforcement requires a task set  transformation that, with current techniques, is subject to exponential time complexity; and \textbf{(iii)}~the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
+This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior (compared to the base case without enforcement) as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~to match the assumptions underlying the analysis of the period enforcer, a schedulability analysis of self-suspending tasks subject to period enforcement requires a task set  transformation for which no solution is known  in the general case, and which is subject to exponential time complexity (with current techniques) in the limited case of a single self-suspending task; and \textbf{(iii)}~the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
 \end{abstract}

### 2016-06-09 untitled.tex
Update second claim

@@
 	\item period enforcement can be a cause of deadline misses in self-suspending task sets that are otherwise schedulable (Section~\ref{sec:unschedulable}); 
-	\item to match the assumptions underlying the analysis of the period enforcer, a schedulability analysis of self-suspending tasks subject to period enforcement requires a task set  transformation that, with current techniques, is subject to exponential time complexity (Section~\ref{sec:convert}); and
+	\item to match the assumptions underlying the analysis of the period enforcer, a schedulability analysis of self-suspending tasks subject to period enforcement requires a task set  transformation for which no solution is known  in the general case, and which  is subject to exponential time complexity (with current techniques)  in the limited case of a single self-suspending task (Section~\ref{sec:convert}); and
 	\item the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed (Section~\ref{sec:locking}).

### 2016-06-12 LITES/response.tex
minor changes in the response letter

@@
 \begin{response}
-Thanks you for the suggestions. We have rephrased those parts accordingly.
+Thank you for the suggestions. We have rephrased those parts accordingly.
 
@@
 \begin{response}
-We do not fully understand this recommendation. In any case, we have completely rewritten this section, which we hope addresses this point as well.
+We do not fully understand this recommendation. In any case, we have
+completely rewritten this section, which we hope that our revision addresses this point as well.
 \end{response}

### 2016-06-12 locking.tex
minor revision

@@
 
+In this section, we take a look at the period enforcer in the context of synchronization protocols.
 Binary semaphores, i.e., suspension-based locks used to realize mutually exclusive access to shared resources, are a common source of self-suspensions in multiprocessor real-time systems.
@@
 Notably, the examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a \emph{shared-memory} locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors.
-One may wonder whether  effects similar to those described in in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the DPCP~\cite{RSL:88} or the DFLP~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}.
+One may wonder whether  effects similar to those described in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the DPCP~\cite{RSL:88} or the DFLP~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}.
 In this case, the self-suspension occurs on the task's \emph{application processor}, which is different from the (remote) synchronization processor on which the critical section is executed.

### 2016-06-12 transformation.tex
minor revision

@@
 To  convert a computation segment of $\tau_k$ into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible arrival time}, relative to the release of a job.
-Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than at time $t+\rho_k^j$.
+Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will not arrive later than time $t+\rho_k^j$.
 
@@
 Furthermore, Nelissen et al.
-\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.\footnote{In fact, in ongoing work, this problem has recently been shown to be coNP-hard in the strong sense~\cite{Chen2016b}.}
+\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.\footnote{In fact, in ongoing work, verifying the schedulability of task $\tau_k$ even for this case has recently been shown to be coNP-hard in the strong sense~\cite{Chen2016b}.}
 
@@
 
-Next, we take a look at the period enforcer in the context of synchronization protocols.
-
 %%% Local Variables:

### 2016-06-12 untitled.tex
minor revision

@@
 We focus exclusively on preemptive fixed-priority scheduling in this note, as the period enforcer is explicitly designed for this setting.
-For simplicity, we assume that tasks are indexed in order of decreasing priority (i.e., $\tau_1$ is the highest-priority task).
+For simplicity, we assume that the given tasks are indexed in a decreasing order with respect to their priorities (i.e., $\tau_1$ is the highest-priority task).
 
@@
 
-Importantly, the tasks in $\tset$ and $\tset'$ do \emph{not} have to be of the same task model, nor does the number of tasks have to remain constant (i.e.,  $|\tset| \neq |\tset'|$ is possible).
+Importantly, the tasks in $\tset$ and $\tset'$ do \emph{not} have to be of the same task model, nor does the number of tasks have to remain the same (i.e.,  $|\tset| \neq |\tset'|$ is possible).
 Specifically, the task set transformation underlying the analysis of the period enforcer maps each \emph{multi-}segmented self-suspending task $\tau_i \in \tset$  to $m_i$ \emph{single-}segmented self-suspending tasks in $\tset'$ (i.e., $|\tset'| = \sum_{\tset} m_i$).
@@
 Further, let $a^k_{i,j}$ denote the segment's arrival time.
-Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$ (i.e., the processor executes only $\tau_i$ or higher-priority tasks throughout the interval $[\mathit{busy}(\tau_i, t'), t']$).
+Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval begins on or prior to time $t'$ (i.e., the processor executes only $\tau_i$ or higher-priority tasks throughout the interval $[\mathit{busy}(\tau_i, t'), t']$).
 The period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
@@
 
-Figure \ref{fig:ok-with-period-enforcement} illustrates how the definition of eligibility time in Equation~\ref{eq:ET-def} restores the schedulability of the task set depicted in Figure~\ref{fig:not-ok-without-period-enforcement}.
+Figure \ref{fig:ok-with-period-enforcement} illustrates how the definition of eligibility time in Equation~\eqref{eq:ET-def} restores the schedulability of the task set depicted in Figure~\ref{fig:not-ok-without-period-enforcement}.
 Consider the eligibility times of the second segment of task $\tau_2$.
@@
 
-This leads to the first question: \emph{Does schedulability of the  original set of segmented self-suspending tasks (without period enforcement) imply schedulability under period enforcement?} That is, can Theorem 5 (in \cite{Raj:suspension1991}) be generalized to multi-segmented self-suspending tasks?
+This leads to the first question: \emph{Does the schedulability of the  original set of segmented self-suspending tasks (without period enforcement) imply the schedulability under period enforcement?} That is, can Theorem 5 (in \cite{Raj:suspension1991}) be generalized to multi-segmented self-suspending tasks?
 In Section \ref{sec:unschedulable}, we answer this question in the negative.
@@
 	\item There exist sets of segmented self-suspending tasks that are schedulable under fixed-priority scheduling without any enforcement, but that are infeasible under period enforcement.
-This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, and not in the context of the original multi-segmented self-suspending task set.
+This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, but not in the context of the original multi-segmented self-suspending task set.
 \end{enumerate}
 
-Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the task-set transformation question: \emph{given a set of segmented self-suspending tasks $\tset$, how do we obtain a corresponding set of single-segment deferrable tasks $\tset'$ such that $\tset'$ is schedulable (without period enforcement)  only if $\tset$ is schedulable (with period enforcement)?} That is, as discussed in Section~\ref{sec:classic-analysis}, the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert multi-segmented self-suspending tasks into corresponding sets of single-segment deferrable tasks, but it is left undefined in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished.
+Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the task-set transformation question: \emph{given a set of segmented self-suspending tasks $\tset$, how do we obtain a corresponding set of single-segment deferrable tasks $\tset'$ such that $\tset'$ is schedulable (without period enforcement)  only if $\tset$ is schedulable (with period enforcement)?} That is, as discussed in Section~\ref{sec:classic-analysis}, the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert a set of multi-segmented self-suspending tasks into a corresponding set of single-segment deferrable tasks, but it is left undefined in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished.
 In Section \ref{sec:convert}, we make a pertinent observation.

### 2016-06-12 transformation.tex
polishing edits

--- 2016-06-09 transformation.tex
@@
 To  convert a computation segment of $\tau_k$ into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible arrival time}, relative to the release of a job.
-Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than at time $t+\rho_k^j$.
+Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will not arrive later than at time $t+\rho_k^j$.
 
@@
 Furthermore, Nelissen et al.
-\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.\footnote{In fact, in ongoing work, this problem has recently been shown to be coNP-hard in the strong sense~\cite{Chen2016b}.}
+\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.\footnote{In fact, in ongoing work, it has recently been shown that verifying the schedulability of task $\tau_k$ is coNP-hard in the strong sense even in the considered simplified case~\cite{Chen2016b}.}
 

### 2016-06-12 untitled.tex
polishing edits

@@
 We focus exclusively on preemptive fixed-priority scheduling in this note, as the period enforcer is explicitly designed for this setting.
-For simplicity, we assume that the given tasks are indexed in a decreasing order with respect to their priorities (i.e., $\tau_1$ is the highest-priority task).
+For simplicity, we assume that tasks are indexed in order of decreasing priority (i.e., $\tau_1$ is the highest-priority task).
 
@@
 
-This leads to the first question: \emph{Does the schedulability of the  original set of segmented self-suspending tasks (without period enforcement) imply the schedulability under period enforcement?} That is, can Theorem 5 (in \cite{Raj:suspension1991}) be generalized to multi-segmented self-suspending tasks?
+This leads to the first question: \emph{If the  original set of segmented self-suspending tasks is schedulable without period enforcement, is it then also schedulable under period enforcement?} That is, can Theorem 5 (in \cite{Raj:suspension1991}) be generalized to multi-segmented self-suspending tasks?
 In Section \ref{sec:unschedulable}, we answer this question in the negative.

### 2016-06-13 locking.tex
polishing edits
improve figure placement
add examples for distributed case and third interpretation

--- 2016-06-08 locking.tex
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=1]{../figures/locking/legend.pdf}
+  \includegraphics[scale=1]{../figures/locking/sched-locking-alt2.pdf}
+  \caption{Example schedule of two tasks $\tau_1$ and $\tau_2$ on two processors sharing one lock-protected resource.
+The example assumes that lock requests take effect immediately, even if the critical section segment is not yet eligible to be scheduled according to the rules of the period enforcer algorithm.
+Under this interpretation, the third job of task $\tau_1$ misses its deadline at time $24$.}
+  \label{fig:locking-alt2}
+\end{figure}
+
 At time $1-\epsilon$, the first job of $\tau_2$ acquires the shared resource because $\tau_1$ does not issue its request until time $1$.
@@
 
+\subsection{Other Protocols and Interpretations}
+
+The examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a \emph{shared-memory} locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors.
+One may wonder whether  effects similar to those described in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the DPCP~\cite{RSL:88} or the DFLP~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}.
+In this case, the self-suspension occurs on the task's \emph{application processor}, which is different from the (remote) synchronization processor on which the critical section is executed.
+
+This separation allows employing period enforcement  only on application processors (while avoiding it on synchronization processors) without incurring the feedback cycle between blocking times and self-suspension times highlighted in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2}.
+
+However, period enforcement still invalidates all existing blocking analyses for distributed real-time semaphore protocols~\cite{RSL:88,Raj:91,Br:13} because it artificially increases blocking times if tasks contain multiple accesses to shared resources.
+An example demonstrating this effect is shown in Figure~\ref{fig:locking-dpcp}.
+Two segmented self-suspending tasks $\tau_1$ and $\tau_2$ share a resource using a distributed real-time locking protocol.
+The choice of protocol is irrelevant; the example works with both the DPCP and the DFLP.
+The tasks have parameters $m_1 = 2$, $C_1^1 = C_1^2 = 1$, $T_1 = D_1 = 25$ and $m_2 = 3$, $C_2^1 = C_2^3 = 1$, $C^2_2 = 6$, and $T_2 = D_2 = 15$.
+The computation segments are separated by self-suspensions that arise while the tasks wait for the completion of critical sections that are executed remotely on a dedicated synchronization processor $P_{sync}$; the corresponding suspension segment parameters $S_1^1$, $S_2^1$, and $S_2^2$ will be defined shortly.
+
+The first jobs of $\tau_1$ and $\tau_2$ are both released at time~0 and attempt to access the shared resource at time~1.
+Task $\tau_1$'s request is serviced first; as a result $\tau_2$ resumes only at time $a^2_{2,1} = ET_{2,1}^2 = 5$ after having been suspended for four time units:
+\begin{align*}
+	ET_{2,1}^2 & = \max\left(ET_{2,0}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,1})\right) =\max(-T_2 + T_2,\ 5) = 5.
+\end{align*}
+Task $\tau_2$ then executes its second computation segment for $C_2^2 = 6$ time units until time~11, when the job accesses the shared resource for a second time.
+Since there is no contention from $\tau_1$ at this time, $\tau_2$ resumes after only two time units at time~13.
+This leaves the job sufficient time to complete at time~14, one time unit before its deadline at time~15.
+
+The second job of $\tau_2$ is released at time~15 and issues a request for the shared resource at time~16.
+Since there is no contention from $\tau_1$ at the time, the second computation segment arrives already at time $a^2_{2,2} = 18$, after having been self-suspended for only two time units.
+However, since the second segment of the first job arrived at time $a^2_{2,1} = 5$, the second segment of the second job is not eligible to start execution until time $ET_{2,2}^2 = 20$ since
+\begin{align*}
+	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,2})\right) =\max(5 + T_2,\ 18) = 20.
+\end{align*}
+As a result, $\tau_2$ faces contention from $\tau_1$ when it issues its second request for the shared resource at time~26, which ultimately leads to a deadline miss at time~30.
+
+In contrast, without period enforcement, $\tau_2$ does not miss its deadline at time~30 because, \emph{across its two requests}, a job of $\tau_2$ is delayed by at most one request of $\tau_1$, for a total self-suspension time of at most two six time units.
+That is, even though the \emph{individual} self-suspension segments of the two tasks are each up to four time units long (i.e., $S_1^1 = S_2^1 = S_2^2 = 4$), the fact that self-suspensions arise due to the same cause (resource contention) means that the \emph{total} self-suspension time is actually less than the sum of the individual per-segment bounds.
+
+ Existing analyses for the DPCP~\cite{RSL:88,Raj:91,Br:13} and the DFLP~\cite{Br:13} exploit this knowledge and therefore predict task $\tau_2$ to be schedulable with a worst-case response time of 14.\footnote{The analyses in \cite{RSL:88,Raj:91} do assume a segmented task model, but bound the total blocking across all segments.
+The analysis in \cite{Br:13} also bounds the total blocking across all segments and can be applied to both the segmented and the dynamic self-suspension model.} The example in Figure~\ref{fig:locking-dpcp} thus demonstrates that the period enforcer invalidates existing blocking bounds for distributed semaphore protocols.
+As an aside, the example in Figure~\ref{fig:locking-dpcp} further highlights limitations of the segmented self-suspension model in the context of synchronization protocols, where the lengths of self-suspensions encountered at runtime are inherently not independent.
+
 \begin{figure}[t]
   \centering
-  \includegraphics[scale=1]{../figures/locking/legend.pdf}
-  \includegraphics[scale=1]{../figures/locking/sched-locking-alt2.pdf}
+  \includegraphics[scale=1]{../figures/locking3/sched-locking-dpcp.pdf}
+  \caption{Example schedule of two tasks $\tau_1$ and $\tau_2$ on two processors sharing a \emph{remote} resource using a distributed semaphore protocol (e.g., the DPCP~\cite{RSL:88,Raj:91} or the DFLP~\cite{Br:13,Br:14}) together with the period enforcer.
+Since critical sections are executed as \emph{remote procedure calls} (RPCs) on a dedicated synchronization processor $P_{sync}$ (not subject to period enforcement), their execution is not delayed by period enforcement.
+However, period enforcement delays $\tau_2$ at time~18, which invalidates existing analyses~\cite{RSL:88,Raj:91,Br:13}: under both the DPCP and the DFLP, $\tau_2$ is predicted to have a maximum response time of~14~\cite{Br:13}, but with period enforcement, the second job of $\tau_2$ has in fact a response time of 16.
+  }
+  \label{fig:locking-dpcp}
+\end{figure}
+
+Returning to the shared-memory case,
+as a third possible interpretation, one could also exclude critical sections from period enforcement such that only the rest of the computation segment \emph{after} a critical section is subject to period enforcement (i.e., making critical sections immediately eligible to execute).\footnote{This interpretation does not fit the assumptions stated in~\cite{Raj:suspension1991,Raj:91}.} This can be understood as making each critical section an individual computation segment (exempt from period enforcement) that is separated from the following computation by a ``virtual'' self-suspension of maximum length zero.
+As in the case of distributed semaphore protocols, this interpretation breaks the feedback cycle highlighted in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2}, but still invalidates all existing blocking analyses as it artificially inflates the synchronization delay.
+
+An example of this effect is shown in  Figure~\ref{fig:locking-alt3}, which depicts the same scenario as in Figure~\ref{fig:locking-dpcp} under the assumption that a shared-memory semaphore protocol is used (i.e., critical sections are executed locally by each job) and that critical sections are exempt from period enforcement.
+As in the distributed case, period enforcement induces a deadline miss, whereas existing blocking analyses~\cite{Br:13,LNR:09} exploit the fact that a remote critical section can block only once, thus arriving at a worst-case response time bound of 14 for $\tau_2$.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=1]{../figures/locking3/sched-locking-alt3.pdf}
   \caption{Example schedule of two tasks $\tau_1$ and $\tau_2$ on two processors sharing one lock-protected resource.
-The example assumes that lock requests take effect immediately, even if the critical section segment is not yet eligible to be scheduled according to the rules of the period enforcer algorithm.
-Under this interpretation, the third job of task $\tau_1$ misses its deadline at time $24$.}
-  \label{fig:locking-alt2}
+The example assumes that lock requests take effect immediately and that critical sections are exempt from period enforcement (i.e., period enforcement is applied only to any computation after a critical section).
+As in Figure~\ref{fig:locking-dpcp}, the second job of task $\tau_2$ misses its deadline at time $30$.
+  }
+  \label{fig:locking-alt3}
 \end{figure}
+
+To conclude, in both Figures~\ref{fig:locking-dpcp} and~\ref{fig:locking-alt3}, period enforcement has an effect  \emph{as if} a single remote critical section can block a given job multiple times, which is fundamentally incompatible with efficient blocking analysis~\cite{Br:13}.
 
@@
 In particular, our examples show that the addition of period enforcement invalidates all existing blocking analyses.
-They also suggest that devising a correct blocking analysis would be a substantial challenge due to the demonstrated feedback cycle between the period enforcer rules and blocking durations.
 
+If critical sections are subject to period enforcement, our examples also suggest that devising a correct blocking analysis would be a substantial challenge due to the demonstrated feedback cycle between the period enforcer rules and blocking durations.
 Fundamentally, the design of the period enforcer algorithm implicitly rests on the assumption that a segment \emph{can} execute as soon as it is eligible to do so.
@@
 
-Notably, the examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a shared-memory locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors.
-One may wonder whether  effects similar to those described in in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the DPCP~\cite{RSL:88} or the DFLP~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}.
-In this case, the self-suspension occurs on the task's \emph{application processor}, which is different from the (remote) synchronization processor on which the critical section is executed.
-This separation allows employing period enforcement  only on application processors (while avoiding it on synchronization processors) without incurring the feedback cycle between blocking times and self-suspension times highlighted in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2}.
-However, period enforcement still invalidates all existing blocking analyses for distributed real-time semaphore protocols~\cite{RSL:88,Raj:91,Br:13} because it artificially increases blocking times if tasks contain multiple accesses to shared resources.
-
 Finally, it is worth noting that our examples can be trivially extended with lower-priority tasks to ensure that no processor idles before the described deadline misses occur.
-It is also not difficult to extend the second example with a task on a third processor such that all segments of $\tau_1$ and $\tau_2$ are separated by a non-zero self-suspension.
+It is also not difficult to extend the examples in Figures~\ref{fig:locking-alt2} and~\ref{fig:locking-alt3} with a task on a third processor such that all critical sections of $\tau_1$ and $\tau_2$ are separated from their predecessor segments by a non-zero self-suspension.

### 2016-06-13 LITES/response.tex
polishing edits
update response to reviewers to mention new examples

@@
 We do not fully understand this recommendation. In any case, we have
-completely rewritten this section, which we hope that our revision addresses this point as well.
+completely rewritten this section; we hope that the current revision addresses this concern as well.
 \end{response}
@@
 \begin{response}
-Thank you for the suggestion; this is indeed an interesting option to explore. However, we believe this to be out of the scope of this paper.
+Thank you for the suggestion; this is indeed an interesting option to explore. However, we believe a comprehensive exploration of all possible interpretations to be out of the scope of this paper.
 
@@
 
-\action{Section 5.4 has been extended to briefly comment on this third possible interpretation. We now also comment on distributed semaphore protocols (e.g., DPCP, DFLP).}
+\action{We have added Section 5.4 to briefly comment on this third possible interpretation.
+We now also comment on distributed semaphore protocols (e.g., DPCP, DFLP).
+Illustrative examples are provided in Figures~7 and~8.}
 

### 2016-06-13 locking.tex
spell out DPCP and DFLP on first use

@@
 The examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a \emph{shared-memory} locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors.
-One may wonder whether  effects similar to those described in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the DPCP~\cite{RSL:88} or the DFLP~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}.
+One may wonder whether  effects similar to those described in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the \emph{Distributed Priority Ceiling Protocol} (DPCP)~\cite{RSL:88,Raj:91} or the \emph{Distributed FIFO Locking Protocol} (DFLP)~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}.
 In this case, the self-suspension occurs on the task's \emph{application processor}, which is different from the (remote) synchronization processor on which the critical section is executed.

### 2016-06-13 locking.tex
proofreading fixes

@@
 Such self-suspensions due to lock contention, just like any other self-suspension, result in deferred execution and thus can detrimentally affect a task's interference on lower-priority tasks.
-It may thus seem natural to apply the period enforcer  to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91}, stated as a motivation and possible use case in the original period enforcer proposal~\cite{Raj:suspension1991}, and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.
-As we show here, this is actually unsafe.} However, as we demonstrate with two examples, it is actually unsafe to apply period enforcement to lock-induced self-suspensions.
+It may thus seem natural to apply the period enforcer  to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91}, stated as a motivation and possible use case in the original period enforcer proposal~\cite{Raj:suspension1991}, and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}.} However, as we demonstrate with two examples, it is actually unsafe to apply period enforcement to lock-induced self-suspensions.
 
@@
 
-The examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a \emph{shared-memory} locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors.
+The examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a \emph{shared-memory} locking protocol: once a lock is granted, tasks execute their own critical sections on their assigned processors.
 One may wonder whether  effects similar to those described in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the \emph{Distributed Priority Ceiling Protocol} (DPCP)~\cite{RSL:88,Raj:91} or the \emph{Distributed FIFO Locking Protocol} (DFLP)~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}.

### 2016-06-13 transformation.tex
proofreading fixes

@@
 
-The original period enforcer proposal~\cite{Raj:suspension1991} is silent on this issue and does not spell out a procedure for converting  multi-segmented self-suspending task to a corresponding set of single-segment deferrable tasks.
+The original period enforcer proposal~\cite{Raj:suspension1991} is silent on this issue and does not spell out a procedure for converting a multi-segmented self-suspending task to a corresponding set of single-segment deferrable tasks.
 However, in our opinion, performing such a transformation without introducing additional pessimism is not at all easy in the general case.

### 2016-06-14 LITES/LITES-Paper.tex
Comments from Reviewer B.
take out unneeded macros (compilation trouble)
add remarks on DPCP/DFLP and some minor edits
reformat for Arxiv

@@
+\pdfoutput=1 
 \input{../preamble}
@@
 
-\papertrue 
-\title{A Note on the Period Enforcer Algorithm  for Self-Suspending Tasks}
+\title{A Note on the Period Enforcer Algorithm  \\ for Self-Suspending Tasks}
 
-\author[1]{Jian-Jia Chen}
-\author[2]{Bj\"orn B.\ Brandenburg}
-\affil[1]{TU Dortmund University, Germany\\
-  \texttt{jian-jia.chen@cs.uni-dortmund.de}}
-\affil[2]{Max Planck Institute for Software Systems (MPI-SWS)\\
-  \texttt{bbb@mpi-sws.org}}
-\authorrunning{Chen and Brandenburg} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
+\date{}
 
-\Copyright{Chen and Brandenburg}%mandatory. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+\author{Jian-Jia Chen \\ TU Dortmund University, Germany \\ \texttt{jian-jia.chen@cs.uni-dortmund.de} \and Bj\"orn B.\ Brandenburg \\ Max Planck Institute for Software Systems (MPI-SWS) \\ \texttt{bbb@mpi-sws.org}}
 
-\subjclass{Real-time systems, Real-time schedulability, Synchronization}% mandatory: Please choose 2012 ACM classifications from http://www.acm.org/about/class/2012 .
+%\author[1]{Jian-Jia Chen}
+%\author[2]{Bj\"orn B.\ Brandenburg}
+%\affil[1]{TU Dortmund University, Germany\\
+%  \texttt{jian-jia.chen@cs.uni-dortmund.de}}
+%\affil[2]{Max Planck Institute for Software Systems (MPI-SWS)\\
+%  \texttt{bbb@mpi-sws.org}}
+%\authorrunning{Chen and Brandenburg} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
+%
+%\Copyright{Chen and Brandenburg}%mandatory. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+%
+%\subjclass{Real-time systems, Real-time schedulability, Synchronization}% mandatory: Please choose 2012 ACM classifications from http://www.acm.org/about/class/2012 .
 E.g., cite as "Integrated circuits".
-\keywords{period enforcer, deferred execution, self-suspension, blocking}% mandatory: Please provide 1-5 keywords
+%\keywords{period enforcer, deferred execution, self-suspension, blocking}% mandatory: Please provide 1-5 keywords
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%EditorialOffice macros (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\Volume{XXX}
-\Issue{YYY}
-\DateSubmission{Date of submission}
-\DateAcceptance{Date of acceptance}
-\DatePublished{Date of publishing}
-\SectionAreaEditor{LITES section area editor}
-\DOI{10.4230/LITES.xxx.yyy.p}
+%%EditorialOffice macros (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\Volume{XXX}
+%\Issue{YYY}
+%\DateSubmission{Date of submission}
+%\DateAcceptance{Date of acceptance}
+%\DatePublished{Date of publishing}
+%\SectionAreaEditor{LITES section area editor}
+%\DOI{10.4230/LITES.xxx.yyy.p}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-06-14 arXiv/make-arxiv.sh

@@ -1,0 +1,31 @@
+#!/bin/bash
+
+OUT=submission
+FIGS=figures.txt
+
+mkdir -p submission
+mkdir -p check
+
+latexpand ../LITES/LITES-paper.tex > $OUT/paper.tex
+grep includegraphics $OUT/paper.tex | sort | uniq | sed -e 's/.*]{//' -e 's/}//' > $FIGS
+
+for F in `cat $FIGS`
+do 
+    cp -v "$F" $OUT
+done
+
+cp -v ../bibliography/biblio.bib $OUT
+
+
+sed -i.bak -e 's|../figures/.*/||' -e 's|../bibliography/||'  $OUT/paper.tex
+
+rm $OUT/paper.tex.bak
+
+cd $OUT
+latexmk -pdf paper.tex
+latexmk -pdf -c paper.tex
+
+mv -v paper.pdf ../check
+rm biblio.bib
+
+

### 2016-08-17 conclusion.tex
minor revision
transformation

@@
 This problem remains open.
-Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a transformation is non-trivial in the general case.
+Taking into account recent developments~\cite{Chen2016b,RTSS2016-suspension,ecrts15nelissen}, we have observed that such a transformation is non-trivial in the general case.
 

### 2016-08-17 transformation.tex
transformation

@@
 In the following, we illustrate the inherent difficulty of the problem by focusing on a special case to which we can apply a recent result of Nelissen et al.
-\cite{ecrts15nelissen}, which allows analyzing the exact worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity.
-Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} is exact under the following conditions:\footnote{We refer to the characteristics of the worst-case release pattern provided in Lemma 2 in \cite{ecrts15nelissen}.
-The exact worst-case response time can be obtained by exploring all  release patterns that satisfy these conditions.}
+\cite{ecrts15nelissen}, which allows analyzing the worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity.
+Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} can be applied under the following conditions:
 \begin{enumerate}
@@
 \end{enumerate}
+
+Even for the restrictive case above, the mixed-integer linear programming (MILP) approach by Nelissen et al. \cite{ecrts15nelissen} does not provide the exact worst-case response time. 
+However, if we apply the characteristics of the worst-case release pattern provided in Lemma 2 by Nelissen et al.
+\cite{ecrts15nelissen},  the exact worst-case response time can be calculated by exploring all  release patterns that satisfy these conditions.
+Detailed discussions can be found in a recent report by Chen \cite{Chen2016b}.
+In fact, it has recently been shown that verifying the schedulability of such a task set is coNP-hard in the strong sense even in the considered simplified case~\cite{RTSS2016-suspension}.
+Therefore, analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense by following the same proof by Chen \cite{RTSS2016-suspension}.
 
@@
 However, deriving an exact bound $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not easy: 
-even for the above ``simple'' case, Nelissen et al.'s solution~\cite{ecrts15nelissen} for calculating the exact worst-case response time requires exponential time complexity if $j \geq 2$.
+even for the above ``simple'' case, Nelissen et al.'s MILP solution~\cite{ecrts15nelissen} for calculating a safe upper bound of the worst-case response time requires exponential time complexity if $j \geq 2$.
+Deriving the exact $W_k^j$ can be achieved by exploring all  release patterns that satisfy the conditions listed in Lemma 2 in \cite{ecrts15nelissen}, with very high time complexity. 
 Furthermore, Nelissen et al.
-\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.\footnote{In fact, in ongoing work, it has recently been shown that verifying the schedulability of task $\tau_k$ is coNP-hard in the strong sense even in the considered simplified case~\cite{Chen2016b}.}
+\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.
 
-Nelissen et al.~\cite{ecrts15nelissen} did not study the period enforcer; rather, they considered unrestricted self-suspensions.
+Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} did not study the period enforcer; rather, they considered unrestricted self-suspensions.
 However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that these observations transfer to the period enforcement case.

### 2016-08-17 transformation.tex
change ref style

@@
 even for the above ``simple'' case, Nelissen et al.'s MILP solution~\cite{ecrts15nelissen} for calculating a safe upper bound of the worst-case response time requires exponential time complexity if $j \geq 2$.
-Deriving the exact $W_k^j$ can be achieved by exploring all  release patterns that satisfy the conditions listed in Lemma 2 in \cite{ecrts15nelissen}, with very high time complexity. 
+Deriving the exact $W_k^j$ can be achieved by exploring all  release patterns that satisfy the conditions listed in Lemma 2 by Nelissen et al. \cite{ecrts15nelissen}, with very high time complexity. 
 Furthermore, Nelissen et al.

### 2016-08-17 transformation.tex
remove by Chen

@@
 In fact, it has recently been shown that verifying the schedulability of such a task set is coNP-hard in the strong sense even in the considered simplified case~\cite{RTSS2016-suspension}.
-Therefore, analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense by following the same proof by Chen \cite{RTSS2016-suspension}.
+Therefore, analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense by following the same proof \cite{RTSS2016-suspension}.
 

### 2016-08-18 transformation.tex
remove a sentence
correct use of 'arrive/arrival' and 'release'
arrival = point in time relevant for period / inter-arrival time constraint
release = point in time that a job/segment is available for execution

@@
 
-To  convert a computation segment of $\tau_k$ into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible arrival time}, relative to the release of a job.
-Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will not arrive later than at time $t+\rho_k^j$.
+To  convert a computation segment of $\tau_k$ into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible release time}, relative to the arrival of the corresponding job.
+Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible release time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will be released at the latest at time $t+\rho_k^j$.
 
@@
 Deriving the exact $W_k^j$ can be achieved by exploring all  release patterns that satisfy the conditions listed in Lemma 2 by Nelissen et al. \cite{ecrts15nelissen}, with very high time complexity. 
-Furthermore, Nelissen et al.
+%Furthermore, Nelissen et al.
 \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.

### 2016-08-18 untitled.tex
busy period -> busy interval
clarify the adopted convention
correct use of 'arrive/arrival' and 'release'
arrival = point in time relevant for period / inter-arrival time constraint
release = point in time that a job/segment is available for execution

@@
 A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen}.
-This penalty, which is maximized when a task defers the completion of one job just until the release of the next job, can manifest as response-time increases and thus may lead to deadline misses.
+This penalty, which is maximized when a task defers the completion of one job just until the arrival of the next job, can manifest as response-time increases and thus may lead to deadline misses.
 
@@
 A straightforward generalization of the periodic task model is to introduce an explicit \emph{relative deadline} parameter $D_i$.
-In this case, each task is represented by a three-tuple $(C_i, T_i, D_i)$, with the interpretation that every job of $\tau_i$ must finish within $D_i$ time units after its release.
+In this case, each task is represented by a three-tuple $(C_i, T_i, D_i)$, with the interpretation that every job of $\tau_i$ must finish within $D_i$ time units after its arrival.
 Task $\tau_i$ is said to have an \emph{implicit deadline}  if $D_i = T_i$, a \emph{constrained deadline} if $D_i \leq T_i$, and an \emph{arbitrary deadline} otherwise.
@@
 However, the sporadic task model relaxes the inter-arrival constraint $T_i$ to specify a \emph{minimum} (rather than an exact) separation between jobs.
-In this interpretation, the first job is not necessarily released at time~0, and the exact release times of future jobs cannot be predicted, which is an appropriate modeling assumption for event-triggered tasks.
+In this interpretation, the first job does not necessarily arrive at time~0, and the exact arrival times of future jobs cannot be predicted, which is an appropriate modeling assumption for event-triggered tasks.
 
-On uniprocessors, the relaxation from periodic to sporadic job arrivals does not introduce additional pessimism:\footnote{Assuming that all periodic tasks synchronously release a job at time zero.} since any two jobs of a sporadic task $\tau_i$ are  known to be released \emph{at least} $T_i$ time units apart, the sporadic task model~\cite{Mo:83} still allows for schedulability analysis that is as accurate as Liu and Layland's analysis of periodic tasks~\cite{LL:73}.
+On uniprocessors, the relaxation from periodic to sporadic job arrivals does not introduce additional pessimism:\footnote{Assuming that all periodic tasks synchronously release a job at time zero.} since any two jobs of a sporadic task $\tau_i$ are  known to arrive \emph{at least} $T_i$ time units apart, the sporadic task model~\cite{Mo:83} still allows for schedulability analysis that is as accurate as Liu and Layland's analysis of periodic tasks~\cite{LL:73}.
 
-Mok retained the assumption that incomplete jobs are always ready for execution (i.e., no suspensions), and that jobs, once released, are \emph{immediately} available for execution.
+Mok retained the assumption that incomplete jobs are always ready for execution (i.e., no suspensions), and that jobs  are \emph{immediately} available for execution after their arrival.
 
@@
 
-In the presence of release jitter, the terms ``job arrival'' and ``job release,'' which are often used interchangeably, take on distinct meanings: a job's \emph{arrival time} denotes the point in time when it actually becomes available for execution, whereas a job's \emph{release time} is the instant that is relevant for the (minimum) inter-arrival time constraint.
-Any job of task $\tau_i$  \emph{arrives} at most $J_i$ time units after it is \emph{released}.
+In the presence of release jitter, the terms ``job arrival'' and ``job release,'' which are often used interchangeably, take on distinct meanings: a job's \emph{release time} denotes the point in time when it actually becomes available for execution, whereas a job's \emph{arrival time} is the instant that is relevant for the (minimum) inter-arrival time constraint.
+Any job of task $\tau_i$  is \emph{released} at most $J_i$ time units after it  \emph{arrives}.
 
@@
 However, even in the presence of release jitter, a key assumption remains that jobs do not self-suspend (e.g., wait for I/O).\footnote{Audsley et al.~\cite{ABRTW:93} do present a response-time analysis that takes into account a limited form of suspensions due to semaphores (``blocking'').
-However, their analysis does not apply to general self-suspensions (i.e., the kind of self-suspensions targeted by the period enforcer algorithm)  and is not relevant in the context of this paper.} That is, Audsley et al.~\cite{ABRTW:93} assume that, once a job has arrived, it continuously remains available for dispatching until it completes.
+However, their analysis does not apply to general self-suspensions (i.e., the kind of self-suspensions targeted by the period enforcer algorithm)  and is not relevant in the context of this paper.} That is, Audsley et al.~\cite{ABRTW:93} assume that, once a job is released, it continuously remains available for dispatching until it completes.
 This restriction is removed next.
@@
 In particular, release jitter and self-suspensions are not interchangeable concepts and it is not safe~\cite{Chen2016,ecrts15nelissen} to simply substitute $J_i$ with $S_i$ in Audsley et al.'s analysis~\cite{ABRTW:93}.
-(Nonetheless, under the dynamic suspension model, it is possible for jobs of self-suspending tasks to defer their entire execution requirement, so self-suspensions can be seen as a generalization of release jitter.)
+(Nonetheless, under the dynamic suspension model, it is possible for jobs of self-suspending tasks to defer their entire execution requirement, so self-suspension can be seen as a generalization of release jitter.)
 
@@
 However, in much of the literature on the segmented self-suspending task model, the segment $S_i^0$ is assumed to be absent (i.e., $S_i^0 = 0$), such that there are only $m_i - 1$ suspension intervals (and jobs arrive jitter-free).
-Unless noted otherwise we adopt this convention.
+Unless noted otherwise we adopt this convention and assume $S_i^0 = 0$.
 
-We say that a segment \emph{arrives} when it becomes available for execution.
-The first computation segment arrives immediately when the job is released (unless $S_i^0 \neq 0$); the second computation segment (if any) arrives when the job resumes from its first self-suspension, \textit{etc.}
+We say that a segment is \emph{released} when it becomes available for execution.
+The first computation segment is released immediately when the job arrives (unless $S_i^0 \neq 0$); the second computation segment (if any) is released when the job resumes from its first self-suspension, \textit{etc.}
 
@@
 
-\subsubsection{Assumptions, Busy Periods, and Task Set Transformations}
+\subsubsection{Assumptions, Busy Intervals, and Task Set Transformations}
 \label{sec:misc-defs}
@@
 \end{align*}
-since the arrival of $\tau_2$'s second segment (and the release of $\tau_1$) starts a new level-2 busy interval at time $a_{2,1}^2 = 5$.
+since the release of $\tau_2$'s second segment (and the arrival of $\tau_1$) starts a new level-2 busy interval at time $a_{2,1}^2 = 5$.
 The second segment of $\tau_2$'s first job is hence immediately eligible to execute; however, due to the presence of a pending higher-priority job, $\tau_2$ is not actually scheduled until time~8 (just as without period enforcement as depicted in Figure \ref{fig:not-ok-without-period-enforcement}).
 
-The second segment of the second job of $\tau_2$ arrives at time $a_{2,2}^2 = 12$. In this case, the segment is \emph{not} immediately eligible to execute since
+The second segment of the second job of $\tau_2$ is released at time $a_{2,2}^2 = 12$. In this case, the segment is \emph{not} immediately eligible to execute since
 \begin{align*}
@@
 \begin{itemize}
-	\item if the first segment of a job of $\tau_2$ is interfered with by $\tau_1$, then the second segment resumes at most after $9$ time units after the release of the job and the response time of task $\tau_2$ is hence $10$; otherwise,
-	\item  if the first segment of a job of $\tau_2$ is not interfered with by $\tau_1$, then the second segment resumes at most $7$ time units after the release of the job and hence the  response time of task $\tau_2$ is at most $10$ even if the second segment is interfered with by $\tau_1$.
+	\item if the first segment of a job of $\tau_2$ is interfered with by $\tau_1$, then the second segment resumes at most $9$ time units after the arrival of the job and the response time of task $\tau_2$ is hence $10$; otherwise,
+	\item  if the first segment of a job of $\tau_2$ is not interfered with by $\tau_1$, then the second segment resumes at most $7$ time units after the arrival of the job and hence the  response time of task $\tau_2$ is at most $10$ even if the second segment is interfered with by $\tau_1$.
 \end{itemize}
@@
 
-In contrast to the first job, the second job of task $\tau_2$ (which is released at time $11$) is affected by period enforcement.
-The first segment of the second job arrives at time $a^1_{2,2} = 11$, incurs interference for one time unit during $[11, 12)$, and suspends at time $13$.
+In contrast to the first job, the second job of task $\tau_2$ (which arrives at time $11$) is affected by period enforcement.
+The first segment of the second job is released at time $a^1_{2,2} = 11$, incurs interference for one time unit during $[11, 12)$, and suspends at time $13$.
 The  second segment of the second job hence resumes only at time $a^2_{2,2} = 19$.

### 2016-08-18 transformation.tex
improve flow in section 4

@@
 \cite{ecrts15nelissen}, which allows analyzing the worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity.
+Specifically,
 Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} can be applied under the following conditions:
@@
 
-Even for the restrictive case above, the mixed-integer linear programming (MILP) approach by Nelissen et al. \cite{ecrts15nelissen} does not provide the exact worst-case response time. 
-However, if we apply the characteristics of the worst-case release pattern provided in Lemma 2 by Nelissen et al.
-\cite{ecrts15nelissen},  the exact worst-case response time can be calculated by exploring all  release patterns that satisfy these conditions.
-Detailed discussions can be found in a recent report by Chen \cite{Chen2016b}.
-In fact, it has recently been shown that verifying the schedulability of such a task set is coNP-hard in the strong sense even in the considered simplified case~\cite{RTSS2016-suspension}.
-Therefore, analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense by following the same proof \cite{RTSS2016-suspension}.
+As an aside, it is interesting to note that, even for the restrictive case above,  Nelissen et al.'s approach \cite{ecrts15nelissen}, which is based on \emph{mixed-integer linear programming} (MILP), provides only an upper bound on the worst-case response time, and not necessarily the \emph{exact} worst-case response time.
+However, at least conceptually, the exact worst-case response time \emph{can} be obtained based on their analysis by combining it with an exhaustive search.
+Specifically, Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen} provides a characterization of the worst-case release pattern that yields a maximal response time.
+Thus, by exploring all  release patterns that satisfy the conditions stated in Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen}, an exact bound can be determined, with very high time complexity.
+A detailed discussion of this approach can be found in a recent report~\cite{Chen2016b}.
 
-For an arbitrary number of tasks $k \geq 2$, 
+Let us now return to the discussion of the task-set transformation needed to apply Theorem 5 in~\cite{Raj:suspension1991}. For an arbitrary number of tasks $k \geq 2$, 
 suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines (i.e., $D_i = T_i$ for all $\tau_i$).
@@
 
-However, deriving an exact bound $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not easy: 
-even for the above ``simple'' case, Nelissen et al.'s MILP solution~\cite{ecrts15nelissen} for calculating a safe upper bound of the worst-case response time requires exponential time complexity if $j \geq 2$.
-Deriving the exact $W_k^j$ can be achieved by exploring all  release patterns that satisfy the conditions listed in Lemma 2 by Nelissen et al. \cite{ecrts15nelissen}, with very high time complexity. 
-%Furthermore, Nelissen et al.
-\cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.
+However, deriving an exact bound $W_k^j$ for task $\tau_k$ and for $j=1,2,\ldots,m_k-1$ is not easy: 
+even for the above ``simple'' case, Nelissen et al.'s MILP solution~\cite{ecrts15nelissen} for calculating a safe upper bound on the worst-case response time requires exponential time complexity if $j \geq 2$.
+In fact, it has recently been shown that the problem of verifying the schedulability of such a task set is coNP-hard in the strong sense~\cite{RTSS2016-suspension}.
+Therefore, it follows analogously that analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense~\cite{RTSS2016-suspension}.
 
-Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} did not study the period enforcer; rather, they considered unrestricted self-suspensions.
+Notably, Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} do not consider the period enforcer; rather, their analyses pertain to unrestricted self-suspensions.
 However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that these observations transfer to the period enforcement case.
@@
 %
-Based on the presented considerations, we conclude that filling in this missing step is non-trivial and observe that the closest known solution by Nelissen et al.~\cite{ecrts15nelissen} requires exponential time even in the greatly simplified special case of a single self-suspending task.
+Based on the presented considerations, we conclude that filling in this missing step is non-trivial and observe that the closest known solution by Nelissen et al.~\cite{ecrts15nelissen} requires exponential time even in the greatly simplified special case of a single self-suspending task, and that the problem of verifying the schedulability of such a task set is in fact coNP-hard in the strong sense~\cite{RTSS2016-suspension}.
 It thus remains unclear how Theorem~5 in~\cite{Raj:suspension1991} can be used for schedulability analysis of sets of multi-segmented self-suspending tasks.

### 2016-08-18 transformation.tex
polishing edits

@@
 Specifically,
-Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} can be applied under the following conditions:
+Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen}, which is based on \emph{mixed-integer linear programming} (MILP), can be applied under the following conditions:
 \begin{enumerate}
@@
 
-As an aside, it is interesting to note that, even for the restrictive case above,  Nelissen et al.'s approach \cite{ecrts15nelissen}, which is based on \emph{mixed-integer linear programming} (MILP), provides only an upper bound on the worst-case response time, and not necessarily the \emph{exact} worst-case response time.
-However, at least conceptually, the exact worst-case response time \emph{can} be obtained based on their analysis by combining it with an exhaustive search.
+As an aside, it is interesting to note that, even for the restrictive case above,  Nelissen et al.'s MILP-based analysis~\cite{ecrts15nelissen} provides only an upper bound on the worst-case response time, and not necessarily the \emph{exact} worst-case response time.
+However, at least conceptually, an exact answer \emph{can} be obtained based on their analysis by combining it with an exhaustive search.
 Specifically, Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen} provides a characterization of the worst-case release pattern that yields a maximal response time.
-Thus, by exploring all  release patterns that satisfy the conditions stated in Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen}, an exact bound can be determined, with very high time complexity.
+Thus, by exploring all  release patterns that satisfy the conditions stated in Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen}, an exact bound can be determined (at great computational cost).
 A detailed discussion of this approach can be found in a recent report~\cite{Chen2016b}.
 
-Let us now return to the discussion of the task-set transformation needed to apply Theorem 5 in~\cite{Raj:suspension1991}. For an arbitrary number of tasks $k \geq 2$, 
+Let us now return to the discussion of the task-set transformation that is needed before Theorem 5 in~\cite{Raj:suspension1991} can be applied. For an arbitrary number of tasks $k \geq 2$, 
 suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines (i.e., $D_i = T_i$ for all $\tau_i$).
@@
 In fact, it has recently been shown that the problem of verifying the schedulability of such a task set is coNP-hard in the strong sense~\cite{RTSS2016-suspension}.
-Therefore, it follows analogously that analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense~\cite{RTSS2016-suspension}.
+It follows analogously that analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense~\cite{RTSS2016-suspension}.
 
 Notably, Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} do not consider the period enforcer; rather, their analyses pertain to unrestricted self-suspensions.
-However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that these observations transfer to the period enforcement case.
+However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that their observations transfer to the period enforcement case.
 

### 2016-08-18 untitled.tex
polishing edits

@@
 A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen}.
-This penalty, which is maximized when a task defers the completion of one job just until the arrival of the next job, can manifest as response-time increases and thus may lead to deadline misses.
+This penalty, which is maximized when a task defers the completion of one job just until the release of the next job, can manifest as response-time increases and thus may lead to deadline misses.
 
@@
 However, in much of the literature on the segmented self-suspending task model, the segment $S_i^0$ is assumed to be absent (i.e., $S_i^0 = 0$), such that there are only $m_i - 1$ suspension intervals (and jobs arrive jitter-free).
-Unless noted otherwise we adopt this convention and assume $S_i^0 = 0$.
+We adopt this convention and assume $S_i^0 = 0$ unless noted otherwise.
 

### 2016-08-18 transformation.tex
analyses -> results

@@
 
-Notably, Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} do not consider the period enforcer; rather, their analyses pertain to unrestricted self-suspensions.
+Notably, Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} do not consider the period enforcer; rather, their results pertain to unrestricted self-suspensions.
 However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that their observations transfer to the period enforcement case.
